### 2016-09-19 slides/slidesSection.tex
slides template

%%
%% The official i6 slide template _demo_
%% created by Philippe Dreuw and Thomas Deselaers, 2005
%%
%% $Id: slidesSection.tex,v 1.23 2008-06-20 12:42:51 dreuw Exp $
%%
%% Possible options for:
%%  - language       : 'english' (default), 'german'
%%  - page numbering : 'nonumber', 'lastpage' to enable automatic page n/m numbering, 
%%                     'userlastpage' to enable a user defined page n/m numbering using \LastPage command,  
%%                      or leave empty for normal page numbering (default)
%%  - itemize        : 'triangle' or leave empty for bullets (default)
%%  - page layout    : 'vertical' to enable vertical centering on each slide using \vfill ... \vfill (default)
%%                     'demo' to compile images in demo mode, i.e. image files are not necessary 
%%  - encoding       : 'utf-8' to enable utf encoding instead of latin1 (default)
%%  - tools          : 'noblackslide' to disable the black slide at the end which is linked with every slide title, e.g. to write a proof on the blackboard
%%
\documentclass[11pt,a4paper,landscape]{article}
\usepackage[english,userlastpage,triangle,utf-8]{NeyDreuwSlides_Oct08}

% math operators and symbols %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\nat}{\ensuremath{\rm I\!N}\xspace}
\newcommand*{\rel}{\ensuremath{\rm I\!R}\xspace}
\newcommand*{\argmin}{\ensuremath{\operatornamewithlimits{argmin}}\xspace}
\newcommand*{\argmax}{\ensuremath{\operatornamewithlimits{argmax}}\xspace}
\newcommand*{\congmod}{\ensuremath{\operatornamewithlimits{\cong}}\xspace}
\newcommand*{\invers}{\ensuremath{\frac{1}}\xspace}
\newcommand*{\ra}{\ensuremath{\Rightarrow}\xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Change TOC style - only display sections with page number in \ToC
\setcounter{tocdepth}{1}  % set tocdepth to 2 to view subsections in ToC
\usepackage{titletoc}
\titlecontents{section}[1em]
  {\itshape}
  {\thecontentslabel\enspace}
  {}
  {}
  [\addvspace{0.6pc}]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% flyspell can read the local ispelldict variable to automatically change the dictionary, 
%% e.g. german-new8, american, english, british, ...
%% 
%% Local IspellDict: american
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% custom packages
\usepackage{fancyvrb} %%% fancy verbatim to enable coloring within verbatim environments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Some example hacks work only in combination with the 'pdflatex -shell-escape'
% mode
\ifpdf
  \usepackage{xspace}
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EDIT P. Dreuw: load minitoc always in last position
\usepackage[insection]{minitoc}
\renewcommand{\stifont}{\normalsize\bf} 
\renewcommand{\stcfont}{\normalsize\bf}
\renewcommand{\stcSSfont}{\normalsize\bf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand*{\title}{\LaTeX\ Slides Template \\ for Talks at the i6}        % main title of the work (used for \TitlePage)
\renewcommand*{\titleshort}{Seminar Template}                                 % short title (used for \lfoot)
\renewcommand*{\occasion}{Seminar medical image processing -- MedBV05 \\ this occasion title can be very long } % (used for \TitlePage)
\renewcommand*{\occasionshort}{MedBV05 -- shorter for footline}               % short occasion title (used for \rfoot)
%\renewcommand*{\date}{11. April 2005}                                         % default is \today (used for \TitlePage and \rfoot)
\renewcommand*{\author}{\underline{Philippe Dreuw}, Thomas Deselaers}         % all the authors of the work, can be long (used for \TitlePage)
\renewcommand*{\authorshort}{Deselaers / Dreuw:\xspace}                       % all the authors of the work, should be short (used for \lfoot)
\renewcommand*{\email}{\url{{deselaers,dreuw}@i6.informatik.rwth-aachen.de}}  % all email address(es) of the authors (used for \TitlePage)
\renewcommand*{\mainauthor}{Philippe Dreuw}                                   % the author(s) who presented the work (used for \TitlePage)
\renewcommand*{\mainauthoremail}{\url{dreuw@i6.informatik.rwth-aachen.de}}    % presenter mail address(es) (used for \FinalPage)
\renewcommand*{\www}{http://www-i6.informatik.rwth-aachen.de/}                % web address (used for \TitlePage _and_ \FinalPage)
\newcommand*{\keywords}{Latex, Slides, Seminar}                               % keywords, can be used for PDF summary

% will be set into the PDF document summary
\hypersetup{
  pdftitle={\title}, 
  pdfsubject={\occasion},  
  pdfauthor={\author}, 
  pdfkeywords={\keywords}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\TitlePage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\dosecttoc
\ToC

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section{Introduction - Requirements - Recommendations} 
\secttoc
\vfill
This latex beamer slide style was created by Philippe Dreuw and Thomas
Deselaers and should be used for talks presented at the Lehrstuhl fuer
Informatik 6 at the RWTH Aachen University.
\vfill
Any requests or comments should be sent to
\begin{itemize}
\item \url{mailto:{deselaers,dreuw}@informatik.rwth-aachen.de}
\item \url{http://www-i6.informatik.rwth-aachen.de/}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Literature} 
\begin{itemize}
\item Talks presented at the
  \href{http://www-i6.rwth-aachen.de/}{\isechsicon}
  should always have a literature part in front of the talk. 
\item You should \red{not just copy} bibitems from your \texttt{*.bbl} files
  into a minipage and an \texttt{\textbackslash itemize} environment.
\item Depending on the public where you will present your talk you should adapt
  your literature slide.
\item you should describe the content of the SOTA literature with your
  own words
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\headline{Literature Examples} 
\vfill 
\begin{description}
\item [T.~Starner, J.~Weaver, A.~Pentland:] Real-time american sign-language
  recognition using desk and wearable computer based video. {\em PAMI 1998}.
  \begin{itemize}
  \item HMM based isolated sign language recognition. Explain the main content here.
  \end{itemize}
\item [Povey:] Papertitle. {\em ICASSP 2007}.
  \begin{itemize}
  \item something new in speech recognition. Explain the main content here.
  \end{itemize}
\item \cite{loupias:_wavel_salien_point_image_retriev}: Papertitle. {\em CONFNAME 2007}.
  \begin{itemize}
  \item image retrieval. Explain the main content here.
  \end{itemize}
\end{description}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{State of the Art} 
\begin{itemize}
\item Talks presented at the
  \href{http:://www-i6.rwth-aachen.de/}{\includegraphics[height=\baselineskip]{logos/i6-hks44}}
  should also always have a state of the art part in front of the talk.
\item Depending on the public where you will present your talk you should
adapt your state of the art slide.
\item This can be in relation with literature section
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Results} 
\vfill 
\begin{itemize}
\item Talks presented at the
  \href{http:://www-i6.rwth-aachen.de/}{\includegraphics[height=\baselineskip]{logos/i6-hks44}}
  should always have a result section somewhere at the end of the talk.
\item Depending on the public should highlight the
  advantages/disadvantages of your presented work in comparison to
  other results achieved by other groups on the same (publicly
  available) benchmark database
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Misc} 
\vfill 
\begin{itemize}
\item you should always check your slites fohr teipos!!!
\item use \eg \texttt{flyspell} within emacs or \texttt{ispell} from
  the command line
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section{The \texttt{HyperSlides} Style}
\secttoc
\vfill
The following slides will explain how to use the HyperSlides style.
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Style Parameters}
\vfill
TODO
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Including Images}
\begin{itemize}
\item You can include your images with the \\ 
  \texttt{\textbackslash includegraphics[options]\{<filename>\}} command. A short caption may be
  written with the \texttt{\textbackslash caption\{<name>\}} command. Examples:
  \begin{figure}
    \centering
    \begin{tabularx}{600pt}{Z Z Z}
      \includegraphics[height=2.5cm]{car}%
      \label{fig:images_car}% 
      \caption{Car}%
      &
      \includegraphics[height=2.5cm]{airplane}%
      \label{fig:images_airplane}% 
      \caption{Airplane}%
      &
      \includegraphics[height=2.5cm]{motorbike}
      \label{fig:images_motorbike}% 
      \caption{Motorbike}%
    \end{tabularx}
    \caption{Images With Caption below}
    \label{fig:images}% 
  \end{figure}
\item Images can be scaled relatively to the \texttt{\textbackslash textwidth} or \texttt{\textbackslash linewidth} of the slides:
  \begin{center}
    \includegraphics[width=.2\textwidth]{salientpoints}
    \includegraphics[width=.2\linewidth]{grid}
  \end{center}
\item You should always provide images in JPEG/PNG/PDF-- \alert{and} EPS/PS--format.
\item you can insert \texttt{xfig} images in pdftex or pstex manner by \\ using the 
  \texttt{\textbackslash inputfigure\{<filename>\}} \\ or the \texttt{\textbackslash
    inputfigurex\{<filename>\}\{<width>\}\{<height>\}}
\item Example:
  \inputfigure{xfigures/xfigtest} and \inputfigurex{10cm}{1cm}{xfigures/xfigtest}
\item convert the \texttt{xfig}- into \texttt{eps}-images:  
  \texttt{/u/hasan/bin/fig2eps <filename>}
\item if you want to change the image paths, you should redefine the
  \texttt{\textbackslash graphicspath} option and the paths using:
  \begin{footnotesize}
    \begin{itemize}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash imagedir\}\{./images/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash imagetemplatedir\}\{/u/figures/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash xfigdir\}\{./xfigures/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash logodir\}\{./logos/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash audiodir\}\{./audio/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash videodir\}\{./video/\}}
    \item[] \texttt{\textbackslash renewcommand*\{\textbackslash sourcedir\}\{./sources/\}}
    \end{itemize}
  \end{footnotesize}
\item if you want to \alert{extend} the path you should use
  \texttt{\textbackslash extendgraphicspath}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{Including Images}
\vfill
\begin{minipage}[b]{.5\linewidth}
  \begin{center}
    \includegraphics[width=.5\linewidth]{car}\\
    car label
  \end{center}
\end{minipage}
\begin{minipage}[b]{.5\linewidth}
  \begin{itemize}
  \item my
  \item image
  \item description
  \item using \texttt{minipage} environment
  \item my
  \item image
  \item description
  \end{itemize}
\end{minipage}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{Including Images}
\vfill
\begin{table}
  \centering
  \begin{tabular}{@{} M{.3\linewidth} M{.7\linewidth} @{}}
    \includegraphics[width=.9\linewidth]{motorbike}
    & 
    \begin{itemize}
    \item my
    \item image
    \item description
    \item using \texttt{tabular} environment
    \item image
    \item description
    \end{itemize}
  \end{tabular}
\end{table}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{Including Images}
\begin{hvcenter}
  \includegraphics[width=.5\linewidth]{airplane}\\
  my image using \texttt{hvcenter} environment
\end{hvcenter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{Recommendation}
\vfill
\begin{itemize}
\item \red{You should always specify the full image path for images from \\
    \texttt{/u/figures/<user>/<name>\_<User>\_<DDMmmYY>}} \\
  \eg \texttt{/u/figures/dreuw/TangentDistance\_Dreuw\_18Jul06}
\item the package \texttt{epsf} and \texttt{psfig} must be replaced by the newer
  \texttt{graphicx.sty}
  %\item \texttt{epsf} is only a wrapper for \texttt{graphicx} for old documents
  % created using \texttt{psfig}
\item you can use \texttt{\textbackslash usepackage\{afterpage\}} to flush all
  images first before the next text part begins, by calling
  \texttt{\textbackslash afterpage\{\textbackslash clearpage\}}
\item do \red{not use the following deprecated commands}, as they will cause
  problems with pdfTeX (pdfTeX can also run in DVI-mode).
\begin{verbatim}
\newif\ifpdf
\ifx\pdfoutput\undefined
    \pdffalse       %NO Pdflatex
\else
    \pdfoutput=1    %PDFLatex
    \pdftrue
\fi 
\end{verbatim}
\item \alert{\texttt{\textbackslash usepackage\{ifpdf\}}} is already part of the
  style and runs also under \hyphenation{MacOSX} MacOSX and is used instead. It
  provides the \texttt{\textbackslash ifpdf} command.
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Default Search Paths}
\vfill
\begin{itemize}
\item \texttt{<filename>} means the path to the file (with or without parent
  folder depending on the \texttt{\textbackslash graphicspath} option
  \alert{without the file extension} (e.g. *.pdf\_t, *.jpg, or *.eps).
\item you can store your \texttt{xfig} figures in the \texttt{./xfigures/}
  folder as this path is already declared in the \texttt{\textbackslash
    graphicspath} option but \texttt{\textbackslash u\textbackslash figures} is
  preferred.
\item you can define \texttt{\textbackslash renewcommand*\{\textbackslash
    defaultaudiodir\}\{\textbackslash audiodir\}} to specify the default search
  directory for audio files. [Default is \defaultaudiodir]
\item you can define \texttt{\textbackslash renewcommand*\{\textbackslash
    defaultvideodir\}\{\textbackslash videodir\}} to specify the default search
  directory for video files. [Default is \defaultvideodir]
\item you \alert{should} set \texttt{\textbackslash
    renewcommand*\{\textbackslash defaultaudiodir\}\{\}} \\ or
  \texttt{\textbackslash
    renewcommand*\{\textbackslash defaultvideodir\}\{\}} to clear the default search path for videos in order to use absolute filenames \eg \\
  \texttt{\textbackslash
    videofilelogobox\{/u/wherever/you/want/i6gesture.avi\}}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Running External Applications}
\begin{itemize}
\item You can call every external application with \texttt{\textbackslash
    href\{run:<script>\}} command.  For this you have to specify a shell
  script, in which you can execute any system commands. This will work with Xpdf
  \href{http://www.foolabs.com/xpdf/}{\includegraphics[scale=0.5]{logos/xpdf}}
  \alert{and} AdobeAcrobat
  \href{http://www.adobe.com/products/acrobat/readstep2.html}{\includegraphics[scale=0.5]{logos/acrobat}}.
\item \href{run:scripts/demo.sh}{Shell demo starting xclock} will open a new
  terminal which will start the \texttt{xclock} on your computer.
\item Also you can use this command to play movies. Clicking on the image will
  play a movie
  \href{run:\videodir/lti.mpg}{\includegraphics[scale=0.15]{\videodir/lti}}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Hyperlinks}
\begin{itemize}
\item You can link to an url with the \texttt{\textbackslash url\{<link>\}}
  command, e.g.\ \url{www.google.com}. If you have a long url name or you want
  to change the name of the url anchor then use \texttt{\textbackslash
    href\{<link>\}\{<name>\}}, e.g.\ \href{http://www.google.com}{Google}.
\item use the \texttt{\textbackslash urllogo\{<link>\}} or
  \texttt{\textbackslash urllogobox\{<link>\}} to create links with an
  additional icon before like \urllogo{www.google.com} or
  \urllogobox{www.google.com}
\item With the \texttt{\textbackslash hyperlink\{<anchor>\}\{<linkname>\}} and \\
  \texttt{\textbackslash hypertarget\{<anchor>\}\{<targetname>\}} commands you
  can create links inside your slides.  Example: this can be usefull to
  \hyperlink{anchorname}{link to slides from the appendix} (e.g.\ the data used
  to create a plot).
\item Use the \texttt{\textbackslash autoref\{<anchor>\}} command to
  automatically refer to \autoref{fig:images}, \autoref{eq:histogram_formula},
  or \autoref{tab:table}.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Playing Audio Files}
\hypertarget{sli:hyperlinks_audio}
\vfill
Audio examples from the Verbmobil corpus:
\begin{footnotesize}
  \begin{itemize}
  \item \texttt{\textbackslash audiofilelogobox\{<file>\}}:
    \audiofilelogobox{example-verbmobil-1.wav}
  \item \texttt{\textbackslash audiofilelogo\{<file>\}}:
    \audiofilelogo{example-verbmobil-3.wav}
  \item \texttt{\textbackslash audiofilebox\{<file>\}}:
    \audiofilebox{example-verbmobil-2.wav}
  \item \texttt{\textbackslash audiofile\{<file>\}}:
    \audiofile{example-verbmobil-1.wav}
  \item \texttt{\textbackslash audiofiletext\{<file>\}\{<text>\}}:
    \audiofiletext{example-verbmobil-2.wav}{Example 2 from Verbmobil2}
  \end{itemize}
\end{footnotesize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Playing Video Files}
\hypertarget{sli:hyperlinks_video}
\vfill
%
%\renewcommand*{\defaultvideodir}{} 
% -> if set, you could write e.g. \videofilelogobox{/u/wherever/you/want/i6gesture.avi} 
%    or \videofilelogobox{./video/i6gesture.avi} instead of only \videofilelogobox{i6gesture.avi}
%
\renewcommand{\defaultvideodir}{}
Video examples from the LTI--Gesture and i6--Gesture database:
\begin{footnotesize}
  \begin{itemize}
  \item \texttt{\textbackslash
      videofilethumbnailbox\{<file>\}\{<param>\}\{<image>\}}:
    \videofilethumbnailbox{\videodir/lti.mpg}{scale=0.1}{\videodir/lti}
  \item \texttt{\textbackslash
      videofilethumbnail\{<file>\}\{<param>\}\{<image>\}}:
    \videofilethumbnail{\videodir/lti.mpg}{scale=0.1}{\videodir/lti} %use scale=1.0 to disable scaling
  \item \texttt{\textbackslash videofilelogobox\{<file>\}}:
    \videofilelogobox{\videodir/i6gesture.avi}
  \item \texttt{\textbackslash videofilelogo\{<file>\}}:
    \videofilelogo{\videodir/i6gesture.avi}
  \item \texttt{\textbackslash videofilebox\{<file>\}}: \videofilebox{\videodir/i6gesture.avi}
  \item \texttt{\textbackslash videofile\{<file>\}}: \videofile{\videodir/i6gesture.avi}
  \item \texttt{\textbackslash videologo\{<file>\}}: \videologo{\videodir/i6gesture.avi}
  \item \texttt{\textbackslash videofiletext\{<file>\}\{<text>\}}:
    \videofiletext{./video/i6gesture.avi}{Example from i6--Gesture database} %here with absolute/relative path!
  \end{itemize}
\end{footnotesize}
\vfill 
There are also icons for CDs \cdicon and DVDs \dvdicon which you can use to play
videos from an external device. Just write a script to mount the device and a
player command which will start playing a file from the mounted device.
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Page Numbering}
\hypertarget{sli:pagenumbering}
\vfill
\begin{itemize}
\item you can use the package option \texttt{lastpage} or \texttt{userlastpage}
  to enable a page numbering like ``n of m'', otherwise the pages will have a
  single page number.
\item if you use the package option \texttt{userlastpage} \alert{you have to} call
  \texttt{\textbackslash LastPage} \\ or \texttt{\textbackslash FinalPage} at the
  end of your last slide to enable a correct numbering of the slides.
\item \texttt{\textbackslash FinalPage} will automatically generate a ``Thank
  you for your attention page'' with your name, email and www address.
\item you can change the layout of \emph{your} last page by using
  \texttt{\textbackslash LastPage} at the end of your last slide. This will
  simply insert a blank page and enable a correct numbering of the
  slides.
\item you can also use \texttt{\textbackslash LastPage} or
  \texttt{\textbackslash FinalPage} without specifying the package option
  \texttt{lastpage}. This won't affect the page numbering.
\item to disable the page numbering you must use the the package option
  \texttt{nonumber}.
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Changing Logos}
\hypertarget{sli:logos}
\vfill
\renewcommand*{\topleftlogo}{\raisebox{-7mm}[0mm][0mm]{\includegraphics[height=10mm, width=2cm]{\logodir TT2logo}}}
\begin{itemize}
\item you can display a third logo in the topleft corner on each slide by redefining 
  \texttt{\textbackslash topleftlogo}
  \begin{itemize}
  \item Example 1: \newline \footnotesize{\texttt{\textbackslash renewcommand*\{\textbackslash topleftlogo\}\{\textbackslash includegraphics[height=6mm]\{\textbackslash logodir YOUR--THIRD-LOGO\}\} }} \normalsize
  \item Example 2: \newline \footnotesize{\texttt{\textbackslash renewcommand*\{\textbackslash topleftlogo\}\{\textbackslash raisebox\{-7mm\}[0mm][0mm]\{\%\\
          \textbackslash includegraphics[height=6mm]\{/u/path/to/your/third/LOGO\}\}\}}} \normalsize
  \end{itemize}
\item Also you could redefine the other logos \texttt{\textbackslash
    toprightlogo} \newline and \texttt{\textbackslash bottomrightlogo} in this way
\item if you are still not happy you might change \texttt{\textbackslash lhead},
  \texttt{\textbackslash rhead} or \texttt{\textbackslash rfoot} at your own
  ``risk''
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Page Titles}
\renewcommand*{\topleftlogo}{} %% remove the third logo defined at the previous slide
\hypertarget{sli:titles} 
\vfill 
You can break pages if you choose the option \texttt{allowpagebreaks}, the title
will be repeated on the next slide ...  
\vfill 
\NewPage
\vfill 
... as you can see (or not, depending on the option) !  
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Overlay Slides}
\vfill
\begin{itemize}
\item you can use the \texttt{\textbackslash NewOverlay} command to create
  overlay slides, i.e.\ to correctly number the pages
\item you will get pdfTeX warnings about duplicate identifiers with
  \texttt{pdflatex}. Use \texttt{make slides.pdf} instead.
  \begin{itemize}
  \item \textcolor{gray}{not yet ...}  % or use white instead of gray
  \end{itemize}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewOverlay\centerline{Overlay Slides}
\vfill
\begin{itemize}
\item you can use the \texttt{\textbackslash NewOverlay} command to create
  overlay slides, i.e.\ to correctly number the pages
\item you will get pdfTeX warnings about duplicate identifiers with
  \texttt{pdflatex}. Use \texttt{make slides.pdf} instead.
  \begin{itemize}
  \item not yet ... but now, and look at the page number
  \end{itemize}
\end{itemize}
\vfill

%  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \NewCenteredPage\headline{Centered Slides}
%  \begin{itemize}
%  \item you can use the \texttt{\textbackslash NewOverlay} command to create vertical centered slides
%  \item this may cause problems when you change between normal and automatically
%    centered slides!
%  \end{itemize}
%  \vfill % because we use only one \NewCenteredPage command here ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Predefined Custom Commands}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}
Changing description layout:
\begin{description}
  \item[FirstDescription] is usually only black and bold.
  \item[SecondDescription] is usually only black and bold.
  \item[ThirdDescription] is usually only black and bold.
\end{description}
\vfill
\end{block}

\begin{block}
  vertical centered Block Environment
\end{block}

\begin{hvcenter}
  vertical and horizontal centered Block Environment
\end{hvcenter}

\begin{block}
  \begin{itemize}
  \item Most commands make usage of the \texttt{\textbackslash xspace} command, which
    allows a context sensitive whitespace placement after macros.
    \begin{itemize}
    \item some math symbols with ensured math mode: \nat, \rel, \congmod
    \item some operators: $\argmin_{x}$, $\argmax_{y}$, $\invers{10}$
    \item arrows: \ra
    \item \eg something, \Eg, \ie nothing, \Ie, \cf page, \Cf, \etc, something
      \vs nothing, \wrt $t$, \dof, Author \etal, \zB auf Deutsch, \ZB
    \end{itemize}
  \end{itemize}
\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section{\LaTeX \ Tips} 
\secttoc
\vfill
Some general \LaTeX \ recommendations and tips:
\begin{itemize}
\item replace: \texttt{\$\$...\$\$} by
  \texttt{\textbackslash[...\textbackslash]}
\item replace: \texttt{\textbackslash centerline\{...\}} by
  \texttt{\{\textbackslash centering ...\}} or the \texttt{center}
  environment
\item replace: \texttt{eqnarray} environment where possible by \texttt{align}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Formulas} 
\hypertarget{sli:formulas} 
\vfill 
Write formulas with the \texttt{\textbackslash begin\{equation\}} environment or
with the double \texttt{\$\$} signs, use a single \texttt{\$} sign if you want
to write a formula on the same text line.  
\vfill
\begin{itemize}
\item numbered equation 
  \begin{equation}
    h_c(X) = \frac{1}{L_X} \sum_{l=1}^{L_X} \delta(c,c(x_l)) \label{eq:histogram_formula}
  \end{equation}
\item unnumbered equation 
  \begin{equation}
    h_c(X) = \frac{1}{L_X} \sum_{l=1}^{L_X} \delta(c,c(x_l)) \nonumber
  \end{equation}
  or use 
  \[ h_c(X) = \frac{1}{L_X} \sum_{l=1}^{L_X} \delta(c,c(x_l)) \] %%  i.e. do not use '$$...$$' which is deprecated
\item with $ h_c(X) = \frac{1}{L_X} \sum_{l=1}^{L_X}
  \delta(c,c(x_l))$ on the same line.  
\end{itemize}
%  with 
%    \begin{center}
%      \begin{tabular}{ll}
%        $L_X$ & number of patches for image $X$\\
%        $x_l$ & $l$-th patch for image $X$\\
%        $c(x_l)$ & closest cluster identifier for $x_l$\\
%        $h_c(X)$ & $c$-th bin of histogram for image $X$\\
%      \end{tabular}
%    \end{center}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Tables} 
\hypertarget{sli:tables} 
\vfill 
You can use the \texttt{\textbackslash begin\{table\}} environment to present
your results. Also you can link the whole table to another slide with the
\texttt{\textbackslash hyperlink} command in combination with
\texttt{\textbackslash textcolor\{black\}}, otherwise the table would appear in
link color.
\vfill
Some results achieved at the \isechsicon.
\begin{table}
  \centering
  \caption{Table caption above. \alert{Click on the table} to jump to the \appendixname}
  \hyperlink{table_data}{\textcolor{black}{%
      \begin{tabular}{|l|l|l|l|l|}
        \hline
                              &  motorbikes   &   bicycles &   people   & cars     \\
        \hline
        Task 1                &  6,8 (17)     &  5,8 (15)  & 5,6 (15)   & 4,5 (17) \\
        Task 2&  2,3 (11)     &  2,3 (9)      & 2,3 (9)    & 2,3 (10)              \\
        \hline
      \end{tabular}
    }
  }
  \label{tab:table}
\end{table}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{Tables} 
\begin{table}      
  %\footnotesize
  \centering
  \caption{Error rates [\%] using the nicer \emph{booktabs style} ordered by decimal position}
  \begin{tabular}{@{} l D{2.4} D{10.10} D{2.2} @{}}
    \toprule  
    \multicolumn{1}{@{}N}{Spatial derivative} & \multicolumn{1}{N}{Original} & \multicolumn{1}{N}{1st time derivative} & \multicolumn{1}{N@{}}{2nd time derivative} \\ 
    \multicolumn{1}{@{}N}{(Sobel)}  &                            &                            &                                                     \\ 
    \midrule

    no                              &  0.0001                    &  0.0001                      & 15.7                                                \\
    horizontal                      &  0.001                     &  0.001                       & 20.0                                                \\
    vertical                        &  1.01                      &  1.01                        & 16.4                                                \\ 
    magnitude                       & 10                         & 10                           &  7.1                                                \\ 
    squared magnitude               & 11.1                       & 11.1                         & 34.2                                                \\ 
    \bottomrule
  \end{tabular}
  \label{tab:table_book}
\end{table}
\vfill
\begin{table}      
  %\footnotesize
  \centering
  \caption{Table using cmidrule command}
  \begin{tabular}{@{} l l D{2.2} D{2.2} @{}}
    \toprule 
    \multicolumn{1}{@{}N}{Densities}  & \multicolumn{1}{N}{Pooling}  & \multicolumn{1}{N}{Gaussian ER[\%]} & \multicolumn{1}{N@{}}{Laplacian ER[\%]} \\ 
    \cmidrule(r){1-1}                   \cmidrule(lr){2-2}              \cmidrule(lr){3-3}                   \cmidrule(l){4-4} 
    Single                            & No                           & 29.2                                & 30.7                                 \\
                                      & Yes                          & 29.2                                & 30.7                                 \\ 
    \addlinespace                                                                                                                  
    Mixture                           & No                           & 21.4                                & 29.2                                 \\ 
                                      & Yes                          & 23.5                                & 27.8                                 \\ 
    \bottomrule
  \end{tabular}
  \label{tab:basic-hmm-difference-0-1}
\end{table}
\vfill
\begin{table}      
  \centering
  \caption{Table using cmidrule command and rotated column heads}
  \begin{tabular}{@{} l M{7cm} V{3cm} D{2.2} c @{}}
    \toprule 
    Type & \multicolumn{2}{c}{Features} & \multicolumn{1}{N}{Error Rate} & \multicolumn{1}{V{5cm}@{}}{Info}\\
    \cmidrule(lr){2-3}
    &
    \multicolumn{1}{R{6em}}{Long Feature description 1} 
    &
    \multicolumn{1}{R{6em}}{Long Feature description 2} 
    &
    &
    \\
    \cmidrule(l){1-1}\cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}\cmidrule(r){5-5}
    nice 
    & 
    feature very long and centered baseline with fixed width feature very long and centered baseline 
    with fixed width feature very long and centered baseline  with fixed width 
    & description very long and NOT centered with fixed width 
    & 1.23\% 
    & horizontal centered w/o width\\
    \bottomrule
  \end{tabular}
  \label{tab:cmidrule-rotated-vcentered}
\end{table}    
\vfill
\begin{table}      
  \centering
  \caption{Table using *-command}
  \begin{tabular}{@{} *{4}{l} @{}}
    \toprule 
    \multicolumn{1}{@{}n}{Nominativ} & \multicolumn{1}{n}{Genetiv}    & \multicolumn{1}{n}{Dativ}     & \multicolumn{1}{@{}n}{Akkusativ} \\
    \midrule
    die Frau  & der Frau   & der Frau  & die Frau  \\
    der Mann  & des Mannes & dem Manne & den Mann  \\
    das Kind  & des Kindes & dem Kinde & das Kind  \\
    \bottomrule
  \end{tabular}
\end{table}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Citing} 
\hypertarget{sli:citing} 
\vfill 
Use the \texttt{\textbackslash cite\{<anchor>\}}
command to refer to an entry in your bibliography. You can click on the citation
to jump to your bibliography.

You can use the backreferences in the bibliography to jump back to your slide.
\vfill
\begin{enumerate}[{Example*} a)]
\item Results on Caltech database:
  \cite{deselaers05:_discr_train_objec_recog_image_patch}\\
  very good on motorbikes and airplanes, quite good on faces
  \vfill
\item Results on medical
  radiographs:\cite{loupias:_wavel_salien_point_image_retriev}\\
  quite good, specialized approaches are better
  \vfill
\end{enumerate}
\vfill
Use the package \texttt{\textbackslash usepackage\{cite\}} to sort your slides
\vfill 
{\footnotesize *The enumerate environment from the \texttt{paralist} package can use
  special labels.}  
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{\textcolor{red}{C}\textcolor{green}{o}\textcolor{blue}{l}\textcolor{red}{o}\textcolor{green}{r}\textcolor{blue}{s}}
\hypertarget{sli:colors}
\vfill
\begin{itemize}
  \setlength\itemsep{2ex}   %%% redefines itemsep only for this section
\item How to \alert{highlight} words? This can be done by the
  \texttt{\textbackslash alert\{<text>\}} command. You should use this command
  for \alert{important} words.
\item You can use \textcolor{red}{additional} \textcolor{green}{colors} with the
  \texttt{\textbackslash textcolor\{<color>\}} \magenta{command} in your
  \blue{slides} to \red{highlight} \green{words}, but don't use \magenta{too}
  \isechsbluedark{much} \isechsblue{colors}! \\ The \texttt{\textbackslash
    alert\{<text>\}} command should always be \alert{preferred}
\end{itemize}
\vfill
Spoken:
\begin{quote}
  also ich vielleicht ist grade zu der Zeit die CeBit das w\"are
  vielleicht f\"ur uns fachlich auch ganz interessant
\end{quote}
Recognized:
\begin{quote}
  also ich vielleicht \textcolor{red}{das} grade zu der Zeit die
  CeBit das w\"are vielleicht \textcolor{red}{---} uns fachlich auch
  \textcolor{green}{noch} ganz interessant
\end{quote}
\textcolor{red}{substitution}
\qquad
\textcolor{green}{insertion}
\qquad
\textcolor{red}{---} deletion
\vfill
\begin{equation*}
  \textrm{WER} =
  \frac{1 \textrm{ deletion} + 1\textrm{ insertion} + 1\textrm{ substitution}}
  {19 \textrm{ spoken words}} = 15.8 \%
\end{equation*}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section{\LaTeX \ Tricks}
\secttoc
\vfill
Some \LaTeX\xspace tricks
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Hide Table Columns}
\def\ignore#1\endignore{} 
\newcolumntype{I}{@{}>{\ignore}l<{\endignore}} 
\hypertarget{sli:tricks}
\vfill
Hide complete columns of your tables without changing the table values, e.g. hide a result of a WER\% table
\vfill
w/o hiding: 
\begin{tabular}{|l|l|l|} 
  \hline
  foo & bar & baz  \\ 
  blu & bli & blo  \\
  \hline
\end{tabular} 
\vfill 
w/ hiding:
\begin{tabular}{|l|Il|} % also change the the vertical lines
  \hline
  foo & bar & baz \\ 
  blu & bli & blo \\
  \hline
\end{tabular} 
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Includegraphics with Clip \& Crop}
\vfill
If you want to crop something from an image or a plot (e.g. the title of a plot)
\begin{figure}
  \centering
  \includegraphics{car}
  \quad
  \includegraphics[clip=true,viewport=4.5cm 3.0cm 9cm 6cm]{car}%
\end{figure}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Phantom vs. Itemize}
\vfill
Phantom: \\
Das ist die erste Zeile\\
\phantom{Das ist die erste Zeile} das ist die zweite zeile\\
\phantom{Das ist die erste Zeile das ist die zweite zeile} und die dritte
\vfill
Itemize:\\
\begin{itemize}
  \setlength{\itemsep}{3ex}
  \setlength{\leftmargini}{5em}   
  \setlength{\leftmarginii}{10em}  
  \setlength{\leftmarginiii}{15em}
\item Das ist die erste Zeile
  \begin{itemize}
  \item das ist die zweite zeile
    \begin{itemize}
    \item und die dritte
    \end{itemize}
  \end{itemize}
\end{itemize}
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Embedded Fonts}
\vfill
converts \texttt{foo.pdf}, a file \textcolor{red}{w/o} embedded fonts, into \texttt{foo2.pdf}, a file \textcolor{green}{w/} embedded fonts (e.g.\ for IEEE PDF eXpress):
\begin{itemize}
\item \texttt{gs -sDEVICE=pdfwrite -dPDFSETTINGS=/prepress -dNOPAUSE -q \\ -dBATCH -sOutputFile=foo2.pdf foo.pdf}
\end{itemize}
\vfill 
\alert{There is a bug in in ESP Ghostscript 8.15.x} that may produce a
``drawing error'' in Acroread7.0 but not in xpdf. The newest subversion revision
from January 2006 or e.g\ AFPL Ghostscript 8.53 fixes this problem.

This can cause also ``drawing error'' problems in Acroread7.0 when you use
\texttt{convert} to convert e.g.\ a JPEG/PNG image into an EPS image.
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{\texttt{write18} Hacks}
\vfill
\begin{verbatim}
\makeatletter
\begingroup
\catcode`\%12\relax
\let\\\relax
\edef\doshell{
date '+%d.%m.%Y %r' |
      awk '{print "\\newcommand*{\\datum}{" $1 "\\xspace}";
            print "\\newcommand*{\\zeit}{" $2 "\\xspace}"
          }' > date.tex
}
\immediate\write18{\doshell}
\endgroup
\makeatother
\input{date}
\end{verbatim}
%%%
\ifpdf
\makeatletter
\begingroup
\catcode`\%12\relax
\let\\\relax
\edef\doshell{
date '+%d.%m.%Y %r' |
      awk '{print "\\newcommand*{\\datum}{" $1 "\\xspace}";
            print "\\newcommand*{\\zeit}{" $2 "\\xspace}"
          }' > date.tex
}
\immediate\write18{\doshell}
\endgroup
\makeatother
\input{date}
This slide was created the \datum at \zeit.
\else
This example needs to be compiled using \\ \alert{\texttt{pdflatex -shell-escape slides.tex}}
\fi
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\centerline{\texttt{write18} Hacks}
\vfill
\begin{verbatim}
\immediate\write18{% 
 echo $USER $HOME $TMPDIR > variables.tex
}
\end{verbatim}
The user, his home-directory und his scratch-directory:
\begin{verbatim}
\input{variables.tex}
\end{verbatim}
\ifpdf
\immediate\write18{% 
 echo $USER $HOME $TMPDIR > variables.tex
}
\alert{\input{variables.tex}}
\else
\vfill
This example needs to be compiled using \\ \alert{\texttt{pdflatex -shell-escape slides.tex}}
\fi
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Fancy verbatim}
\vfill
\begin{Verbatim}[commandchars=\\\{\}] 
  A \textcolor{red}{red word}  within a verbatim environment. 
\end{Verbatim} 
\vfill
program code can be colored by using the listings package, see dante-faq 7.3.4
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\subsection{Embedding graphics within a \LaTeX-document}
\begin{block}
  TODO ...  
\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section{Converting And Printing The Slides}
\secttoc
\begin{itemize}
\item XEmacs editor:
  \begin{itemize}
  \item change into PDF-mode with \texttt{C-c C-t C-p} if you want to create PDF
    slides, otherwise PS slides will be created
  \item run \LaTeX with \texttt{C-c C-c}
  \item run again to open the standard viewer \texttt{xpdf} or \texttt{xdvi}
    depending on the mode
  \end{itemize}
\item Creating PDF or PS slides on the command line:
  \begin{itemize}
  \item type \texttt{pdflatex slides} and \texttt{xpdf slides.pdf} to view the
    result
  \item type \texttt{latex slides}, \texttt{dvips slides}, and \texttt{gv
      slides.ps} to view the result
  \end{itemize}
\item Converting:
  \begin{itemize}
  \item use \texttt{dvipdf slides} to convert the created PS-dvi files into
    PDF-slides
  \item use \texttt{dvips slides} to convert the created PS-dvi files into
    PS-slides
  \end{itemize}
\item Handout Printing:
  \begin{itemize}
  \item PDF-slides
    \begin{itemize}
    \item use the Acrobat Reader to print the slides with printer
      option\\ \texttt{/u/hasan/bin/pp -4sup}
    \item or convert the slides with \texttt{pdfnup --nup 2x2
        slides.pdf} and \\print the generated output
      \texttt{slides-2x2.pdf}
    \end{itemize}
  \item PS-slides: use gv with printer option \texttt{/usr/bin/lpz
      -4slidessea} or try \texttt{/u/hasan/bin/pp -4sup}
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\FinalPage
%%%
\vfill 
\centerline{\footnotesize PS: This page was generated automatically by calling
\texttt{\textbackslash FinalPage}.} 
\vfill
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage
%\headline{\refname}
%\renewcommand*{\refname}{}
\bibliographystyle{i6bibliostyle}
\bibliography{slides}

%\clearpage
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section*{\appendixname: First Slide}
\vfill
\hypertarget{anchorname}{Hyper Target} on the first appendix slide. 
Look at the current page number.
\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NewPage\section*{\appendixname: Table Data}
\vfill
\hypertarget{table_data}{Table Data} on the second appendix slide.
Look at the current page number.
\vfill

\end{document}

### 2016-09-19 paper/paper.tex
paper template

\documentclass[twoside,11pt,a4paper]{article}

% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx,curves,float,rotating}

\usepackage{amsmath, amssymb, latexsym}  % math stuff
\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
\usepackage[ngerman]{babel}                     % otherwise use british or american
\usepackage{theorem}                            % instead of \usepackage{amsthm}
\usepackage{dcolumn}
\usepackage{hyperref}

% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{xspace}                             % context sensitive space after macros
\makeatletter 
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
\def\etc{{etc}\onedot} \def\vs{{vs}\onedot} 
\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
\def\etal{{et al}\onedot}
\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Allgemeines ueber LaTeX
%% =======================
%%
%% 
%% LaTeX (sprich: 'latech') ist ein Textsatzsystem im Gegensatz zu
%% einem Textverarbeitungsprogramm wie beispielsweise Word. Wie im
%% professionellen Buchdruck werden Texte nicht nur geschrieben,
%% sondern muessen auch 'gesetzt' werden. Da LaTeX wie eine Program-
%% miersprache gehandhabt wird, ermoeglicht es dem Benutzer auch
%% eine groesstmoegliche Freiheit in der Gestaltung einer Seite.
%% 
%%
%%
%% Eine der Staerken von LaTeX bildet die Vielzahl von Befehlen
%% zur Beschreibung mathematischer Formeln. Wie das im einzelnen
%% geht wird spaeter unten beschrieben.
%%
%%
%%
%% Die Vorteile von LaTeX haben natuerlich ihren Preis. Anders als
%% bei Word kann man das Resultat nicht direkt zur Eingabezeit
%% begutachten. Statt dessen muss das TeX-File (das ist die Datei,
%% die Ihr gerade lest und immer auf den Namen <Dateiname>.tex enden
%% muss) zuerst mit dem Befehl
%%
%%
%%                  latex  <Dateiname>.tex
%%         
%%
%% compiliert werden. Dabei wird ein file erzeugt, das den Namen
%% 
%%
%%    --->          <Dateiname>.dvi          
%%
%%
%% traegt. Die Endung dvi steht fuer 'device independent'und enthaelt
%% bereits die geraeteunabhaengige Darstellung, die man sich mit dem
%% Aufruf
%%
%%
%%    --->          xdvi  <Dateiname>.dvi
%%
%%
%% anschauen kann. Um die *.dvi Datei nach PostScript zu konvertieren,
%% verwendet man den Aufruf
%%
%%
%%    --->          dvips <Dateiname>.dvi -o <Name>.ps
%%
%%
%% Diese Datei kann von jedem PostScript faehigen Drucker ausgedruckt
%% werden.

%% Aufbau eines LaTeX Dokuments
%% ============================
%%
%% Jedes LaTeX Dokument beginnt mit der Angabe der Dokumentklasse in
%% geschweiften Klammern (ganz oben in der ersten Zeile). Dies ist
%% hier die article Klasse. In die eckigen Klammern und den
%% nachfolgenden usepackages koennen weitere Pakete eingebunden
%% werden, die neben dem reinen LaTeX Standard zusaetzliche Befehle
%% zur Verfueugung stellen.
%%
%% Die folgenden Befehle fuehren neue Umgebungen ein, stellen
%% Macros zur Verfuegung und modifizieren das Seitenlayout.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%	Macros fuer neue Umgebungen
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
\newlength{\textwd}
\newlength{\oddsidemargintmp}
\newlength{\evensidemargintmp}
\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
\newlength{\textht}
\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}

\newenvironment{deflist}[1][\quad]%
{  \begin{list}{}{%
      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
      \settowidth{\labelwidth}{\textbf{#1}}%
      \setlength{\leftmargin}{\labelwidth}
      \addtolength{\leftmargin}{\labelsep}}}
{  \end{list}}

\newenvironment{Quote}% Definition of Quote
{  \begin{list}{}{%
      \setlength{\rightmargin}{0pt}}
      \item[]\ignorespaces}
{\unskip\end{list}}

\theoremstyle{break}
\theorembodyfont{\itshape}	
\theoremheaderfont{\scshape}

\newtheorem{Cor}{Corollary}
\newtheorem{Def}{Definition}
%\newtheorem{Def}[Cor]{Definition}

\newcolumntype{.}{D{.}{.}{-1}}

\pagestyle{headings}
\textwidth 15cm
\textheight 23cm
\oddsidemargin 1cm
\evensidemargin 0cm
%\parindent 0mm

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%       Jetzt geht's los
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%% Mit dem Befehl \begin{document} weist man LaTeX an, dass das Dokument
%% beginnt. Ganz am Ende des Dokuments gibt es einen entsprechenden Befehl
%% \end{document}
%%
%% Aber beginnen wir zunaechst mit der Titelseite unseres Seminars. Eine
%% Titelseite soll natuerlich die folgenden Dinge enthalten:
%%
%%   1. Titel des Seminars
%%   2. Bezeichnung des Lehrstuhls, an dem das Seminar angefertigt wurde
%%   3. Name des Seminaristen
%%   4. Name des Betreuers
%%
%% Wie realisiert man das in LaTeX?
%%
%% Zunaechst einmal soll eine Titelseite natuerlich keine Numerierung
%% erhalten. LaTeX kann selbstaendig Zeilen und Seiten umbrechen und
%% fuehrt auch die Seitenzaehlung selbstaendig durch. Da wir die Zaehlung
%% fuer die Titelseite unterruecken wollen, geben wir mit dem Befehl
%% \pagestyle{empty} vor, dass LaTeX weder Seitenzahl noch Kopf- oder
%% Fusszeilen einfuegt. Da wir den Titel gerne zentriert haetten, benutzen
%% wir die Befehle \begin{center}  ...  \end{center} . Alle Eingaben, die
%% wir innerhalb dieser 'center' Umgebung eingeben, werden dann automatisch
%% von LaTeX zentriert.
%%
%% Der Befehl \LARGE beschreibt die Font-Groesse. Fuer die Groesse des
%% Schriftsatzes gibt es mehrere Befehle. Diese sind in aufsteigender
%% Reihenfolge:
%%
%%	\tiny 		\normalsize	\huge
%%	\scriptsize   	\large		\Huge
%%	\footnotesize	\Large
%%	\small		\LARGE
%%
%%
%% Um den Font zu aendern, stehen folgende Befehle zur Verfuegung:
%%
%%	\textrm{}	roman font
%%	\textsf{}	sans  serif
%%	\texttt{}	typewriter
%%	\textbf{}	bold series
%%	\textit{}	italic		(kursiv)
%%	\textsc{}	SMALL CAPS
%%	\textnormal{}	typeset text in the normal font
%%
%%
%% Mit dem Kommando \\  gibt man EXPLIZIT(!) an, dass man eine Zeile
%% umbrechen will (wie gesagt, normalerweise macht LaTeX das selbst).
%% Optional kann man hier \\ einen Wert angeben wie z.B. \\[4ex]
%% Damit legt man fest, wieviel Zwischenraum bis zur naechsten
%% Zeile bestehen soll. Das 'ex' steht fuer die Groesse des Symbols x.
%% Hier kann man auch die Einheiten 'pt' fuer Punkte, cm, ... benutzen.
%%
%% Der \vfill Befehl fuegt soviel Leerraum ein, das der restliche Text
%% bis an den unteren Rand gedrueckt wird (die Leerzeile davor wird
%% benoetigt). Der \newpage Befehl erzeugt einen EXPLIZITEN(!) Seiten-
%% umbruch. Da wir eine Leerseite haben wolllen wird mit '\ ' ein
%% Space-Symbol auf die Seite eingefuegt, so dass wir wieder einen
%% \newpage Befehl einsetzen koennen.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%               Title
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{empty}

\begin{center}

    Rheinisch-Westf\"alische Technische Hochschule Aachen \\
    Lehrstuhl f\"ur Informatik 6 \\
    Prof. Dr.-Ing. Hermann Ney\\[6ex]
    Seminar Titel im SS 20XX\\[12ex]                          % auch Seminar Titel und Datum ändern!!!

    \LARGE
    \textbf{Titel des Vortrages} \\[6ex]
    \textit{Vorname Name} \\[6ex]
    \Large
    Matrikelnummer 123 456 \\[6ex]
    Datum des Vortrages

    \vfill
    \Large Betreuer: Vorname Name

\end{center}

\newpage
\ 
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%               Inhaltsverzeichnis / Tabellenverzeichnis / Abbildungsverz.
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{headings}
\tableofcontents
\listoftables
\listoffigures
\newpage
\pagestyle{empty}
\ 
\newpage
\pagestyle{headings}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%               Abschnitt 1
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Name des ersten Abschnitts}
\label{sec:Abschnitt-1}

%% Der Befehl \section{} erstellt den ersten Abschnitt. Unter-
%% abschnitte werden mit \subsection{} generiert. Mit dem \label
%% Befehl erzeugt man eine symbolische Marke auf den Abschnitt.
%% Will man spaeter im Text auf den ersten Abschnitt referenzieren,
%% so gibt man das wie folgt an:

Dies ist eine Referenz auf Abschnitt~\ref{sec:Abschnitt-1}

%% LaTeX uebernimmt damit automatisch die Nummerierung der Referenzen.
%% Der Vorteil liegt darin, dass wenn man irgendwann mitten im Text einen
%% neuen Abschnitt einfuegt, muss man nicht manuell alle Referenzen
%% erneuern. Das Verwenden solcher symbolischer Referenzen sorgt dann
%% selber fuer die korrekte Referenzierung
%%
%%
%% Es kann sein, dass beim Compilieren des TeX-Files folgende Warnung
%% ausgegeben wird
%%
%% 	LaTeX Warning: There were undefined references.
%%
%%
%%	LaTeX Warning: Label(s) may have changed. Rerun to get ...
%%
%% Einfach der Anweisung folgen und neu compilieren.

\subsection{Unterabschnitt: Texte in \TeX}

Erzeugen wir erst etwas Text, um uns mit der Schreibweise in \LaTeX\ 
vertraut zu machen. \LaTeX\ bricht selbst\"andig Zeilen und Seiten um.
Selbst in einem Neogloismus wie dem ziemlich langen Wort aus Michael
Endes zuletzt geschriebenem Buch \glqq Der
satanarch\"aol\"ugenialkoh\"ollische Wunschpunsch\grqq\ findet \LaTeX\ eine
gute Trennstelle. Manchmal ist es aber sinnvoll, selber eine
Trennstelle vorzugeben. Dies gibt man mit $\backslash-$ im Wort im
\TeX-File selbst an, also \zB Trenn$\backslash-$stelle.
%
%
%
\\ \ \\   		% das erzeugt eine Leerzeile

L\"asst man eine Zeile im \TeX-File frei so wird automatisch die
n\"achste Zeile einger\"uckt. Wem das nicht gef\"allt, der kann im
\TeX-File den Befehl $\backslash$\texttt{parindent0pt} auskommentieren
und das damit unterdr\"ucken. 
Text mit \"a, \"o, \"U, Stra{"s}e, Caf\'e.\\[4ex]

\subsection{Unterabschnitt: mathematische Formeln}

%% Betrachten wir nun die Erzeugung mathematischer Formeln. In LaTeX
%% wir zwischen dem Text-Modus und dem mathematischen Modus
%% unterschieden. Um im laufenden Text Formeln zu verwenden, benutzt
%% man die $...$ Umgebung, z.B. wie folgt:

Dies ist ein laufender-Text mit Einsteins Formel $E = m \cdot c^2 $.
%
%
%% Natuerlich sind auch kompliziertere Formeln realisierbar.
%% Setzen wir z.B. die geometrische Reihe in eine eigene
%% Formelumgebung. Dies leitet man mit \[ ... \]  ein
\\ \ \\
Eine etwas kompliziertere Formel:
\[
     a \Frac{1 - q^{n+1}} {1-q}  =  \sum_{i=0}^{n} aq^{n} \qquad
     \text{mit\ } \quad a,q \in \mathcal{R},  q \neq 1
\]
%% Das Prinzip eines solchen Formelaufbaus ist relativ einfach.
%% Zunaechst schreiben wir unser ersten Symbol a hin. Dann folgt
%% ein Bruch. Dieser wird mit dem Befehl \Frac{}{} erzeugt. In
%% das erste Paar geschweifter Klammern schreibt man den Zaehler
%% rein, in das zweite Paar schreibt man den Nenner rein. Der
%% Zaehler enthaelt einen Exponenten. Exponenten werden immer
%% mit ^{} beschrieben. Alles was man dann in die Klammern
%% schreibt, steht dann im Exponenten. Analog gibt es natuerlich
%% auch Indizes. Diese notiert man mit _{}
%%
%%
%% Dieselbe Notation verwendet man auch fuer das Summenzeichen
%% (analog fuer das Produktzeichen \prod
%%
%%
%% Mit dem Befehl \mathcal erzeugt man ein kaligraphisches R im
%% mathematischen Modus. Der Befehl \neq erzeugt ein Ungleichheits-
%% symbol. Es gibt zahlreiche weitere Befehle zum Editieren von
%% Formeln in LaTeX. Fuer eine Uebersicht schaue man in der Literatur
%% nach, die auf der Webseite genannt ist.
%%
%%
%% Will man Formeln durchnumerieren, so kann man die equation-array
%% Umgebung benutzen.
%%
%%
\\ \ \\
Nun folgt die Bayessche Entscheidungsregel:\\

\begin{eqnarray}
\label{eq:bayes-decision-rule}
        r(X) = \underset{w_1 \dots w_N}{\mbox{argmax}}
        \bigg\{Pr(w_1^N | x_1^T)\bigg\}
        \quad \mbox{mit} \quad
        Pr(w_1^N | x_1^T) = \Frac{Pr(x_1^T | w_1^N) \cdot Pr(w_1^N)}
                            {Pr(x_1^T)}
        \quad .
\end{eqnarray}
%% Um Formeln richtig einzuruecken kann in der equation array Umgebung
%% das & Symbol verwendet werden. Gleichzeitig wollen wir die
%% Nummerierung der Formel unterdruecken. Dies geht mit einem *
%%
\\ \ \\
Beispiel f\"ur ein lineares Gleichungssystem:\\
\begin{eqnarray*}
	a_{11} x_1 + a_{12} x_2 + a_{13} x_3 &=& y_1\\
	a_{21} x_1 + a_{22} x_2		     &=& y_2\\
	a_{31} x_1			     &=& y_3 
\end{eqnarray*}

%% Statt \text kann man auch den etwas allgemeineren Befehl \mbox im
%% mathematischen Modus verweden, um reinen Text darzustellen.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%               Abschnitt 2
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Name des zweiten Abschnitts}
\subsection{Unterabschnitt}
\subsubsection{Unter-Unterabschnitt: Beispiel f\"ur Tabellen}

%% Tabellen werden mit der tabular Umgebung erzeugt. Mit der table
%% Umgebung werden wir unserer Tabelle zusaetzlich eine Ueberschrift
%% geben und ihr mit \label einen symbolischen Namen zuordnen. Der
%% Aufbau einer Tabelle ist wie folgt: zunaechst wird in das zweite
%% Paar geschweifter Klammern eingetragen, wieviele Spalten man
%% haben will (hier |l|c|r|  , also 3). Das l weist LaTeX an, dass
%% die Eintraege der Spalte 1 linkbuendig, die der Spalte 2 zentriert
%% und die Eintraege der Spalte 3 rechtsbuendig dargestellt werden.
%% Das Symbol  |  steht fuer einen vertikalen Trennstrich. Laesst
%% man diesen weg, so enthaelt die Tabelle an der entsprechenden
%% Stelle auch keine Trennlinie mehr. Der Befehl \hline erzeugt eine
%% horizontale Trennlinie und wird nach einem Zeilenumbruch \\
%% angegeben.

In der Umgebung zum Plazieren von Tabellen (und auch Grafiken) kann
man optional die Werte t, b, h und H verwenden. Damit lassen sich
Tabellen am oberen Rand (top), unten (bottom) und an der aktuellen
Position (here) plazieren. Je nach Seitenlayout funktioniert die
Angabe h nicht immer. Mittels H l\"asst sich dann die aktuelle
Position erzwingen.

\begin{table}%[H,h,t,b,p]
\begin{center}
\caption{Beispiel Tabelle} % Tabellen_über_schrift
\label{tab:beispiel}
\vspace{2ex}
\begin{tabular}{|l|c|r|}
\hline
links		&	zentriert	&	rechts  \\ \hline\hline
links2		&	zentriert2	&	rechts2 \\ \hline
XXXXXXXXXXX	& XXXXXXXXXXXXXXXXXXXXX	& XXXXXXXXXXXX  \\ \hline

\end{tabular}
\end{center}
\end{table}

\subsection{Unterabschnitt: Beispiel f\"ur Grafiken}
Grafiken werden \"ublicherweise mit dem (frei verf\"ugbaren) Programm
tgif erstellt. Alternativ hierzu l\"asst sich auch das Programm xfig
verwenden. tgif ist zwar nicht ganz einfach zu bedienen, stellt dem
(erfahrenen) Anwender aber wesentlich mehr M\"oglichkeiten zur
Gestaltung von Grafiken zur Ver\"ugung. Die Konvertierung der *.obj
Dateien nach encapsulated PostScript erfolgt \"uber die Kommandozeile mit
\begin{center}
	tgif \quad -print \quad -eps \quad $<$Datei$>$.obj
\end{center}

\begin{figure}
\begin{center}
  \includegraphics[width=9cm]{./architektur}           % keine file extension angeben
  \caption{Architektur eines Spracherkennungssystems}  % Bild_unter_schrift
  \label{fig:Erkenner-Architektur}
\end{center}
\end{figure}

\newpage
Nun erzeugen wir mittels \textsc{Bib}\TeX\ eine
Literaturliste. Dazu erzeugen wir uns ein File namens
$<$Dateiname$>$.bib, dass denselben Dateinamen wie unser \TeX-File
hat. Dieses wird dann mit
\begin{center}
	bibtex \quad  $<$Dateiname$>$
\end{center}
compiliert. Die im laufenden Text vorkommenden symbolischen
Literaturverweise zeigen dann auf den entsprechenden Eintrag
im Literaturverzeichnis. Dies geschieht mit dem Befehl
$\backslash$cite\{\}. Das sieht dann \zB wie folgt aus.
\\ \ \\
Zitat aus \cite{Name:1991}, Zitat aus \cite{Ende:1989}, Zitat aus
\cite{Name:1996}. Wichtig ist, das zuerst das \TeX-File compiliert
wird und dann das \textsc{Bib}\TeX-File. Da sich dann die Referenzen
meist \"andern, ist ein erneutes Compilieren des \TeX-Files notwendig.
\\ \ \\
Es gibt zahlreiche style-files, mit denen man das Layout eines
Literaturverzeichnisses \"andern und die Darstellung der
Literatur-Refererenzen modifizieren kann.
\\ \ \\
Letzter Hinweis: \LaTeX-Neulinge sollten auch die Kommentare im
\TeX-File lesen.

\addcontentsline{toc}{section}{Literaturverzeichnis}
\bibliographystyle{plain}
\bibliography{ausarbeitung_vorlage}

\end{document}

### 2016-10-12 paper/paper.tex
clear comments, start with introductory text

@@
 \usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[ngerman]{babel}                     % otherwise use british or american
+\usepackage[american]{babel}                     % otherwise use british or american
 \usepackage{theorem}                            % instead of \usepackage{amsthm}
@@
 \usepackage{hyperref}
+\usepackage[utf8]{inputenc}
 
@@
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Allgemeines ueber LaTeX
-%% =======================
-%%
-%% 
-%% LaTeX (sprich: 'latech') ist ein Textsatzsystem im Gegensatz zu
-%% einem Textverarbeitungsprogramm wie beispielsweise Word. Wie im
-%% professionellen Buchdruck werden Texte nicht nur geschrieben,
-%% sondern muessen auch 'gesetzt' werden. Da LaTeX wie eine Program-
-%% miersprache gehandhabt wird, ermoeglicht es dem Benutzer auch
-%% eine groesstmoegliche Freiheit in der Gestaltung einer Seite.
-%% 
-%%
-%%
-%% Eine der Staerken von LaTeX bildet die Vielzahl von Befehlen
-%% zur Beschreibung mathematischer Formeln. Wie das im einzelnen
-%% geht wird spaeter unten beschrieben.
-%%
-%%
-%%
-%% Die Vorteile von LaTeX haben natuerlich ihren Preis. Anders als
-%% bei Word kann man das Resultat nicht direkt zur Eingabezeit
-%% begutachten. Statt dessen muss das TeX-File (das ist die Datei,
-%% die Ihr gerade lest und immer auf den Namen <Dateiname>.tex enden
-%% muss) zuerst mit dem Befehl
-%%
-%%
-%%                  latex  <Dateiname>.tex
-%%         
-%%
-%% compiliert werden. Dabei wird ein file erzeugt, das den Namen
-%% 
-%%
-%%    --->          <Dateiname>.dvi          
-%%
-%%
-%% traegt. Die Endung dvi steht fuer 'device independent'und enthaelt
-%% bereits die geraeteunabhaengige Darstellung, die man sich mit dem
-%% Aufruf
-%%
-%%
-%%    --->          xdvi  <Dateiname>.dvi
-%%
-%%
-%% anschauen kann. Um die *.dvi Datei nach PostScript zu konvertieren,
-%% verwendet man den Aufruf
-%%
-%%
-%%    --->          dvips <Dateiname>.dvi -o <Name>.ps
-%%
-%%
-%% Diese Datei kann von jedem PostScript faehigen Drucker ausgedruckt
-%% werden.
-
-%% Aufbau eines LaTeX Dokuments
-%% ============================
-%%
-%% Jedes LaTeX Dokument beginnt mit der Angabe der Dokumentklasse in
-%% geschweiften Klammern (ganz oben in der ersten Zeile). Dies ist
-%% hier die article Klasse. In die eckigen Klammern und den
-%% nachfolgenden usepackages koennen weitere Pakete eingebunden
-%% werden, die neben dem reinen LaTeX Standard zusaetzliche Befehle
-%% zur Verfueugung stellen.
-%%
-%% Die folgenden Befehle fuehren neue Umgebungen ein, stellen
-%% Macros zur Verfuegung und modifizieren das Seitenlayout.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%       Jetzt geht's los
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
 
-%% Mit dem Befehl \begin{document} weist man LaTeX an, dass das Dokument
-%% beginnt. Ganz am Ende des Dokuments gibt es einen entsprechenden Befehl
-%% \end{document}
-%%
-%% Aber beginnen wir zunaechst mit der Titelseite unseres Seminars. Eine
-%% Titelseite soll natuerlich die folgenden Dinge enthalten:
-%%
-%%   1. Titel des Seminars
-%%   2. Bezeichnung des Lehrstuhls, an dem das Seminar angefertigt wurde
-%%   3. Name des Seminaristen
-%%   4. Name des Betreuers
-%%
-%% Wie realisiert man das in LaTeX?
-%%
-%% Zunaechst einmal soll eine Titelseite natuerlich keine Numerierung
-%% erhalten. LaTeX kann selbstaendig Zeilen und Seiten umbrechen und
-%% fuehrt auch die Seitenzaehlung selbstaendig durch. Da wir die Zaehlung
-%% fuer die Titelseite unterruecken wollen, geben wir mit dem Befehl
-%% \pagestyle{empty} vor, dass LaTeX weder Seitenzahl noch Kopf- oder
-%% Fusszeilen einfuegt. Da wir den Titel gerne zentriert haetten, benutzen
-%% wir die Befehle \begin{center}  ...  \end{center} . Alle Eingaben, die
-%% wir innerhalb dieser 'center' Umgebung eingeben, werden dann automatisch
-%% von LaTeX zentriert.
-%%
-%% Der Befehl \LARGE beschreibt die Font-Groesse. Fuer die Groesse des
-%% Schriftsatzes gibt es mehrere Befehle. Diese sind in aufsteigender
-%% Reihenfolge:
-%%
-%%	\tiny 		\normalsize	\huge
-%%	\scriptsize   	\large		\Huge
-%%	\footnotesize	\Large
-%%	\small		\LARGE
-%%
-%%
-%% Um den Font zu aendern, stehen folgende Befehle zur Verfuegung:
-%%
-%%	\textrm{}	roman font
-%%	\textsf{}	sans  serif
-%%	\texttt{}	typewriter
-%%	\textbf{}	bold series
-%%	\textit{}	italic		(kursiv)
-%%	\textsc{}	SMALL CAPS
-%%	\textnormal{}	typeset text in the normal font
-%%
-%%
-%% Mit dem Kommando \\  gibt man EXPLIZIT(!) an, dass man eine Zeile
-%% umbrechen will (wie gesagt, normalerweise macht LaTeX das selbst).
-%% Optional kann man hier \\ einen Wert angeben wie z.B. \\[4ex]
-%% Damit legt man fest, wieviel Zwischenraum bis zur naechsten
-%% Zeile bestehen soll. Das 'ex' steht fuer die Groesse des Symbols x.
-%% Hier kann man auch die Einheiten 'pt' fuer Punkte, cm, ... benutzen.
-%%
-%% Der \vfill Befehl fuegt soviel Leerraum ein, das der restliche Text
-%% bis an den unteren Rand gedrueckt wird (die Leerzeile davor wird
-%% benoetigt). Der \newpage Befehl erzeugt einen EXPLIZITEN(!) Seiten-
-%% umbruch. Da wir eine Leerseite haben wolllen wird mit '\ ' ein
-%% Space-Symbol auf die Seite eingefuegt, so dass wir wieder einen
-%% \newpage Befehl einsetzen koennen.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%               Title
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \pagestyle{empty}
@@
 
-    Rheinisch-Westf\"alische Technische Hochschule Aachen \\
-    Lehrstuhl f\"ur Informatik 6 \\
+    RWTH Aachen University\\
+    Chair of Computer Science 6 \\
     Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Seminar Titel im SS 20XX\\[12ex]                          % auch Seminar Titel und Datum ändern!!!
+    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
 
     \LARGE
-    \textbf{Titel des Vortrages} \\[6ex]
-    \textit{Vorname Name} \\[6ex]
-    \Large
-    Matrikelnummer 123 456 \\[6ex]
-    Datum des Vortrages
+    \textbf{Deep Directed Generative Models} \\[6ex]
+    \textit{André Merboldt} \\[6ex]
+    \today
 
     \vfill
-    \Large Betreuer: Vorname Name
+    \Large Supervisor: Tobias Menne
 
@@
 
+% blank page
 \newpage
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%               Inhaltsverzeichnis / Tabellenverzeichnis / Abbildungsverz.
-%
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% contents
 \pagestyle{headings}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%               Abschnitt 1
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Name des ersten Abschnitts}
-\label{sec:Abschnitt-1}
+%----------------------------------
+% ABSTRACT
+%----------------------------------
+\section{Abstract}
+\label{sec:abstract}
+In this seminar paper we provide an overview and introduction to directed generative models,
+which are able to produce data following a distribution (...?).
+% more info what generative models try to achieve
+After we have introduced several architectures, we will explore sigmoid belief networks (SBN)
+as an early generative model.\\
 
-%% Der Befehl \section{} erstellt den ersten Abschnitt. Unter-
-%% abschnitte werden mit \subsection{} generiert. Mit dem \label
-%% Befehl erzeugt man eine symbolische Marke auf den Abschnitt.
-%% Will man spaeter im Text auf den ersten Abschnitt referenzieren,
-%% so gibt man das wie folgt an:
+Then we will turn to more recent proposals, in particular variational autoencoders (VAE) and
+generative adversarial networks (GAN).
+Variants of GANs have been shown to produce images which are able to fool a human discriminator 40\% of the time (citation needed, LAPGAN? or DCGAN?).
+Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.
 
-Dies ist eine Referenz auf Abschnitt~\ref{sec:Abschnitt-1}
+%----------------------------------
+% Motivation
+%----------------------------------
+\section{Motivation}
+\label{sec:motivation}
+Generative models have been proposed as one of the most promising approaches towards
+learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).
 
-%% LaTeX uebernimmt damit automatisch die Nummerierung der Referenzen.
-%% Der Vorteil liegt darin, dass wenn man irgendwann mitten im Text einen
-%% neuen Abschnitt einfuegt, muss man nicht manuell alle Referenzen
-%% erneuern. Das Verwenden solcher symbolischer Referenzen sorgt dann
-%% selber fuer die korrekte Referenzierung
-%%
-%%
-%% Es kann sein, dass beim Compilieren des TeX-Files folgende Warnung
-%% ausgegeben wird
-%%
-%% 	LaTeX Warning: There were undefined references.
-%%
-%%
-%%	LaTeX Warning: Label(s) may have changed. Rerun to get ...
-%%
-%% Einfach der Anweisung folgen und neu compilieren.
+In recent years supervised learning has yielded impressive results,
+however for these models to suceed huge amount of labelled data is needed.
+Unsupervised learning, that is learning from unlabelled data, has been expressed
+as one hurdles toward general articial intelligence (citation needed).
+Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
 
-\subsection{Unterabschnitt: Texte in \TeX}
+In generative models the approach is different in a way that
+we search for a good internal representation of the data.
+This resembles the way humans learn about the world where we incrementally build
+a model of the world.
 
-Erzeugen wir erst etwas Text, um uns mit der Schreibweise in \LaTeX\ 
-vertraut zu machen. \LaTeX\ bricht selbst\"andig Zeilen und Seiten um.
-Selbst in einem Neogloismus wie dem ziemlich langen Wort aus Michael
-Endes zuletzt geschriebenem Buch \glqq Der
-satanarch\"aol\"ugenialkoh\"ollische Wunschpunsch\grqq\ findet \LaTeX\ eine
-gute Trennstelle. Manchmal ist es aber sinnvoll, selber eine
-Trennstelle vorzugeben. Dies gibt man mit $\backslash-$ im Wort im
-\TeX-File selbst an, also \zB Trenn$\backslash-$stelle.
-%
-%
-%
-\\ \ \\   		% das erzeugt eine Leerzeile
+%----------------------------------
+% Overview of Directed Generative Models
+%----------------------------------
+\section{Overview of Directed Generative Models}
+\label{sec:overview}
 
-L\"asst man eine Zeile im \TeX-File frei so wird automatisch die
-n\"achste Zeile einger\"uckt. Wem das nicht gef\"allt, der kann im
-\TeX-File den Befehl $\backslash$\texttt{parindent0pt} auskommentieren
-und das damit unterdr\"ucken. 
-Text mit \"a, \"o, \"U, Stra{"s}e, Caf\'e.\\[4ex]
+\paragraph{Variational Autoencoders\cite{vae:2013}}
+\label{par:overview_vae}
+Autoencoders comprise of two components, the encoder and decoder
+where the encoder takes input data and translates it into a latent structure while
+the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
 
-\subsection{Unterabschnitt: mathematische Formeln}
+In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
+% intractable -> use neural network, which we can train
+% --> no closed form possible
+We use the KL-Divergence to train the network.
 
-%% Betrachten wir nun die Erzeugung mathematischer Formeln. In LaTeX
-%% wir zwischen dem Text-Modus und dem mathematischen Modus
-%% unterschieden. Um im laufenden Text Formeln zu verwenden, benutzt
-%% man die $...$ Umgebung, z.B. wie folgt:
+% TODO read chapter 19, approximate inference
+In VAEs however, this
 
-Dies ist ein laufender-Text mit Einsteins Formel $E = m \cdot c^2 $.
-%
-%
-%% Natuerlich sind auch kompliziertere Formeln realisierbar.
-%% Setzen wir z.B. die geometrische Reihe in eine eigene
-%% Formelumgebung. Dies leitet man mit \[ ... \]  ein
-\\ \ \\
-Eine etwas kompliziertere Formel:
-\[
-     a \Frac{1 - q^{n+1}} {1-q}  =  \sum_{i=0}^{n} aq^{n} \qquad
-     \text{mit\ } \quad a,q \in \mathcal{R},  q \neq 1
-\]
-%% Das Prinzip eines solchen Formelaufbaus ist relativ einfach.
-%% Zunaechst schreiben wir unser ersten Symbol a hin. Dann folgt
-%% ein Bruch. Dieser wird mit dem Befehl \Frac{}{} erzeugt. In
-%% das erste Paar geschweifter Klammern schreibt man den Zaehler
-%% rein, in das zweite Paar schreibt man den Nenner rein. Der
-%% Zaehler enthaelt einen Exponenten. Exponenten werden immer
-%% mit ^{} beschrieben. Alles was man dann in die Klammern
-%% schreibt, steht dann im Exponenten. Analog gibt es natuerlich
-%% auch Indizes. Diese notiert man mit _{}
-%%
-%%
-%% Dieselbe Notation verwendet man auch fuer das Summenzeichen
-%% (analog fuer das Produktzeichen \prod
-%%
-%%
-%% Mit dem Befehl \mathcal erzeugt man ein kaligraphisches R im
-%% mathematischen Modus. Der Befehl \neq erzeugt ein Ungleichheits-
-%% symbol. Es gibt zahlreiche weitere Befehle zum Editieren von
-%% Formeln in LaTeX. Fuer eine Uebersicht schaue man in der Literatur
-%% nach, die auf der Webseite genannt ist.
-%%
-%%
-%% Will man Formeln durchnumerieren, so kann man die equation-array
-%% Umgebung benutzen.
-%%
-%%
-\\ \ \\
-Nun folgt die Bayessche Entscheidungsregel:\\
+\paragraph{Generative Adversarial Networks}
+\label{par:overview_gan}
+GANs are a game theoretic approach to generative modelling
+in which two networks compete against each other.
+The discriminator $D$ is trying to distinguish the generated samples
+by the generator $G$ from real world data.
+The goal for $G$ however is to produce data as realistic as possible.\\
+During the early phase of training, $G$ produces random noise which
+is easy to differentiate from groundtruth data.
+Both networks are trained in parallel, where both networks try to get better
+at their respective objective.
+In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
+case the discriminator will only be able to guess correctly with a 50\% chance.
+This plateau is called Nash equilibrium (citation needed).
 
-\begin{eqnarray}
-\label{eq:bayes-decision-rule}
-        r(X) = \underset{w_1 \dots w_N}{\mbox{argmax}}
-        \bigg\{Pr(w_1^N | x_1^T)\bigg\}
-        \quad \mbox{mit} \quad
-        Pr(w_1^N | x_1^T) = \Frac{Pr(x_1^T | w_1^N) \cdot Pr(w_1^N)}
-                            {Pr(x_1^T)}
-        \quad .
-\end{eqnarray}
-%% Um Formeln richtig einzuruecken kann in der equation array Umgebung
-%% das & Symbol verwendet werden. Gleichzeitig wollen wir die
-%% Nummerierung der Formel unterdruecken. Dies geht mit einem *
-%%
-\\ \ \\
-Beispiel f\"ur ein lineares Gleichungssystem:\\
-\begin{eqnarray*}
-	a_{11} x_1 + a_{12} x_2 + a_{13} x_3 &=& y_1\\
-	a_{21} x_1 + a_{22} x_2		     &=& y_2\\
-	a_{31} x_1			     &=& y_3 
-\end{eqnarray*}
+%----------------------------------
+% Sigmoid Belief Networks
+%----------------------------------
+\section{Sigmoid Belief Networks}
+\label{sec:sbn}
 
-%% Statt \text kann man auch den etwas allgemeineren Befehl \mbox im
-%% mathematischen Modus verweden, um reinen Text darzustellen.
+%----------------------------------
+% Variational Autoencoders
+%----------------------------------
+\section{Variational Autoencoders}
+\label{sec:vae}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%               Abschnitt 2
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Name des zweiten Abschnitts}
-\subsection{Unterabschnitt}
-\subsubsection{Unter-Unterabschnitt: Beispiel f\"ur Tabellen}
+\subsection{Architecture}
+\label{sub:vae_architecture}
 
-%% Tabellen werden mit der tabular Umgebung erzeugt. Mit der table
-%% Umgebung werden wir unserer Tabelle zusaetzlich eine Ueberschrift
-%% geben und ihr mit \label einen symbolischen Namen zuordnen. Der
-%% Aufbau einer Tabelle ist wie folgt: zunaechst wird in das zweite
-%% Paar geschweifter Klammern eingetragen, wieviele Spalten man
-%% haben will (hier |l|c|r|  , also 3). Das l weist LaTeX an, dass
-%% die Eintraege der Spalte 1 linkbuendig, die der Spalte 2 zentriert
-%% und die Eintraege der Spalte 3 rechtsbuendig dargestellt werden.
-%% Das Symbol  |  steht fuer einen vertikalen Trennstrich. Laesst
-%% man diesen weg, so enthaelt die Tabelle an der entsprechenden
-%% Stelle auch keine Trennlinie mehr. Der Befehl \hline erzeugt eine
-%% horizontale Trennlinie und wird nach einem Zeilenumbruch \\
-%% angegeben.
+\subsection{Extensions}
+\label{sub:vae_extensions}
 
-In der Umgebung zum Plazieren von Tabellen (und auch Grafiken) kann
-man optional die Werte t, b, h und H verwenden. Damit lassen sich
-Tabellen am oberen Rand (top), unten (bottom) und an der aktuellen
-Position (here) plazieren. Je nach Seitenlayout funktioniert die
-Angabe h nicht immer. Mittels H l\"asst sich dann die aktuelle
-Position erzwingen.
+\subsubsection{Deep Recurrent Attention Writer (DRAW)}
+\label{ssub:deep_recurrent_attention_writer_draw_}
 
-\begin{table}%[H,h,t,b,p]
-\begin{center}
-\caption{Beispiel Tabelle} % Tabellen_über_schrift
-\label{tab:beispiel}
-\vspace{2ex}
-\begin{tabular}{|l|c|r|}
-\hline
-links		&	zentriert	&	rechts  \\ \hline\hline
-links2		&	zentriert2	&	rechts2 \\ \hline
-XXXXXXXXXXX	& XXXXXXXXXXXXXXXXXXXXX	& XXXXXXXXXXXX  \\ \hline
+%----------------------------------
+% Generative Adversarial Netoworks
+%----------------------------------
+\section{Generative Adversarial Networks}
+\label{sec:gan}
 
-\end{tabular}
-\end{center}
-\end{table}
+\subsection{Architecture}
+\label{sub:gan_arch}
 
-\subsection{Unterabschnitt: Beispiel f\"ur Grafiken}
-Grafiken werden \"ublicherweise mit dem (frei verf\"ugbaren) Programm
-tgif erstellt. Alternativ hierzu l\"asst sich auch das Programm xfig
-verwenden. tgif ist zwar nicht ganz einfach zu bedienen, stellt dem
-(erfahrenen) Anwender aber wesentlich mehr M\"oglichkeiten zur
-Gestaltung von Grafiken zur Ver\"ugung. Die Konvertierung der *.obj
-Dateien nach encapsulated PostScript erfolgt \"uber die Kommandozeile mit
-\begin{center}
-	tgif \quad -print \quad -eps \quad $<$Datei$>$.obj
-\end{center}
+\subsection{Training}
+\label{sub:gan_training}
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=9cm]{./architektur}           % keine file extension angeben
-  \caption{Architektur eines Spracherkennungssystems}  % Bild_unter_schrift
-  \label{fig:Erkenner-Architektur}
-\end{center}
-\end{figure}
+\subsection{Stability and Performance}
+\label{sub:gan_stability}
 
-\newpage
-Nun erzeugen wir mittels \textsc{Bib}\TeX\ eine
-Literaturliste. Dazu erzeugen wir uns ein File namens
-$<$Dateiname$>$.bib, dass denselben Dateinamen wie unser \TeX-File
-hat. Dieses wird dann mit
-\begin{center}
-	bibtex \quad  $<$Dateiname$>$
-\end{center}
-compiliert. Die im laufenden Text vorkommenden symbolischen
-Literaturverweise zeigen dann auf den entsprechenden Eintrag
-im Literaturverzeichnis. Dies geschieht mit dem Befehl
-$\backslash$cite\{\}. Das sieht dann \zB wie folgt aus.
-\\ \ \\
-Zitat aus \cite{Name:1991}, Zitat aus \cite{Ende:1989}, Zitat aus
-\cite{Name:1996}. Wichtig ist, das zuerst das \TeX-File compiliert
-wird und dann das \textsc{Bib}\TeX-File. Da sich dann die Referenzen
-meist \"andern, ist ein erneutes Compilieren des \TeX-Files notwendig.
-\\ \ \\
-Es gibt zahlreiche style-files, mit denen man das Layout eines
-Literaturverzeichnisses \"andern und die Darstellung der
-Literatur-Refererenzen modifizieren kann.
-\\ \ \\
-Letzter Hinweis: \LaTeX-Neulinge sollten auch die Kommentare im
-\TeX-File lesen.
+\subsubsection{Freeze Learning}
+\label{ssub:gan_freeze_learning}
 
-\addcontentsline{toc}{section}{Literaturverzeichnis}
+\subsubsection{Feature Matching}
+\label{ssub:gan_feature_matching}
+
+\subsubsection{Minibatch discrimination}
+\label{ssub:gan_minibatch_discrimination}
+
+\subsubsection{Historical Averaging}
+\label{ssub:gan_historical_averaging}
+
+\subsubsection{Label Smoothing}
+\label{ssub:gan_label_smoothing}
+
+\subsection{Application}
+\label{sub:gan_application}
+
+\subsection{Extensions}
+\label{sub:gan_extensions}
+
+\subsubsection{LAPGAN}
+\label{ssub:lapgan}
+
+\subsubsection{DCGAN}
+\label{ssub:dcgan}
+
+%----------------------------------
+% More Generative Directed Models
+%----------------------------------
+\section{More Generative Directed Models}
+\label{sec:more}
+
+\subsection{Generative Moment Matching Networks}
+\label{sub:more_mmn}
+
+\subsection{Auto-Regressive Networks}
+\label{sub:more_arn}
+
+See \cite{gan_openai:2016} and \cite{gan_training:2016}.
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+Generative models have shown great promise to tackle interesting problems.
+Academic work is moving incredibly fast and many new ideas and combinations
+are being proposed almost daily (citation needed, meta paper?).
+For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
+which produced way larger images than GANs alone.
+
+% not relevant, but wanted to write about it :p
+As mentioned in the motivation, further advances to generative architectures are
+likely to solve current issues with depending on large labelled data.
+Another large and increasing area of research is semi-supervised learning
+(combining learning from labelled data as well as unlabelled data)
+and also one-shot learning, which is where humans excel and machines fail currently.
+\\\\
+Exciting times :)
+
+\addcontentsline{toc}{section}{References}
 \bibliographystyle{plain}
-\bibliography{ausarbeitung_vorlage}
+\bibliography{paper}
 

### 2016-11-22 paper/paper.tex
add VAE first attempt

@@
 
+%Variational Autoencoders (VAE) have been a popular choice for unsupervised learning of complicated distributions (citation needed) and generative modelling.
+In order to generate data from unknown and mostly intractable distributions, we need approximations.
+There are basically two approaches for sampling from these distributions,
+first there are approximate samples (MCMC, gibbs sampling, etc) which try to directly approximate $p(x)$.
+Variational Autoencoders instead try to match an easier to compute distribution $q(x)$ to $p(x)$.
+By using this approach, VAEs are computationally less intensive (citation?) but have the drawback of being more restricted in their modelling approach.
+Practically, this means that with more computation MCMC methods approach $p(x)$, while there is no such guarantee for variational methods (--> not for all).
+
+VAE can be learned with just backpropagation (paper,TODO), but they differ from denoising and sparse autoencoders due to the different loss function.
+
+%VAEs are built on top of neural networks and are designed in a way to allow training with gradient-based methods.
+%Learning and inference are reasonable efficient and relatively easy to implement and show decent results, but have been overshadowed by more recent adversarial approaches (citation needed!!,see \ref{sec:gan}).
+
 \subsection{Architecture}
 \label{sub:vae_architecture}
+VAE just like other autoencoders encode the input data into a smaller latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
+But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.
+This allows to sample from this distribution and generate output which will look similar to the data on which the VAE has been trained.
+\ref{fig:vae_architecture}.
+
+\subsubsection{Reparameterization Trick}
+To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
+Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
+Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
+\begin{equation}
+  \label{eq:rep_trick}
+  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
+\end{equation}
+This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
+
+The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
+In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
+
+Variational Autoencoders achieve this by using a measurement of inequality between $q(x)$ and $p(x)$ called Kullback-Leibler Divergence (KL divergence).
+
+%VAEs are unlike \emph{traditional} auto-encoders (sparse autoencoders or denoising autoencoders for example) because they have a different architecture
+%VAEs resembles traditional autoencoders
+%\begin{equation}
+%P(x) = \int P(x|z;\theta)P(z) dz
+  %\label{eq:max_likelihood}
+%\end{equation}
+
+We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
+The prior distribution $p_\theta(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
+The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2013} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
+Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
+\begin{equation}
+  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p_\theta(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
+\end{equation}
+\paragraph{Derivation of lower bound}
+We want to minimize the KL divergence from $P(z|x)$ to $Q(z)$ for some arbitrary $Q$.
+  \begin{align*}
+    \label{eq:der_lower_bound}
+    \mathcal{D}[Q(z)||P(z|x)] &= \mathbb{E}_{z~q_\theta}(\log Q(z) - \log P(z|x))
+    &= \mathbb{E}_{z ~ q_\theta}[]
+  \end{align*}
+Where $D_{KL}(q_\theta(z|x) || p_\theta(z))$ is defined as the Kullback-Leibler divergence, which can be intuitively thought of as a measurement for the distance between two probabilistic distributions (even though it lacks the symmetric property).
+
+\subsection{Learning}
+\label{sub:vae_learning}
+In practice, the usual choice for $q_\theta(z|x)$ is the multivariate gaussian distribution.
+$$
+q_\theta(z|x) = \mathcal{N}(z|\mu(X;\vartheta), \Sigma(X;\vartheta))
+$$
+Because we assume that $P(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
+\begin{align*}
+  \mathcal{D}_{\mathrm{KL}}\big[Q(z) || P(z|x)\big] &= \mathbb{E}\big[\log \frac{p}{q}\big]\\
+  \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]
+  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)
+\end{align*}
+
+\subsection{Inference}
+\label{sub:vae_inference}
+
+\subsection{Performance}
+\label{sub:vae_performance}
 
@@
 
-\subsubsection{Deep Recurrent Attention Writer (DRAW)}
-\label{ssub:deep_recurrent_attention_writer_draw_}
+\subsubsection{Deep Recurrent Attention Writer}
+\label{ssub:vae_deep_recurrent_attention_writer}
+
+\subsubsection{Importance Weighted Autoencoder}
+\label{ssub:vae_importance_weighted_autoencoder}
+
+\subsubsection{Conditional VAE}
+\label{ssub:vae_conditional_vae}
 

### 2016-11-22 paper/paper.tex
started writing on GAN

@@
 
+Generative Adversarial Networks (GAN) are a recently proposed framework by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
+GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
+The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
+Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
+% --> declare distribution p_data, p_model, x, z
+During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
+In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
+We don't have access to that distribution.
+
 \subsection{Architecture}
 \label{sub:gan_arch}
+
+\subsubsection{Generator Network}
+The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
+
+\subsubsection{Discriminator Network}
 

### 2016-11-28 paper/paper.tex
more stuff on GANs

@@
 \usepackage{amsmath, amssymb, latexsym}  % math stuff
+\usepackage{amsfonts}
 \usepackage{amsopn}                             % um mathe operatoren zu deklarieren
@@
 \usepackage[utf8]{inputenc}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+
+\newcommand*\Let[2]{\State #1 $\gets$ #2}
 
@@
 
-Generative Adversarial Networks (GAN) are a recently proposed framework by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
+Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
 GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
@@
 The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
+This process can be formally written as minimizing the following cost function.
+$$
+\mathcal{C}_g = \log(1 - D(G(z)))
+$$
+$z$ is produced by sampling from the noise prior $p_g(z)$.\\
+
+Note that this loss function is fully differentiable which means it can be trained with backpropagation.
 
 \subsubsection{Discriminator Network}
+The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
+This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
+$$
+\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
+$$
 
@@
 \label{sub:gan_training}
+The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
+
+Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
+$$
+\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
+$$
+In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
+\begin{algorithm}
+  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
+  \label{alg:gan_online}
+  \begin{algorithmic}[1]
+    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
+    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
+    \For{number of training iterations}
+      \Let{$z_d$}{sample from noise prior $p_g(z)$}
+      \Let{$z_g$}{sample from noise prior $p_g(z)$}
+      \Let{$x$}{example from training set}
+      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
+      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
+    \EndFor
+  \end{algorithmic}
+\end{algorithm}
+
+In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
+Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
+Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
+Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
 
@@
 
+\subsubsection{Normalizing Flows}
+\label{ssub:gan_nf}
+
 \subsection{Application}
 \label{sub:gan_application}
-
+GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
 \subsection{Extensions}

### 2016-11-28 paper/paper.tex
more on VAE (introduction)

@@
 \label{sec:vae}
+Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
+To define VAEs, we first declare the observed variables $x$, as well as latent variables $z$.
+VAEs are directed graphical models where we wish to infer \emph{good} low-dimensional latent variables from the input data.
+In this setup we define $p(x,z)$ as the joint probability distribution over both input and latent variables.
+Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
+
+In this setup we have observed variables $x$ from which we would like to infer latent variables $z$
 

### 2016-11-29 paper/paper.tex
added sigmoid belief nets (short intro)

@@
 \label{sec:sbn}
+Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
+In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
+$$
+p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
+$$
+
+While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
+% show why?
+
+There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
 

### 2016-11-29 paper/paper.tex
more text on VAE

@@
 \label{par:overview_vae}
-Autoencoders comprise of two components, the encoder and decoder
-where the encoder takes input data and translates it into a latent structure while
-the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
+Variational Autoencoders (VAE) use two networks called encoder and decoder where the encoder transform the given data into a latent space and the decoder translates this latent space back into the data space.
+The input and output space are usually high-dimensional data spaces, for example images while the latent space is low-dimensional in comparison.
+Like other autoencoders, VAEs try to reconstruct the input data using the latent variables as closely as possible.
+But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
+This encouragement is done using a variational lower bound on the distribution output by the encoder network.
 
-In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
+%where the encoder takes input data and translates it into a latent structure while
+%the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
+
+%In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
 % intractable -> use neural network, which we can train
 % --> no closed form possible
-We use the KL-Divergence to train the network.
+%We use the KL-Divergence to train the network.
 
 % TODO read chapter 19, approximate inference
-In VAEs however, this
+%In VAEs however, this
 
@@
 \ref{fig:vae_architecture}.
+
+\subsection{Loss function}
+The loss function of the VAE consists of two terms, the negative log-likelihood or the reconstruction error $\mathcal{C}_{\mathrm{rec}}$ and the second term $\mathcal{C}_{\mathrm{KL}}$ serves as the regularizer.
+We will discuss both terms in detail below.
+
+\paragraph{Reconstruction error}
+$\mathcal{C}_{\mathrm{rec}}$ measures how well the decoder reconstructs the input data using the latent space $z$.
+This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\varphi(z | x_i)$.
+$\mathcal{C}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+
+\paragraph{KL divergence (regularization)}
+The second term in $\mathcal{C}$ is the KL divergence between the approximated distribution $q_\varphi(z|x)$ and $p(z)$.
+This serves as the regularizer in the VAE as it encourages $q_\varphi(z|x)$ to match $p(z)$.\\
+% move to architecture:
+%In the original VAE paper, the unit gaussian distribution with diagonal variance was proposed for $p(z)$ and although there have been multiple proposals (citations, multi-modal etc) for different distributions we will stick with $\mathcal{N}(0,I)$ for the sake of simplicity.
+% ---
+
+For two probability distributions $P$ and $Q$, the KL divergence $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can informally be described as the amount of information which is lost when using $Q$ to represent $P$.
+The KL divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
+$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the VAE to let $q_\varphi(z|x)$ approach $p(z)$.
 

### 2016-11-30 paper/chapters/abstract.tex
moved abstract to file

--- 2016-11-28 paper/paper.tex
@@
-\documentclass[twoside,11pt,a4paper]{article}
-
-% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx,curves,float,rotating}
-
-\usepackage{amsmath, amssymb, latexsym}  % math stuff
-\usepackage{amsfonts}
-\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[american]{babel}                     % otherwise use british or american
-\usepackage{theorem}                            % instead of \usepackage{amsthm}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage[utf8]{inputenc}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-
-\newcommand*\Let[2]{\State #1 $\gets$ #2}
-
-% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{xspace}                             % context sensitive space after macros
-\makeatletter 
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
-\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
-\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
-\def\etc{{etc}\onedot} \def\vs{{vs}\onedot} 
-\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
-\def\etal{{et al}\onedot}
-\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
-\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%	Macros fuer neue Umgebungen
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-\newlength{\textwd}
-\newlength{\oddsidemargintmp}
-\newlength{\evensidemargintmp}
-\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
-\newlength{\textht}
-\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
-\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-
-\newenvironment{deflist}[1][\quad]%
-{  \begin{list}{}{%
-      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
-      \settowidth{\labelwidth}{\textbf{#1}}%
-      \setlength{\leftmargin}{\labelwidth}
-      \addtolength{\leftmargin}{\labelsep}}}
-{  \end{list}}
-
-\newenvironment{Quote}% Definition of Quote
-{  \begin{list}{}{%
-      \setlength{\rightmargin}{0pt}}
-      \item[]\ignorespaces}
-{\unskip\end{list}}
-
-\theoremstyle{break}
-\theorembodyfont{\itshape}	
-\theoremheaderfont{\scshape}
-
-\newtheorem{Cor}{Corollary}
-\newtheorem{Def}{Definition}
-%\newtheorem{Def}[Cor]{Definition}
-
-\newcolumntype{.}{D{.}{.}{-1}}
-
-\pagestyle{headings}
-\textwidth 15cm
-\textheight 23cm
-\oddsidemargin 1cm
-\evensidemargin 0cm
-%\parindent 0mm
-
-\begin{document}
-
-\pagestyle{empty}
-
-\begin{center}
-
-    RWTH Aachen University\\
-    Chair of Computer Science 6 \\
-    Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
-
-    \LARGE
-    \textbf{Deep Directed Generative Models} \\[6ex]
-    \textit{André Merboldt} \\[6ex]
-    \today
-
-    \vfill
-    \Large Supervisor: Tobias Menne
-
-\end{center}
-
-% blank page
-\newpage
-\ 
-\newpage
-
-% contents
-\pagestyle{headings}
-\tableofcontents
-\listoftables
-\listoffigures
-\newpage
-\pagestyle{empty}
-\ 
-\newpage
-\pagestyle{headings}
-
-%----------------------------------
-% ABSTRACT
-%----------------------------------
 \section{Abstract}
@@
 Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.
-
-%----------------------------------
-% Motivation
-%----------------------------------
-\section{Motivation}
-\label{sec:motivation}
-Generative models have been proposed as one of the most promising approaches towards
-learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).
-
-In recent years supervised learning has yielded impressive results,
-however for these models to suceed huge amount of labelled data is needed.
-Unsupervised learning, that is learning from unlabelled data, has been expressed
-as one hurdles toward general articial intelligence (citation needed).
-Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
-
-In generative models the approach is different in a way that
-we search for a good internal representation of the data.
-This resembles the way humans learn about the world where we incrementally build
-a model of the world.
-
-%----------------------------------
-% Overview of Directed Generative Models
-%----------------------------------
-\section{Overview of Directed Generative Models}
-\label{sec:overview}
-
-\paragraph{Variational Autoencoders\cite{vae:2013}}
-\label{par:overview_vae}
-Autoencoders comprise of two components, the encoder and decoder
-where the encoder takes input data and translates it into a latent structure while
-the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
-
-In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
-% intractable -> use neural network, which we can train
-% --> no closed form possible
-We use the KL-Divergence to train the network.
-
-% TODO read chapter 19, approximate inference
-In VAEs however, this
-
-\paragraph{Generative Adversarial Networks}
-\label{par:overview_gan}
-GANs are a game theoretic approach to generative modelling
-in which two networks compete against each other.
-The discriminator $D$ is trying to distinguish the generated samples
-by the generator $G$ from real world data.
-The goal for $G$ however is to produce data as realistic as possible.\\
-During the early phase of training, $G$ produces random noise which
-is easy to differentiate from groundtruth data.
-Both networks are trained in parallel, where both networks try to get better
-at their respective objective.
-In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
-case the discriminator will only be able to guess correctly with a 50\% chance.
-This plateau is called Nash equilibrium (citation needed).
-
-%----------------------------------
-% Sigmoid Belief Networks
-%----------------------------------
-\section{Sigmoid Belief Networks}
-\label{sec:sbn}
-
-%----------------------------------
-% Variational Autoencoders
-%----------------------------------
-\section{Variational Autoencoders}
-\label{sec:vae}
-Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
-To define VAEs, we first declare the observed variables $x$, as well as latent variables $z$.
-VAEs are directed graphical models where we wish to infer \emph{good} low-dimensional latent variables from the input data.
-In this setup we define $p(x,z)$ as the joint probability distribution over both input and latent variables.
-Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
-
-In this setup we have observed variables $x$ from which we would like to infer latent variables $z$
-
-%Variational Autoencoders (VAE) have been a popular choice for unsupervised learning of complicated distributions (citation needed) and generative modelling.
-In order to generate data from unknown and mostly intractable distributions, we need approximations.
-There are basically two approaches for sampling from these distributions,
-first there are approximate samples (MCMC, gibbs sampling, etc) which try to directly approximate $p(x)$.
-Variational Autoencoders instead try to match an easier to compute distribution $q(x)$ to $p(x)$.
-By using this approach, VAEs are computationally less intensive (citation?) but have the drawback of being more restricted in their modelling approach.
-Practically, this means that with more computation MCMC methods approach $p(x)$, while there is no such guarantee for variational methods (--> not for all).
-
-VAE can be learned with just backpropagation (paper,TODO), but they differ from denoising and sparse autoencoders due to the different loss function.
-
-%VAEs are built on top of neural networks and are designed in a way to allow training with gradient-based methods.
-%Learning and inference are reasonable efficient and relatively easy to implement and show decent results, but have been overshadowed by more recent adversarial approaches (citation needed!!,see \ref{sec:gan}).
-
-\subsection{Architecture}
-\label{sub:vae_architecture}
-VAE just like other autoencoders encode the input data into a smaller latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
-But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.
-This allows to sample from this distribution and generate output which will look similar to the data on which the VAE has been trained.
-\ref{fig:vae_architecture}.
-
-\subsubsection{Reparameterization Trick}
-To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
-Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
-Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
-\begin{equation}
-  \label{eq:rep_trick}
-  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
-\end{equation}
-This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
-
-The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
-In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
-
-Variational Autoencoders achieve this by using a measurement of inequality between $q(x)$ and $p(x)$ called Kullback-Leibler Divergence (KL divergence).
-
-%VAEs are unlike \emph{traditional} auto-encoders (sparse autoencoders or denoising autoencoders for example) because they have a different architecture
-%VAEs resembles traditional autoencoders
-%\begin{equation}
-%P(x) = \int P(x|z;\theta)P(z) dz
-  %\label{eq:max_likelihood}
-%\end{equation}
-
-We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
-The prior distribution $p_\theta(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
-The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2013} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
-Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
-\begin{equation}
-  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p_\theta(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
-\end{equation}
-\paragraph{Derivation of lower bound}
-We want to minimize the KL divergence from $P(z|x)$ to $Q(z)$ for some arbitrary $Q$.
-  \begin{align*}
-    \label{eq:der_lower_bound}
-    \mathcal{D}[Q(z)||P(z|x)] &= \mathbb{E}_{z~q_\theta}(\log Q(z) - \log P(z|x))
-    &= \mathbb{E}_{z ~ q_\theta}[]
-  \end{align*}
-Where $D_{KL}(q_\theta(z|x) || p_\theta(z))$ is defined as the Kullback-Leibler divergence, which can be intuitively thought of as a measurement for the distance between two probabilistic distributions (even though it lacks the symmetric property).
-
-\subsection{Learning}
-\label{sub:vae_learning}
-In practice, the usual choice for $q_\theta(z|x)$ is the multivariate gaussian distribution.
-$$
-q_\theta(z|x) = \mathcal{N}(z|\mu(X;\vartheta), \Sigma(X;\vartheta))
-$$
-Because we assume that $P(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
-\begin{align*}
-  \mathcal{D}_{\mathrm{KL}}\big[Q(z) || P(z|x)\big] &= \mathbb{E}\big[\log \frac{p}{q}\big]\\
-  \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]
-  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)
-\end{align*}
-
-\subsection{Inference}
-\label{sub:vae_inference}
-
-\subsection{Performance}
-\label{sub:vae_performance}
-
-\subsection{Extensions}
-\label{sub:vae_extensions}
-
-\subsubsection{Deep Recurrent Attention Writer}
-\label{ssub:vae_deep_recurrent_attention_writer}
-
-\subsubsection{Importance Weighted Autoencoder}
-\label{ssub:vae_importance_weighted_autoencoder}
-
-\subsubsection{Conditional VAE}
-\label{ssub:vae_conditional_vae}
-
-%----------------------------------
-% Generative Adversarial Netoworks
-%----------------------------------
-\section{Generative Adversarial Networks}
-\label{sec:gan}
-
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
-Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
-% --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-
-\subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
-$$
-\mathcal{C}_g = \log(1 - D(G(z)))
-$$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
-
-Note that this loss function is fully differentiable which means it can be trained with backpropagation.
-
-\subsubsection{Discriminator Network}
-The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
-$$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
-
-\subsection{Training}
-\label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
-$$
-\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
-
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
-
-\subsection{Stability and Performance}
-\label{sub:gan_stability}
-
-\subsubsection{Freeze Learning}
-\label{ssub:gan_freeze_learning}
-
-\subsubsection{Feature Matching}
-\label{ssub:gan_feature_matching}
-
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
-
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
-\subsection{Extensions}
-\label{sub:gan_extensions}
-
-\subsubsection{LAPGAN}
-\label{ssub:lapgan}
-
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
-
-%----------------------------------
-% More Generative Directed Models
-%----------------------------------
-\section{More Generative Directed Models}
-\label{sec:more}
-
-\subsection{Generative Moment Matching Networks}
-\label{sub:more_mmn}
-
-\subsection{Auto-Regressive Networks}
-\label{sub:more_arn}
-
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
-
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
-
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2016-11-30 paper/paper.tex
remove new page for now
fix whitespace issues
add tikz package
added matrikelnr
moved abstract to file

@@
 \usepackage{hyperref}
-\usepackage[utf8]{inputenc}
+\usepackage[utf8x]{inputenc}
+\usepackage{tikz}
 \usepackage{algpseudocode}
@@
     \textit{André Merboldt} \\[6ex]
+    \textit{346703} \\[6ex]
     \today
@@
     \Large Supervisor: Tobias Menne
-
 \end{center}
@@
 \listoffigures
-\newpage
-\pagestyle{empty}
-\ 
+%\newpage
+%\pagestyle{empty}
+%\
 \newpage
@@
 %----------------------------------
-\section{Abstract}
-\label{sec:abstract}
-In this seminar paper we provide an overview and introduction to directed generative models,
-which are able to produce data following a distribution (...?).
-% more info what generative models try to achieve
-After we have introduced several architectures, we will explore sigmoid belief networks (SBN)
-as an early generative model.\\
-
-Then we will turn to more recent proposals, in particular variational autoencoders (VAE) and
-generative adversarial networks (GAN).
-Variants of GANs have been shown to produce images which are able to fool a human discriminator 40\% of the time (citation needed, LAPGAN? or DCGAN?).
-Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.
+\input{chapters/abstract}
 

### 2016-11-30 paper/chapters/vae.tex
VAE to file, added derivation for loss function

--- 2016-11-30 paper/paper.tex
@@
-\documentclass[twoside,11pt,a4paper]{article}
-
-% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx,curves,float,rotating}
-
-\usepackage{amsmath, amssymb, latexsym}  % math stuff
-\usepackage{amsfonts}
-\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[american]{babel}                     % otherwise use british or american
-\usepackage{theorem}                            % instead of \usepackage{amsthm}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage[utf8x]{inputenc}
-\usepackage{tikz}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-
-\newcommand*\Let[2]{\State #1 $\gets$ #2}
-
-% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{xspace}                             % context sensitive space after macros
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
-\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
-\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
-\def\etc{{etc}\onedot} \def\vs{{vs}\onedot}
-\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
-\def\etal{{et al}\onedot}
-\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
-\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%	Macros fuer neue Umgebungen
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-\newlength{\textwd}
-\newlength{\oddsidemargintmp}
-\newlength{\evensidemargintmp}
-\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
-\newlength{\textht}
-\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
-\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-
-\newenvironment{deflist}[1][\quad]%
-{  \begin{list}{}{%
-      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
-      \settowidth{\labelwidth}{\textbf{#1}}%
-      \setlength{\leftmargin}{\labelwidth}
-      \addtolength{\leftmargin}{\labelsep}}}
-{  \end{list}}
-
-\newenvironment{Quote}% Definition of Quote
-{  \begin{list}{}{%
-      \setlength{\rightmargin}{0pt}}
-      \item[]\ignorespaces}
-{\unskip\end{list}}
-
-\theoremstyle{break}
-\theorembodyfont{\itshape}
-\theoremheaderfont{\scshape}
-
-\newtheorem{Cor}{Corollary}
-\newtheorem{Def}{Definition}
-%\newtheorem{Def}[Cor]{Definition}
-
-\newcolumntype{.}{D{.}{.}{-1}}
-
-\pagestyle{headings}
-\textwidth 15cm
-\textheight 23cm
-\oddsidemargin 1cm
-\evensidemargin 0cm
-%\parindent 0mm
-
-\begin{document}
-
-\pagestyle{empty}
-
-\begin{center}
-
-    RWTH Aachen University\\
-    Chair of Computer Science 6 \\
-    Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
-
-    \LARGE
-    \textbf{Deep Directed Generative Models} \\[6ex]
-    \textit{André Merboldt} \\[6ex]
-    \textit{346703} \\[6ex]
-    \today
-
-    \vfill
-    \Large Supervisor: Tobias Menne
-\end{center}
-
-% blank page
-\newpage
-\
-\newpage
-
-% contents
-\pagestyle{headings}
-\tableofcontents
-\listoftables
-\listoffigures
-%\newpage
-%\pagestyle{empty}
-%\
-\newpage
-\pagestyle{headings}
-
-%----------------------------------
-% ABSTRACT
-%----------------------------------
-\input{chapters/abstract}
-
-%----------------------------------
-% Motivation
-%----------------------------------
-\section{Motivation}
-\label{sec:motivation}
-Generative models have been proposed as one of the most promising approaches towards
-learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).
-
-In recent years supervised learning has yielded impressive results,
-however for these models to suceed huge amount of labelled data is needed.
-Unsupervised learning, that is learning from unlabelled data, has been expressed
-as one hurdles toward general articial intelligence (citation needed).
-Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
-
-In generative models the approach is different in a way that
-we search for a good internal representation of the data.
-This resembles the way humans learn about the world where we incrementally build
-a model of the world.
-
-%----------------------------------
-% Overview of Directed Generative Models
-%----------------------------------
-\section{Overview of Directed Generative Models}
-\label{sec:overview}
-
-\paragraph{Variational Autoencoders\cite{vae:2013}}
-\label{par:overview_vae}
-Variational Autoencoders (VAE) use two networks called encoder and decoder where the encoder transform the given data into a latent space and the decoder translates this latent space back into the data space.
-The input and output space are usually high-dimensional data spaces, for example images while the latent space is low-dimensional in comparison.
-Like other autoencoders, VAEs try to reconstruct the input data using the latent variables as closely as possible.
-But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
-This encouragement is done using a variational lower bound on the distribution output by the encoder network.
-
-%where the encoder takes input data and translates it into a latent structure while
-%the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
-
-%In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
-% intractable -> use neural network, which we can train
-% --> no closed form possible
-%We use the KL-Divergence to train the network.
-
-% TODO read chapter 19, approximate inference
-%In VAEs however, this
-
-\paragraph{Generative Adversarial Networks}
-\label{par:overview_gan}
-GANs are a game theoretic approach to generative modelling
-in which two networks compete against each other.
-The discriminator $D$ is trying to distinguish the generated samples
-by the generator $G$ from real world data.
-The goal for $G$ however is to produce data as realistic as possible.\\
-During the early phase of training, $G$ produces random noise which
-is easy to differentiate from groundtruth data.
-Both networks are trained in parallel, where both networks try to get better
-at their respective objective.
-In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
-case the discriminator will only be able to guess correctly with a 50\% chance.
-This plateau is called Nash equilibrium (citation needed).
-
-%----------------------------------
-% Sigmoid Belief Networks
-%----------------------------------
-\section{Sigmoid Belief Networks}
-\label{sec:sbn}
-Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
-In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
-$$
-p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
-$$
-
-While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
-% show why?
-
-There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
-
-%----------------------------------
-% Variational Autoencoders
-%----------------------------------
 \section{Variational Autoencoders}
@@
 Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
-To define VAEs, we first declare the observed variables $x$, as well as latent variables $z$.
-VAEs are directed graphical models where we wish to infer \emph{good} low-dimensional latent variables from the input data.
-In this setup we define $p(x,z)$ as the joint probability distribution over both input and latent variables.
-Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
+% maybe extra section/sub?
+We assume that the high-dimensional input data $x$ can be explained using low-dimensional latent variables $z$.\\
+While this might be surprising at first, most high-dimensional data we encounter can be explained using far fewer dimensions.
+Let us take a look at different examples:\\
+\begin{itemize}
+  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler.
+For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
+There are a few more dimensions, stroke width, cursive, position, size.
+But still far lower than 784.
+We will take a look at this example later on in \ref{sec:vae}.
+  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
+  \item \textbf{...}
+\end{itemize}
+Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling (?).
+%The VAE framework provides a directed model which aims to infer low-dimensional latent variables $z$ from high-dimensional input data $x$ and reconstruct the input.
+Formally, $p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.
+Inferring the posterior distribution $p(z|x)$ is particularly interesting, because it means to enable learning parameters for good latent space representation.
+$p(z|x)$ can be expanded using bayes rule to $\frac{p(x|z) p(z)}{p(x)}$.
+Computing the nominator is straightforward, $p(z)$ is the probability distribution we chose for the latent space, oftenmost simply a gaussian.
+$p(x|z)$ is easy to compute as well, ...? (forward pass?).\\
+Meanwhile is the denominator difficult to evaluate, because it requires to consider all possible input combinations.
+What we do instead in the variational autoencoders is to derive a lower-bound on $p(z|x)$ using a auxiliary distribution called $q_\theta(z|x)$.
+The subscript $\theta$ indicates that the distribution is parametrized by an variational term, so that for one $\theta$, $q_\theta(z|x) \approxeq p(z|x)$ holds ($q(z|x)$ doesn't need to depend on $x$).
+%Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
 
-In this setup we have observed variables $x$ from which we would like to infer latent variables $z$
+%In this setup we have observed variables $x$ from which we would like to infer latent variables $z$
 
 %Variational Autoencoders (VAE) have been a popular choice for unsupervised learning of complicated distributions (citation needed) and generative modelling.
-In order to generate data from unknown and mostly intractable distributions, we need approximations.
-There are basically two approaches for sampling from these distributions,
-first there are approximate samples (MCMC, gibbs sampling, etc) which try to directly approximate $p(x)$.
-Variational Autoencoders instead try to match an easier to compute distribution $q(x)$ to $p(x)$.
-By using this approach, VAEs are computationally less intensive (citation?) but have the drawback of being more restricted in their modelling approach.
-Practically, this means that with more computation MCMC methods approach $p(x)$, while there is no such guarantee for variational methods (--> not for all).
+%In order to generate data from unknown and mostly intractable distributions, we need approximations.
+%There are basically two approaches for sampling from these distributions,
+%first there are approximate samples (MCMC, gibbs sampling, etc) which try to directly approximate $p(x)$.
+%Variational Autoencoders instead try to match an easier to compute distribution $q(x)$ to $p(x)$.
+%By using this approach, VAEs are computationally less intensive (citation?) but have the drawback of being more restricted in their modelling approach.
+%Practically, this means that with more computation MCMC methods approach $p(x)$, while there is no such guarantee for variational methods (--> not for all).
 
-VAE can be learned with just backpropagation (paper,TODO), but they differ from denoising and sparse autoencoders due to the different loss function.
+%VAE can be learned with just backpropagation (paper,TODO), but they differ from denoising and sparse autoencoders due to the different loss function.
 
@@
 \label{sub:vae_architecture}
+
+\begin{figure}[htb]
+\centering
+\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
+  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
+\end{figure}
 VAE just like other autoencoders encode the input data into a smaller latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
@@
 
+\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
+$\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
+The KL divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
+$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the VAE to approximate the true posterior distribution.
+
 \subsection{Loss function}
@@
 
-\paragraph{Reconstruction error}
-$\mathcal{C}_{\mathrm{rec}}$ measures how well the decoder reconstructs the input data using the latent space $z$.
-This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\varphi(z | x_i)$.
-$\mathcal{C}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+\paragraph{Derivation of lower bound}
+\input{chapters/vae/derivation_loss}
 
-\paragraph{KL divergence (regularization)}
-The second term in $\mathcal{C}$ is the KL divergence between the approximated distribution $q_\varphi(z|x)$ and $p(z)$.
-This serves as the regularizer in the VAE as it encourages $q_\varphi(z|x)$ to match $p(z)$.\\
+\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative Kullback-Leibler (KL) divergence between the learned variational distribution $q_\theta(z|x)$ and the latent space prior distribution $p(z)$.
+
 % move to architecture:
@@
 
-For two probability distributions $P$ and $Q$, the KL divergence $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can informally be described as the amount of information which is lost when using $Q$ to represent $P$.
-The KL divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
-$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the VAE to let $q_\varphi(z|x)$ approach $p(z)$.
+%For two probability distributions $P$ and $Q$, the KL divergence $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can informally be described as the amount of information which is lost when using $Q$ to represent $P$.
+
+\paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
+This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\varphi(z | x_i)$.
+$\mathcal{C}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
 
@@
 
-Variational Autoencoders achieve this by using a measurement of inequality between $q(x)$ and $p(x)$ called Kullback-Leibler Divergence (KL divergence).
-
-%VAEs are unlike \emph{traditional} auto-encoders (sparse autoencoders or denoising autoencoders for example) because they have a different architecture
-%VAEs resembles traditional autoencoders
-%\begin{equation}
-%P(x) = \int P(x|z;\theta)P(z) dz
-  %\label{eq:max_likelihood}
-%\end{equation}
-
 We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
@@
 \end{equation}
-\paragraph{Derivation of lower bound}
-We want to minimize the KL divergence from $P(z|x)$ to $Q(z)$ for some arbitrary $Q$.
-  \begin{align*}
-    \label{eq:der_lower_bound}
-    \mathcal{D}[Q(z)||P(z|x)] &= \mathbb{E}_{z~q_\theta}(\log Q(z) - \log P(z|x))
-    &= \mathbb{E}_{z ~ q_\theta}[]
-  \end{align*}
-Where $D_{KL}(q_\theta(z|x) || p_\theta(z))$ is defined as the Kullback-Leibler divergence, which can be intuitively thought of as a measurement for the distance between two probabilistic distributions (even though it lacks the symmetric property).
 
@@
 \label{sub:vae_inference}
+\begin{figure}[htb]
+\centering
+%\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
+\includegraphics{media/manifold.pdf}
+  \caption{MNIST 2-dimensional manifold}\label{fig:vae_manifold}
+\end{figure}
 
@@
 \label{ssub:vae_conditional_vae}
-
-%----------------------------------
-% Generative Adversarial Netoworks
-%----------------------------------
-\section{Generative Adversarial Networks}
-\label{sec:gan}
-
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
-Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
-% --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-
-\subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
-$$
-\mathcal{C}_g = \log(1 - D(G(z)))
-$$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
-
-Note that this loss function is fully differentiable which means it can be trained with backpropagation.
-
-\subsubsection{Discriminator Network}
-The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
-$$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
-
-\subsection{Training}
-\label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
-$$
-\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
-
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
-
-\subsection{Stability and Performance}
-\label{sub:gan_stability}
-
-\subsubsection{Freeze Learning}
-\label{ssub:gan_freeze_learning}
-
-\subsubsection{Feature Matching}
-\label{ssub:gan_feature_matching}
-
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
-
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
-\subsection{Extensions}
-\label{sub:gan_extensions}
-
-\subsubsection{LAPGAN}
-\label{ssub:lapgan}
-
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
-
-%----------------------------------
-% More Generative Directed Models
-%----------------------------------
-\section{More Generative Directed Models}
-\label{sec:more}
-
-\subsection{Generative Moment Matching Networks}
-\label{sub:more_mmn}
-
-\subsection{Auto-Regressive Networks}
-\label{sub:more_arn}
-
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
-
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
-
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2016-11-30 paper/paper.tex
VAE to file, added derivation for loss function

--- 2016-11-29 paper/paper.tex
@@
 %----------------------------------
-\section{Motivation}
-\label{sec:motivation}
-Generative models have been proposed as one of the most promising approaches towards
-learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).
-
-In recent years supervised learning has yielded impressive results,
-however for these models to suceed huge amount of labelled data is needed.
-Unsupervised learning, that is learning from unlabelled data, has been expressed
-as one hurdles toward general articial intelligence (citation needed).
-Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
-
-In generative models the approach is different in a way that
-we search for a good internal representation of the data.
-This resembles the way humans learn about the world where we incrementally build
-a model of the world.
+\input{chapters/introduction}
 
@@
 %----------------------------------
-\section{Overview of Directed Generative Models}
-\label{sec:overview}
-
-\paragraph{Variational Autoencoders\cite{vae:2013}}
-\label{par:overview_vae}
-Autoencoders comprise of two components, the encoder and decoder
-where the encoder takes input data and translates it into a latent structure while
-the decoder takes the latent structure and produces data which is as close as possible to the initial input data.
-
-In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
-% intractable -> use neural network, which we can train
-% --> no closed form possible
-We use the KL-Divergence to train the network.
-
-% TODO read chapter 19, approximate inference
-In VAEs however, this
-
-\paragraph{Generative Adversarial Networks}
-\label{par:overview_gan}
-GANs are a game theoretic approach to generative modelling
-in which two networks compete against each other.
-The discriminator $D$ is trying to distinguish the generated samples
-by the generator $G$ from real world data.
-The goal for $G$ however is to produce data as realistic as possible.\\
-During the early phase of training, $G$ produces random noise which
-is easy to differentiate from groundtruth data.
-Both networks are trained in parallel, where both networks try to get better
-at their respective objective.
-In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
-case the discriminator will only be able to guess correctly with a 50\% chance.
-This plateau is called Nash equilibrium (citation needed).
+\input{chapters/overview}
 
@@
 %----------------------------------
-\section{Sigmoid Belief Networks}
-\label{sec:sbn}
-Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
-In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
-$$
-p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
-$$
-
-While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
-% show why?
-
-There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
+\input{chapters/sigmoid_belief_nets}
 
@@
 %----------------------------------
-\section{Variational Autoencoders}
-\label{sec:vae}
-Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
-To define VAEs, we first declare the observed variables $x$, as well as latent variables $z$.
-VAEs are directed graphical models where we wish to infer \emph{good} low-dimensional latent variables from the input data.
-In this setup we define $p(x,z)$ as the joint probability distribution over both input and latent variables.
-Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
-
-In this setup we have observed variables $x$ from which we would like to infer latent variables $z$
-
-%Variational Autoencoders (VAE) have been a popular choice for unsupervised learning of complicated distributions (citation needed) and generative modelling.
-In order to generate data from unknown and mostly intractable distributions, we need approximations.
-There are basically two approaches for sampling from these distributions,
-first there are approximate samples (MCMC, gibbs sampling, etc) which try to directly approximate $p(x)$.
-Variational Autoencoders instead try to match an easier to compute distribution $q(x)$ to $p(x)$.
-By using this approach, VAEs are computationally less intensive (citation?) but have the drawback of being more restricted in their modelling approach.
-Practically, this means that with more computation MCMC methods approach $p(x)$, while there is no such guarantee for variational methods (--> not for all).
-
-VAE can be learned with just backpropagation (paper,TODO), but they differ from denoising and sparse autoencoders due to the different loss function.
-
-%VAEs are built on top of neural networks and are designed in a way to allow training with gradient-based methods.
-%Learning and inference are reasonable efficient and relatively easy to implement and show decent results, but have been overshadowed by more recent adversarial approaches (citation needed!!,see \ref{sec:gan}).
-
-\subsection{Architecture}
-\label{sub:vae_architecture}
-VAE just like other autoencoders encode the input data into a smaller latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
-But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.
-This allows to sample from this distribution and generate output which will look similar to the data on which the VAE has been trained.
-\ref{fig:vae_architecture}.
-
-\subsubsection{Reparameterization Trick}
-To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
-Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
-Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
-\begin{equation}
-  \label{eq:rep_trick}
-  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
-\end{equation}
-This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
-
-The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
-In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
-
-Variational Autoencoders achieve this by using a measurement of inequality between $q(x)$ and $p(x)$ called Kullback-Leibler Divergence (KL divergence).
-
-%VAEs are unlike \emph{traditional} auto-encoders (sparse autoencoders or denoising autoencoders for example) because they have a different architecture
-%VAEs resembles traditional autoencoders
-%\begin{equation}
-%P(x) = \int P(x|z;\theta)P(z) dz
-  %\label{eq:max_likelihood}
-%\end{equation}
-
-We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
-The prior distribution $p_\theta(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
-The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2013} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
-Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
-\begin{equation}
-  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p_\theta(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
-\end{equation}
-\paragraph{Derivation of lower bound}
-We want to minimize the KL divergence from $P(z|x)$ to $Q(z)$ for some arbitrary $Q$.
-  \begin{align*}
-    \label{eq:der_lower_bound}
-    \mathcal{D}[Q(z)||P(z|x)] &= \mathbb{E}_{z~q_\theta}(\log Q(z) - \log P(z|x))
-    &= \mathbb{E}_{z ~ q_\theta}[]
-  \end{align*}
-Where $D_{KL}(q_\theta(z|x) || p_\theta(z))$ is defined as the Kullback-Leibler divergence, which can be intuitively thought of as a measurement for the distance between two probabilistic distributions (even though it lacks the symmetric property).
-
-\subsection{Learning}
-\label{sub:vae_learning}
-In practice, the usual choice for $q_\theta(z|x)$ is the multivariate gaussian distribution.
-$$
-q_\theta(z|x) = \mathcal{N}(z|\mu(X;\vartheta), \Sigma(X;\vartheta))
-$$
-Because we assume that $P(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
-\begin{align*}
-  \mathcal{D}_{\mathrm{KL}}\big[Q(z) || P(z|x)\big] &= \mathbb{E}\big[\log \frac{p}{q}\big]\\
-  \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]
-  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)
-\end{align*}
-
-\subsection{Inference}
-\label{sub:vae_inference}
-
-\subsection{Performance}
-\label{sub:vae_performance}
-
-\subsection{Extensions}
-\label{sub:vae_extensions}
-
-\subsubsection{Deep Recurrent Attention Writer}
-\label{ssub:vae_deep_recurrent_attention_writer}
-
-\subsubsection{Importance Weighted Autoencoder}
-\label{ssub:vae_importance_weighted_autoencoder}
-
-\subsubsection{Conditional VAE}
-\label{ssub:vae_conditional_vae}
+\input{chapters/vae}
 

### 2016-11-30 paper/chapters/conclusion.tex
moved other chapters to files

--- 2016-11-30 paper/paper.tex
@@
-\documentclass[twoside,11pt,a4paper]{article}
-
-% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx,curves,float,rotating}
-
-\usepackage{amsmath, amssymb, latexsym}  % math stuff
-\usepackage{amsfonts}
-\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[american]{babel}                     % otherwise use british or american
-\usepackage{theorem}                            % instead of \usepackage{amsthm}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage[utf8x]{inputenc}
-\usepackage{tikz}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-
-\newcommand*\Let[2]{\State #1 $\gets$ #2}
-
-% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{xspace}                             % context sensitive space after macros
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
-\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
-\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
-\def\etc{{etc}\onedot} \def\vs{{vs}\onedot}
-\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
-\def\etal{{et al}\onedot}
-\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
-\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%	Macros fuer neue Umgebungen
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-\newlength{\textwd}
-\newlength{\oddsidemargintmp}
-\newlength{\evensidemargintmp}
-\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
-\newlength{\textht}
-\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
-\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-
-\newenvironment{deflist}[1][\quad]%
-{  \begin{list}{}{%
-      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
-      \settowidth{\labelwidth}{\textbf{#1}}%
-      \setlength{\leftmargin}{\labelwidth}
-      \addtolength{\leftmargin}{\labelsep}}}
-{  \end{list}}
-
-\newenvironment{Quote}% Definition of Quote
-{  \begin{list}{}{%
-      \setlength{\rightmargin}{0pt}}
-      \item[]\ignorespaces}
-{\unskip\end{list}}
-
-\theoremstyle{break}
-\theorembodyfont{\itshape}
-\theoremheaderfont{\scshape}
-
-\newtheorem{Cor}{Corollary}
-\newtheorem{Def}{Definition}
-%\newtheorem{Def}[Cor]{Definition}
-
-\newcolumntype{.}{D{.}{.}{-1}}
-
-\pagestyle{headings}
-\textwidth 15cm
-\textheight 23cm
-\oddsidemargin 1cm
-\evensidemargin 0cm
-%\parindent 0mm
-
-\begin{document}
-
-\pagestyle{empty}
-
-\begin{center}
-
-    RWTH Aachen University\\
-    Chair of Computer Science 6 \\
-    Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
-
-    \LARGE
-    \textbf{Deep Directed Generative Models} \\[6ex]
-    \textit{André Merboldt} \\[6ex]
-    \textit{346703} \\[6ex]
-    \today
-
-    \vfill
-    \Large Supervisor: Tobias Menne
-\end{center}
-
-% blank page
-\newpage
-\
-\newpage
-
-% contents
-\pagestyle{headings}
-\tableofcontents
-\listoftables
-\listoffigures
-%\newpage
-%\pagestyle{empty}
-%\
-\newpage
-\pagestyle{headings}
-
-%----------------------------------
-% ABSTRACT
-%----------------------------------
-\input{chapters/abstract}
-
-%----------------------------------
-% Motivation
-%----------------------------------
-\input{chapters/introduction}
-
-%----------------------------------
-% Overview of Directed Generative Models
-%----------------------------------
-\input{chapters/overview}
-
-%----------------------------------
-% Sigmoid Belief Networks
-%----------------------------------
-\input{chapters/sigmoid_belief_nets}
-
-%----------------------------------
-% Variational Autoencoders
-%----------------------------------
-\input{chapters/vae}
-
-%----------------------------------
-% Generative Adversarial Netoworks
-%----------------------------------
-\section{Generative Adversarial Networks}
-\label{sec:gan}
-
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
-Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
-% --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-
-\subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
-$$
-\mathcal{C}_g = \log(1 - D(G(z)))
-$$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
-
-Note that this loss function is fully differentiable which means it can be trained with backpropagation.
-
-\subsubsection{Discriminator Network}
-The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
-$$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
-
-\subsection{Training}
-\label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
-$$
-\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
-
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
-
-\subsection{Stability and Performance}
-\label{sub:gan_stability}
-
-\subsubsection{Freeze Learning}
-\label{ssub:gan_freeze_learning}
-
-\subsubsection{Feature Matching}
-\label{ssub:gan_feature_matching}
-
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
-
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
-\subsection{Extensions}
-\label{sub:gan_extensions}
-
-\subsubsection{LAPGAN}
-\label{ssub:lapgan}
-
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
-
-%----------------------------------
-% More Generative Directed Models
-%----------------------------------
-\section{More Generative Directed Models}
-\label{sec:more}
-
-\subsection{Generative Moment Matching Networks}
-\label{sub:more_mmn}
-
-\subsection{Auto-Regressive Networks}
-\label{sub:more_arn}
-
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
 \section{Conclusion}
@@
 Exciting times :)
-
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2016-11-30 paper/chapters/gan.tex
moved other chapters to files

--- 2016-11-30 paper/paper.tex
@@
-\documentclass[twoside,11pt,a4paper]{article}
-
-% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx,curves,float,rotating}
-
-\usepackage{amsmath, amssymb, latexsym}  % math stuff
-\usepackage{amsfonts}
-\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[american]{babel}                     % otherwise use british or american
-\usepackage{theorem}                            % instead of \usepackage{amsthm}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage[utf8x]{inputenc}
-\usepackage{tikz}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-
-\newcommand*\Let[2]{\State #1 $\gets$ #2}
-
-% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{xspace}                             % context sensitive space after macros
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
-\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
-\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
-\def\etc{{etc}\onedot} \def\vs{{vs}\onedot}
-\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
-\def\etal{{et al}\onedot}
-\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
-\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%	Macros fuer neue Umgebungen
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-\newlength{\textwd}
-\newlength{\oddsidemargintmp}
-\newlength{\evensidemargintmp}
-\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
-\newlength{\textht}
-\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
-\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-
-\newenvironment{deflist}[1][\quad]%
-{  \begin{list}{}{%
-      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
-      \settowidth{\labelwidth}{\textbf{#1}}%
-      \setlength{\leftmargin}{\labelwidth}
-      \addtolength{\leftmargin}{\labelsep}}}
-{  \end{list}}
-
-\newenvironment{Quote}% Definition of Quote
-{  \begin{list}{}{%
-      \setlength{\rightmargin}{0pt}}
-      \item[]\ignorespaces}
-{\unskip\end{list}}
-
-\theoremstyle{break}
-\theorembodyfont{\itshape}
-\theoremheaderfont{\scshape}
-
-\newtheorem{Cor}{Corollary}
-\newtheorem{Def}{Definition}
-%\newtheorem{Def}[Cor]{Definition}
-
-\newcolumntype{.}{D{.}{.}{-1}}
-
-\pagestyle{headings}
-\textwidth 15cm
-\textheight 23cm
-\oddsidemargin 1cm
-\evensidemargin 0cm
-%\parindent 0mm
-
-\begin{document}
-
-\pagestyle{empty}
-
-\begin{center}
-
-    RWTH Aachen University\\
-    Chair of Computer Science 6 \\
-    Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
-
-    \LARGE
-    \textbf{Deep Directed Generative Models} \\[6ex]
-    \textit{André Merboldt} \\[6ex]
-    \textit{346703} \\[6ex]
-    \today
-
-    \vfill
-    \Large Supervisor: Tobias Menne
-\end{center}
-
-% blank page
-\newpage
-\
-\newpage
-
-% contents
-\pagestyle{headings}
-\tableofcontents
-\listoftables
-\listoffigures
-%\newpage
-%\pagestyle{empty}
-%\
-\newpage
-\pagestyle{headings}
-
-%----------------------------------
-% ABSTRACT
-%----------------------------------
-\input{chapters/abstract}
-
-%----------------------------------
-% Motivation
-%----------------------------------
-\input{chapters/introduction}
-
-%----------------------------------
-% Overview of Directed Generative Models
-%----------------------------------
-\input{chapters/overview}
-
-%----------------------------------
-% Sigmoid Belief Networks
-%----------------------------------
-\input{chapters/sigmoid_belief_nets}
-
-%----------------------------------
-% Variational Autoencoders
-%----------------------------------
-\input{chapters/vae}
-
-%----------------------------------
-% Generative Adversarial Netoworks
-%----------------------------------
 \section{Generative Adversarial Networks}
@@
 \label{ssub:dcgan}
-
-%----------------------------------
-% More Generative Directed Models
-%----------------------------------
-\section{More Generative Directed Models}
-\label{sec:more}
-
-\subsection{Generative Moment Matching Networks}
-\label{sub:more_mmn}
-
-\subsection{Auto-Regressive Networks}
-\label{sub:more_arn}
-
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
-
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
-
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2016-11-30 paper/chapters/introduction.tex
moved other chapters to files

\section{Introduction}
\label{sec:motivation}
Generative models have been proposed as one of the most promising approaches towards
learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).

In recent years supervised learning has yielded impressive results,
however for these models to suceed huge amount of labelled data is needed.
Unsupervised learning, that is learning from unlabelled data, has been expressed
as one hurdles toward general articial intelligence (citation needed).
Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).

In generative models the approach is different in a way that
we search for a good internal representation of the data.
This resembles the way humans learn about the world where we incrementally build
a model of the world.

### 2016-11-30 paper/chapters/more.tex
moved other chapters to files

--- 2016-11-30 paper/paper.tex
@@
-\documentclass[twoside,11pt,a4paper]{article}
-
-% packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx,curves,float,rotating}
-
-\usepackage{amsmath, amssymb, latexsym}  % math stuff
-\usepackage{amsfonts}
-\usepackage{amsopn}                             % um mathe operatoren zu deklarieren
-\usepackage[american]{babel}                     % otherwise use british or american
-\usepackage{theorem}                            % instead of \usepackage{amsthm}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage[utf8x]{inputenc}
-\usepackage{tikz}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-
-\newcommand*\Let[2]{\State #1 $\gets$ #2}
-
-% @ environment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{xspace}                             % context sensitive space after macros
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{{e.g}\onedot} \def\Eg{{E.g}\onedot}
-\def\ie{{i.e}\onedot} \def\Ie{{I.e}\onedot}
-\def\cf{{c.f}\onedot} \def\Cf{{C.f}\onedot}
-\def\etc{{etc}\onedot} \def\vs{{vs}\onedot}
-\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
-\def\etal{{et al}\onedot}
-\def\zB{z.B\onedot} \def\ZB{Z.B\onedot}
-\def\dh{d.h\onedot} \def\Dh{D.h\onedot}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%	Macros fuer neue Umgebungen
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*{\Frac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-\newlength{\textwd}
-\newlength{\oddsidemargintmp}
-\newlength{\evensidemargintmp}
-\newcommand*{\hspaceof}[2]{\settowidth{\textwd}{#1}\mbox{\hspace{#2\textwd}}}
-\newlength{\textht}
-\newcommand*{\vspaceof}[3]{\settoheight{\textht}{#1}\mbox{\raisebox{#2\textht}{#3}}}
-\newcommand*{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-
-\newenvironment{deflist}[1][\quad]%
-{  \begin{list}{}{%
-      \renewcommand{\makelabel}[1]{\textbf{##1}\hfil}%
-      \settowidth{\labelwidth}{\textbf{#1}}%
-      \setlength{\leftmargin}{\labelwidth}
-      \addtolength{\leftmargin}{\labelsep}}}
-{  \end{list}}
-
-\newenvironment{Quote}% Definition of Quote
-{  \begin{list}{}{%
-      \setlength{\rightmargin}{0pt}}
-      \item[]\ignorespaces}
-{\unskip\end{list}}
-
-\theoremstyle{break}
-\theorembodyfont{\itshape}
-\theoremheaderfont{\scshape}
-
-\newtheorem{Cor}{Corollary}
-\newtheorem{Def}{Definition}
-%\newtheorem{Def}[Cor]{Definition}
-
-\newcolumntype{.}{D{.}{.}{-1}}
-
-\pagestyle{headings}
-\textwidth 15cm
-\textheight 23cm
-\oddsidemargin 1cm
-\evensidemargin 0cm
-%\parindent 0mm
-
-\begin{document}
-
-\pagestyle{empty}
-
-\begin{center}
-
-    RWTH Aachen University\\
-    Chair of Computer Science 6 \\
-    Prof. Dr.-Ing. Hermann Ney\\[6ex]
-    Selected Topics in Human Language Technology and Pattern Recognition WS 16/17\\[12ex]
-
-    \LARGE
-    \textbf{Deep Directed Generative Models} \\[6ex]
-    \textit{André Merboldt} \\[6ex]
-    \textit{346703} \\[6ex]
-    \today
-
-    \vfill
-    \Large Supervisor: Tobias Menne
-\end{center}
-
-% blank page
-\newpage
-\
-\newpage
-
-% contents
-\pagestyle{headings}
-\tableofcontents
-\listoftables
-\listoffigures
-%\newpage
-%\pagestyle{empty}
-%\
-\newpage
-\pagestyle{headings}
-
-%----------------------------------
-% ABSTRACT
-%----------------------------------
-\input{chapters/abstract}
-
-%----------------------------------
-% Motivation
-%----------------------------------
-\input{chapters/introduction}
-
-%----------------------------------
-% Overview of Directed Generative Models
-%----------------------------------
-\input{chapters/overview}
-
-%----------------------------------
-% Sigmoid Belief Networks
-%----------------------------------
-\input{chapters/sigmoid_belief_nets}
-
-%----------------------------------
-% Variational Autoencoders
-%----------------------------------
-\input{chapters/vae}
-
-%----------------------------------
-% Generative Adversarial Netoworks
-%----------------------------------
-\section{Generative Adversarial Networks}
-\label{sec:gan}
-
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
-Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
-% --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-
-\subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
-$$
-\mathcal{C}_g = \log(1 - D(G(z)))
-$$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
-
-Note that this loss function is fully differentiable which means it can be trained with backpropagation.
-
-\subsubsection{Discriminator Network}
-The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
-$$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
-
-\subsection{Training}
-\label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
-$$
-\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
-
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
-
-\subsection{Stability and Performance}
-\label{sub:gan_stability}
-
-\subsubsection{Freeze Learning}
-\label{ssub:gan_freeze_learning}
-
-\subsubsection{Feature Matching}
-\label{ssub:gan_feature_matching}
-
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
-
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
-\subsection{Extensions}
-\label{sub:gan_extensions}
-
-\subsubsection{LAPGAN}
-\label{ssub:lapgan}
-
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
-
-%----------------------------------
-% More Generative Directed Models
-%----------------------------------
 \section{More Generative Directed Models}
@@
 See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
-
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
-
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2016-11-30 paper/chapters/overview.tex
moved other chapters to files

\section{Overview of Directed Generative Models}
\label{sec:overview}

\paragraph{Variational Autoencoders\cite{vae:2013}}
\label{par:overview_vae}
Variational Autoencoders (VAE) use two networks called encoder and decoder where the encoder transform the given data into a latent space and the decoder translates this latent space back into the data space.
The input and output space are usually high-dimensional data spaces, for example images while the latent space is low-dimensional in comparison.
Like other autoencoders, VAEs try to reconstruct the input data using the latent variables as closely as possible.
But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
This encouragement is done using a variational lower bound on the distribution output by the encoder network.

%where the encoder takes input data and translates it into a latent structure while
%the decoder takes the latent structure and produces data which is as close as possible to the initial input data.

%In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
% intractable -> use neural network, which we can train
% --> no closed form possible
%We use the KL-Divergence to train the network.

% TODO read chapter 19, approximate inference
%In VAEs however, this

\paragraph{Generative Adversarial Networks}
\label{par:overview_gan}
GANs are a game theoretic approach to generative modelling
in which two networks compete against each other.
The discriminator $D$ is trying to distinguish the generated samples
by the generator $G$ from real world data.
The goal for $G$ however is to produce data as realistic as possible.\\
During the early phase of training, $G$ produces random noise which
is easy to differentiate from groundtruth data.
Both networks are trained in parallel, where both networks try to get better
at their respective objective.
In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
case the discriminator will only be able to guess correctly with a 50\% chance.
This plateau is called Nash equilibrium (citation needed).

### 2016-11-30 paper/chapters/sigmoid_belief_nets.tex
moved other chapters to files

\section{Sigmoid Belief Networks}
\label{sec:sbn}
Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
$$
p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
$$

While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
% show why?

There have been several attempts on solving the inference problem, using inference networks which have shown promising results.

### 2016-11-30 paper/paper.tex
moved other chapters to files

@@
 %----------------------------------
-\section{Generative Adversarial Networks}
-\label{sec:gan}
-
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
-Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
-% --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-
-\subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
-$$
-\mathcal{C}_g = \log(1 - D(G(z)))
-$$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
-
-Note that this loss function is fully differentiable which means it can be trained with backpropagation.
-
-\subsubsection{Discriminator Network}
-The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
-$$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
-
-\subsection{Training}
-\label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
-$$
-\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
-
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
-
-\subsection{Stability and Performance}
-\label{sub:gan_stability}
-
-\subsubsection{Freeze Learning}
-\label{ssub:gan_freeze_learning}
-
-\subsubsection{Feature Matching}
-\label{ssub:gan_feature_matching}
-
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
-
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
-\subsection{Extensions}
-\label{sub:gan_extensions}
-
-\subsubsection{LAPGAN}
-\label{ssub:lapgan}
-
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
+\input{chapters/gan}
 
@@
 %----------------------------------
-\section{More Generative Directed Models}
-\label{sec:more}
+\input{chapters/more}
 
-\subsection{Generative Moment Matching Networks}
-\label{sub:more_mmn}
-
-\subsection{Auto-Regressive Networks}
-\label{sub:more_arn}
-
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
-
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
+\input{chapters/conclusion}
 

### 2016-12-01 paper/chapters/vae.tex
add paragraph 'latent space' and 'formal setup'

@@
 % maybe extra section/sub?
+
+\paragraph{Latent Space}
 We assume that the high-dimensional input data $x$ can be explained using low-dimensional latent variables $z$.\\
@@
 \end{itemize}
+
 Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling (?).
 %The VAE framework provides a directed model which aims to infer low-dimensional latent variables $z$ from high-dimensional input data $x$ and reconstruct the input.
+\paragraph{Formal Setup}
 Formally, $p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.

### 2016-12-01 paper/chapters/vae.tex
latent space doesn't have to be smaller than input

@@
 \end{figure}
-VAE just like other autoencoders encode the input data into a smaller latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
+VAE just like other autoencoders encode the input data into a latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
 But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.

### 2016-12-01 paper/chapters/vae.tex
add relation to other AEs

@@
 \ref{fig:vae_architecture}.
+
+\paragraph{Relation to Auto-Encoders}
+The Variational autoencoder has the same overall architecture than other autoencoders, for example sparse or denoising autoencoders.
+Similar to other autoencoder frameworks VAE provides an encoder as well as a decoder model, but in contrast to other autoencoders both networks are probabilistic.
+This means that the model described by the VAE framework can be seen as a jointly trained probabilistic encoder and probabilistic decoder.
+%One of the main differences is that the VAE framework enforces a specific prior distribution on the latent space, most of the time simply an isotropic gaussian.
 

### 2016-12-01 paper/chapters/vae.tex
loss -> objective

@@
 
-\subsection{Loss function}
-The loss function of the VAE consists of two terms, the negative log-likelihood or the reconstruction error $\mathcal{C}_{\mathrm{rec}}$ and the second term $\mathcal{C}_{\mathrm{KL}}$ serves as the regularizer.
+\subsection{Objective}
+The objective function of the VAE consists of the reconstruction error and the regularizer.
+%$\mathcal{L}$ learns both the encoder as well as the decoder with their respective parameters $\phi$ and $\theta$.
+% probabilistic encoder $q_\phi(z|x)$ (produces z values from which x could've been generated)
+% probabilistic decoder $p_\theta(x|z)$
 We will discuss both terms in detail below.

### 2016-12-01 paper/chapters/vae.tex
C -> L

@@
 \paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
-This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\varphi(z | x_i)$.
-$\mathcal{C}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\phi(z | x_i)$.
+$\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
 

### 2016-12-01 paper/chapters/abstract.tex
better abstract (slightly)

@@
 \label{sec:abstract}
-In this seminar paper we provide an overview and introduction to directed generative models,
-which are able to produce data following a distribution (...?).
+Deep Learning has advanced at an incredible speed the last few years and has achieved human-level performance in some discrimination and classification tasks (object recognition, speech recognition).
+These supervised methods have the huge drawback of requiring a vast amount of labeled data which is often expensive or unfeasible to obtain.
+Learning hidden structure from unlabelled data known as unsupervised learning has had less of an impact recently.
+However, it can be argued that this learning is most similar to human learning, in particular inferring information from observations and building an appropriate model from it.
+We will discuss and explore some models for performing aforementioned inference as well as generating new data from these representations.
+Building reasonable representation of the world from unlabelled data is arguably one of the key aspects to machine intelligence.
+%However one of the key aspects to machine intelligence is learning a rich (?) model about the world and draw conclusions from it.
+%The above mentioned tasks require huge amount of labelled data to perform well, which is one of the restrictions of these supervised learning mechanisms.
+%Unsupervised learning has the potential to make use of the vast unlabelled data surrounding the world.
+%This data can be used to learn representations
+
+%In this seminar paper we provide an overview and introduction to directed generative models,
+%which are able to produce data following a distribution (...?).
 % more info what generative models try to achieve
-After we have introduced several architectures, we will explore sigmoid belief networks (SBN)
-as an early generative model.\\
+%After we have introduced several architectures, we will explore sigmoid belief networks (SBN)
+%as an early generative model.\\
 
-Then we will turn to more recent proposals, in particular variational autoencoders (VAE) and
-generative adversarial networks (GAN).
-Variants of GANs have been shown to produce images which are able to fool a human discriminator 40\% of the time (citation needed, LAPGAN? or DCGAN?).
-Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.
+%Then we will turn to more recent proposals, in particular variational autoencoders (VAE) and
+%generative adversarial networks (GAN).
+%Variants of GANs have been shown to produce images which are able to fool a human discriminator 40\% of the time (citation needed, LAPGAN? or DCGAN?).
+%Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.

### 2016-12-01 paper/chapters/vae.tex
objective +-> function

@@
 
-\subsection{Objective}
+\subsection{Objective Function}
 The objective function of the VAE consists of the reconstruction error and the regularizer.
@@
 % probabilistic decoder $p_\theta(x|z)$
-We will discuss both terms in detail below.
+We will discuss both terms in detail below following the derivation of the lower variational bound $\mathcal{L}$ and the rewritten objective function.
 

### 2016-12-01 paper/chapters/abstract.tex
minor grammatic fixes

@@
 Deep Learning has advanced at an incredible speed the last few years and has achieved human-level performance in some discrimination and classification tasks (object recognition, speech recognition).
-These supervised methods have the huge drawback of requiring a vast amount of labeled data which is often expensive or unfeasible to obtain.
+These supervised methods have the huge drawback of requiring a vast amount of labelled data which is often expensive or unfeasible to obtain.
 Learning hidden structure from unlabelled data known as unsupervised learning has had less of an impact recently.
-However, it can be argued that this learning is most similar to human learning, in particular inferring information from observations and building an appropriate model from it.
-We will discuss and explore some models for performing aforementioned inference as well as generating new data from these representations.
-Building reasonable representation of the world from unlabelled data is arguably one of the key aspects to machine intelligence.
+However, it can be argued that this method is most similar to human learning, in particular inferring information from observations and building an appropriate model from it.
+We will discuss and explore some models for performing this inference as well as generating new data from these representations.
+Building reasonable representation of the world from unlabelled data is arguably one of the key aspects of machine intelligence.
 %However one of the key aspects to machine intelligence is learning a rich (?) model about the world and draw conclusions from it.

### 2016-12-01 paper/chapters/overview.tex
add citation to deep latent gaussian
rewritten overview for VAE

@@
 
-\paragraph{Variational Autoencoders\cite{vae:2013}}
+\paragraph{Variational Autoencoders (VAE) \cite{vae:2013}\cite{rezende:2014}}
 \label{par:overview_vae}
-Variational Autoencoders (VAE) use two networks called encoder and decoder where the encoder transform the given data into a latent space and the decoder translates this latent space back into the data space.
-The input and output space are usually high-dimensional data spaces, for example images while the latent space is low-dimensional in comparison.
-Like other autoencoders, VAEs try to reconstruct the input data using the latent variables as closely as possible.
-But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
-This encouragement is done using a variational lower bound on the distribution output by the encoder network.
+Variational Autoencoders (VAE) use two networks which are trained jointly using approximated learned inference.
+The objective function is aimed at building a reasonable simple latent space approximating the intractable posterior distribution using variational inference.
+It consists of the recognition and inference network which can be understood in the autoencoder jargon as probabilistic encoder and decoder.
+Using variational bayesian methods the intractable posterior distribution can be approximated and learned using backpropagation and gradient-based methods.
+
+%But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
+%This encouragement is done using a variational lower bound on the distribution output by the encoder network.
 

### 2016-12-06 paper/chapters/introduction.tex
modify introduction

@@
 \section{Introduction}
-\label{sec:motivation}
-Generative models have been proposed as one of the most promising approaches towards
-learning representations of real world data by some of the leading researches (YannLeCun, OpenAI blogpost).
+\label{sec:introduction}
+Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
+More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
+This assumption is reasonable given the following examples:\\
 
-In recent years supervised learning has yielded impressive results,
-however for these models to suceed huge amount of labelled data is needed.
-Unsupervised learning, that is learning from unlabelled data, has been expressed
-as one hurdles toward general articial intelligence (citation needed).
-Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
+\begin{itemize}
+  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler.
+For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
+There are a few more dimensions, stroke width, cursive, position, size.
+But still far lower than 784.
+We will take a look at this example later on in \ref{sec:vae}.
+  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
+  \item \textbf{...}
+\end{itemize}
 
-In generative models the approach is different in a way that
-we search for a good internal representation of the data.
-This resembles the way humans learn about the world where we incrementally build
-a model of the world.
+After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two promising frameworks, the variational autoencoder (VAE) and the generative adversarial network (GAN)
+
+%Generative models have been proposed as one of the most promising approaches towards
+%learning representations of real world data by some of the leading researchers (YannLeCun, OpenAI blogpost).
+
+%In recent years supervised learning has yielded impressive results,
+%however for these models to succeed huge amount of labelled data is needed.
+%Unsupervised learning, that is learning from unlabelled data, has been expressed
+%as one hurdles toward general articial intelligence (citation needed).
+%Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
+
+%In generative models the approach is different in a way that
+%we search for a good internal representation of the data.
+%This resembles the way humans learn about the world where we incrementally build
+%a model of the world.

### 2016-12-06 paper/chapters/overview.tex
add citation to GAN paper
add 'and'

@@
 Variational Autoencoders (VAE) use two networks which are trained jointly using approximated learned inference.
-The objective function is aimed at building a reasonable simple latent space approximating the intractable posterior distribution using variational inference.
+The objective function is aimed at building a reasonable simple latent space and approximating the intractable posterior distribution using variational inference.
 It consists of the recognition and inference network which can be understood in the autoencoder jargon as probabilistic encoder and decoder.
@@
 
-\paragraph{Generative Adversarial Networks}
+\paragraph{Generative Adversarial Networks\cite{gan:2014}}
 \label{par:overview_gan}

### 2016-12-06 paper/chapters/sigmoid_belief_nets.tex
small fixes & added citations

@@
 \label{sec:sbn}
-Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
+Sigmoid Belief Networks (SBN) \cite{neal:1992} have been proposed in 1992 by R.
+M.
+Neal and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
 In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:

### 2016-12-06 paper/chapters/vae.tex
add IWAE paper
moved latent space stuff to introduction

@@
 Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
-% maybe extra section/sub?
+Deep latent models are based on the assumption that input data can be explained using non-linear transformations coming from simple distributions. \cite{rezende:2014} (check!?).
 
-\paragraph{Latent Space}
-We assume that the high-dimensional input data $x$ can be explained using low-dimensional latent variables $z$.\\
-While this might be surprising at first, most high-dimensional data we encounter can be explained using far fewer dimensions.
-Let us take a look at different examples:\\
-\begin{itemize}
-  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler.
-For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
-There are a few more dimensions, stroke width, cursive, position, size.
-But still far lower than 784.
-We will take a look at this example later on in \ref{sec:vae}.
-  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
-  \item \textbf{...}
-\end{itemize}
+Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling in practice(?).
 
-Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling (?).
-%The VAE framework provides a directed model which aims to infer low-dimensional latent variables $z$ from high-dimensional input data $x$ and reconstruct the input.
 \paragraph{Formal Setup}
@@
 
-\subsubsection{Importance Weighted Autoencoder}
+\subsubsection{Importance Weighted Autoencoder \cite{iwae:2015}}
 \label{ssub:vae_importance_weighted_autoencoder}

### 2016-12-06 paper/chapters/sigmoid_belief_nets.tex
make relation to belief nets clear

--- 2016-11-30 paper/chapters/sigmoid_belief_nets.tex
@@
 \label{sec:sbn}
-Sigmoid Belief Networks (SBN) have been proposed in 1990 by Neal (citation) and thus are one of the earliest generative models and also predating all other presented models presented in this paper.
+Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a a specific type of bayesian networks \cite{pearl:1985}.
+They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
 In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:

### 2016-12-06 paper/chapters/sigmoid_belief_nets.tex
added tikz picture for sbn node

@@
 They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
+
+These random variables are interpreted in a bayesian way where each one may be observed, hidden (or latent) or unknown representing different states of knowledge or belief \cite{definetti:1974}.
+
+\begin{figure}[htb]
+\centering
+\input{media/sbn_node.tikz}
+  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
+\end{figure}
 In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:

### 2016-12-06 paper/media/sbn_node.tikz.tex
added tikz picture for sbn node

%& /home/spotlight/.config/TikzEdtWForms/TikzEdtWForms/0.2.1.0/temp_header
%\begin{document}
%\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta,calc,fit,backgrounds}
\tikzstyle{c} = [circle, draw, fill=blue!20, node distance=3cm, text width=2.5em, text centered]
\tikzstyle{b} = [rectangle, draw, fill=blue!20, node distance=3cm, text width=2.5em, text centered]
\tikzstyle{l} = [draw, -latex',thick]  
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=3mm]}]
\begin{tikzpicture}

\node [c] (prior_1) {\Large $s_j$};
\node [c] (prior_2) at ([shift={(3,0)}] prior_1) {\Large $s_{j+1}$};
\node (prior_others) at ([shift={(1.5,0)}] prior_2) {\Large $\dots$};
\node [c] (prior_n) at ([shift={(1.5,0)}] prior_others) {\Large $s_{i-1}$};

\node (layer_n) at ([shift={(-2,0)}] prior_1) {\Large Layer $n$};
%\node (layer_n_1) at ([shift={(-2,5)}] prior_1) {\Large Layer $n-1$};
\begin{pgfonlayer}{background}
\node[
      fill=black!20,node distance=4cm,,inner sep = 1em,
      fit=(layer_n) (prior_1) (prior_2) (prior_n)] (layer_n_node){};

%TODO: set width according to layer_n_node
\node[
      fill=black!20] (layer_n_1) at ([shift={(0,2)}] layer_n_node){\Large Layer $n-1$};
\end{pgfonlayer}

\node [c] (state_i) at ([shift={(0,-3)}] prior_2) {\Large $s_i$};

\draw[arr] (prior_1) -- node [left]{\Large $W_{i,j}$} (state_i);
\draw[arr] (prior_2) -- (state_i);
\draw[arr] (prior_n) -- (state_i);

% maybe reduce to \draw?
\draw[-{latex}] ($(prior_1)+(-1.5,1.5)$) -- (prior_1);
\draw[-{latex}]($(prior_1)+(0,2)$) -- (prior_1);
\draw[-{latex}]($(prior_1)+(1.5,1.5)$) -- (prior_1);

\end{tikzpicture}

%\end{document}

### 2016-12-06 paper/chapters/sigmoid_belief_nets.tex
add SBN architecture and small fixes for sbn node

@@
 These random variables are interpreted in a bayesian way where each one may be observed, hidden (or latent) or unknown representing different states of knowledge or belief \cite{definetti:1974}.
-
 \begin{figure}[htb]
 \centering
-\input{media/sbn_node.tikz}
-  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
+\input{media/sbn_architecture}
+  \caption{SBN architecture}
+  \label{fig:sbn_arch}
+  \medskip
+  \small
+  Sigmoid Belief Network described as a acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
+  Arrows show stochastic dependency.
+The dashed arrow indicates that it is possible in the original definition (and in the formula below) to have inter-layer dependencies but they are usually not modelled to simplify the architecture.
+\end{figure}
+
+\subsection{Learning}
+\begin{figure}[htb]
+\centering
+\includestandalone[mode=buildnew]{media/sbn_node}
+  \caption{Sigmoid Belief Net  architecture}
+  \label{fig:sbn_node}
 \end{figure}
@@
 $$
-p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
+p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
 $$

### 2016-12-07 paper/chapters/gan.tex
added image for gan architecture (stolen)
additions to GAN stability/performance

@@
 \label{sub:gan_arch}
+
+% http://www.slideshare.net/xavigiro/deep-learning-for-computer-vision-generative-models-and-adversarial-training-upc-2016
+% TODO: recreate in tikz
+\begin{figure}[htb]
+\centering
+  \includegraphics[width=\linewidth]{media/deep-learning-for-computer-vision-generative-models-and-adversarial-training-upc-2016-5-638.jpg}
+  \caption{Conceptual idea of generative adversarial networks (source: Summer seminar UPC TelecomBCN)}\label{fig:gan_architecture}
+\end{figure}
 
@@
 \label{sub:gan_stability}
+In the practical case where both function approximators $G$ and $D$ are represented using a neural network,
+GANs have been shown to be difficult to train using gradient descent.
+These issues have been credited to the problem of finding the nash equilibrium in the minimax-game with continious, high-dimensional parameters whereas current gradient-based methods are [designed] to reach an minimum.
+Additionally to these hypothetical thoughts, \cite{gan_distinguish_crit:2014} have shown that these methods may fail to converge.
+Acknoledging these problems, Salimans et al \cite{improved_gan:2016} has studied the stability and performance of GAN in detail resulting in a number of practical methods improving sample generation and improved semi-supervised learning performance.
+We will take a look at a selection of these methods in the following, focussing on neural networks as approximators.
 
@@
 \label{ssub:gan_freeze_learning}
+One of the main issues prominent during early GAN research is the disparity between $G$ and $D$.
+Commonly during training one of these networks will outpace the other resulting in degraded gradients where the other player is unable to learn from.
+Freeze learning is one of the natural counter to this problem, just freezing either $G$ or $D$ until the other network has catched up (formulation!).
+In practice the generator is far slower to train than the discriminator in most cases (citation).
+
+\subsubsection{Label Smoothing}
+\label{ssub:gan_label_smoothing}
+Another basic idea to improve gradient flow is to replace the binary classification values $0/1$ with smoothed values $0.1/0.9$ or $\epsilon/1-\epsilon$.
+This guarantees to provide the generator with an informative gradient to improve learning performance
+Additionally, smoothed labels have been shown to reduce the attack surface of neural networks to adversarial examples \cite{adv_examples:2016}.
+% TODO replace 0/1 with e/1 − e ??
+% TODO one-sided label smoothing
 
@@
 \label{ssub:gan_historical_averaging}
-
-\subsubsection{Label Smoothing}
-\label{ssub:gan_label_smoothing}
 

### 2016-12-07 paper/chapters/gan.tex
added GAN architecture image

@@
 % TODO: recreate in tikz
+
 \begin{figure}[htb]
 \centering
-  \includegraphics[width=\linewidth]{media/deep-learning-for-computer-vision-generative-models-and-adversarial-training-upc-2016-5-638.jpg}
-  \caption{Conceptual idea of generative adversarial networks (source: Summer seminar UPC TelecomBCN)}\label{fig:gan_architecture}
+\includestandalone[width=\linewidth,mode=buildnew]{media/gan_conceptual}
+  \caption{Conceptual figure of generative adversarial network}
+  \label{fig:gan_a}
+  \medskip
+  \small
+  Figure is illustrating the information and gradient flow in the GAN framework.\\
+  \emph{data space} denotes the real-world data, which we aim to model.
+  Dashed arrows denote the gradient flow.
 \end{figure}
+
+%\begin{figure}[htb]
+%\centering
+%\includestandalone[mode=buildnew,width=\linewidth]{media/gan_conceptual}
+  %\caption{Conceptual idea of generative adversarial networks (source: Summer seminar UPC TelecomBCN)}\label{fig:gan_architecture}
+%\end{figure}
 

### 2016-12-07 paper/media/gan_conceptual.tex
added GAN architecture image

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgf}
%\pgfplotsset{compat=1.14}

\begin{document}
\usetikzlibrary{arrows.meta}
% why does the figure seems to have lots of whitespace to the left when shapes[.symbols] is included??
%\usetikzlibrary{shapes.symbols}
%\tikzstyle{c} = [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1,gb]
\tikzstyle{gb} = [draw,outer sep=7,inner sep=7,minimum size=57,line width=1, very thick, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\begin{tikzpicture}

% latent space
\node[circle,gb] (latent_space) {Latent space $z$};

% nodes
\node [gb] (generator) at ([shift={(5,0)}] latent_space) {Generator $G$};
\node [cloud] (data) at ([shift={(0,4)}] generator) {data space};
\node [gb] (discriminator) at ([shift={(5,2)}] generator) {Discriminator $D$};
  \node[gb,dashed] (loss) at ([shift={(5,0)}] discriminator) {Loss $\mathcal{L}$};

% flow
\draw[arr] (latent_space) -- (generator) node[midway,label=below:sample] {};
\draw[arr] (generator) -- (discriminator) node[midway,label=below:generate] {};
\draw[arr] (data) -- (discriminator) node[midway,pos=0.4,label=below:sample] {};
\draw[arr] (discriminator) -- (loss) node[midway,label=below:Real/Fake] {};

% gradient flow
\path[dashed,arr] (loss) edge[bend right] node [label=$\nabla \mathcal{L}$ with regard to $D$] {} (discriminator);
\path[dashed,arr] (loss) edge[bend left] node [label=below:$\nabla \mathcal{L}$ with regard to $G$] {} (generator);

\end{tikzpicture}
\end{document}

### 2016-12-07 paper/chapters/gan.tex
normalizing flows -> VAE, oooops!
small fixes not worth an own commit

@@
   \small
-  Figure is illustrating the information and gradient flow in the GAN framework.\\
-  \emph{data space} denotes the real-world data, which we aim to model.
+  The above figure is illustrating the information and gradient flow in the GAN framework.\\
+  \emph{data space} denotes the real-world data sampled from $p_{data}$, which we aim to model.
   Dashed arrows denote the gradient flow.
@@
 
-\subsubsection{Normalizing Flows}
-\label{ssub:gan_nf}
-
 \subsection{Application}

### 2016-12-07 paper/chapters/sigmoid_belief_nets.tex
small fixes not worth an own commit

@@
   Sigmoid Belief Network described as a acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
-  Arrows show stochastic dependency.
-The dashed arrow indicates that it is possible in the original definition (and in the formula below) to have inter-layer dependencies but they are usually not modelled to simplify the architecture.
+  Arrows show stochastic dependency. The dashed arrow indicates that it is possible in the original definition, even though most commonly the hidden variables are divided into layers.
 \end{figure}
@@
 \includestandalone[mode=buildnew]{media/sbn_node}
-  \caption{Sigmoid Belief Net  architecture}
+  \caption{SBN node}
   \label{fig:sbn_node}

### 2016-12-07 paper/chapters/gan.tex
added (short) BN, DCGAN and LAPGAN texts

@@
 
+\subsubsection{(Virtual) Batch normalization}
+\label{ssub:gan_batch_norm}
+Using batch normalization \cite{batch_norm:2015} helped stabilizing learning DCGAN \cite{dcgan:2015}.
+\cite{improved_gan:2016} introduced virtual batch normalization where each input $x$ is normalized on one reference batch chosen at the start of learning instead of the other inputs in the mini-batch.
+
 \subsubsection{Feature Matching}
@@
 GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
+
 \subsection{Extensions}
@@
 
-\subsubsection{LAPGAN}
+\subsubsection{DCGAN \cite{dcgan:2015}}
+\label{ssub:dcgan}
+Early implementations of the GAN framework have been implemented using fully-connected layers,
+which have proven to be difficult and slow to train on real-world data (citation, original convnet papers?).
+Using deep convolutional networks has helped reach human-level performance in object recognition and classification problems on images (citations!).
+As the discriminator in the GAN model implements a binary classification task and the generator performance is heavily dependent on the discriminator, using convnets to train the model should improve performance of image generation immensely.
+\emph{Deep convolutional generative adversarial network} (DCGAN) has applied convolutional networks to both generator and discriminator, resulting in quantitively evualuated better results.
+DCGAN uses no fully connected or pooling layers, but instead fractionally strided convolutions (often wrongfully called deconvolutions).
+
+\subsubsection{LAPGAN \cite{lapgan:2015}}
 \label{ssub:lapgan}
+LAPGAN is an extension of GAN applied to images using laplacian pyramids \cite{laplacian:1983}.
+Instead of using a single generator network, LAPGAN uses a series of deep convolutional networks as generators.
 
-\subsubsection{DCGAN}
-\label{ssub:dcgan}
+\paragraph{Training} is heavily modified from the regular GAN framework.
+In each iteration a training sample is drawn from the dataset and a downsamples version from it is computed.
+The discriminator receives now either the computed high-pass image (from the original image and the down-then-up sampled image) or an high-pass image from the generator.
+These steps are repeated for other steps, downscaling the image.
+In the last stage, a traditional generator and discriminator is used for the $8x8$ image.
+\paragraph{Sampling} uses a conditional GAN to produce high-resolution images through different stages of deep generator networks.
+
+\paragraph{Results}
+As GANs have no direct measurement of the quality of generations due to their lack of objective function, human evaluation of the authenticity of generated images is used for quantitive results.
+LAPGAN has pushed the state of the art significantly, with generated images which are high-resolution in comparison to earlier methods and arguebly high variety of generations.
+Furthermore, humans have evaluated the generated images for real images 40\% of the time.

### 2016-12-07 paper/chapters/vae/cvae.tex
added some text to cvae and iwae

Instead of randomly generated data, generation of output conditioned on some prior information is needed.
A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks

Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.

To support conditional distributions, the following relations have to be rewritten:
\begin{itemize}
\item \textbf{Sampling from data space}: $x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
\item \textbf{Probabilistic encoder}: $q_\theta(z|x,y)$ instead of just $q_\theta(z|x)$.\\
\item \textbf{Probabilistic decoder}: $p_\phi(x|y,z)$ instead of just $p_\phi(x|z)$.\\
\end{itemize}

### 2016-12-07 paper/chapters/vae/iwae.tex
added some text to cvae and iwae

Importance Weighted Autoencoders (IWAE) extends on the VAE framework using $k$-sample importance weighting of the log-likelihood.
The authors of the papers argue that using more samples will only tighten the lower variational bound.
$$
\log p(x) \geq \mathcal{L}_{k+1}(x) \geq \mathcal{L}_{k}(x)
$$

Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum:
$$
\lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
$$

Of course, using more samples will result in higher computational cost.

### 2016-12-08 code/Layer/Dense.py
add linear fc layer

@@ -22,7 +22,7 @@ class Dense():
 
     def __call__(self, x):
         """apply Dense layer
-        
+
         Args:
             x(tf.Tensor): input data
         Returns:
@@ -34,7 +34,7 @@ class Dense():
 
     def _init(self, x):
         """Initialize Dense layer with modified Xavier initialization for non-linear units.
-        See https://arxiv.org/pdf/1502.01852.pdf 
+        See https://arxiv.org/pdf/1502.01852.pdf
 
         Args:
             x(tf.Tensor): input tensor
@@ -61,3 +61,13 @@ def xavier_init(shape, dtype=dtypes.float32):
     w = tf.random_normal([shape[0], shape[1]], stddev = sd)
     return w
 
+# Linear
+def Linear(input_, output_size, scope=None, stddev=0.02, bias_start=0.0, with_w=False):
+    shape = input_.get_shape().as_list()
+
+    with tf.variable_scope(scope or "Linear"):
+        matrix = tf.get_variable("Matrix", [shape[1], output_size], tf.float32,
+                                 tf.random_normal_initializer(stddev=stddev))
+        bias = tf.get_variable("bias", [output_size],
+            initializer=tf.constant_initializer(bias_start))
+        return tf.matmul(input_, matrix) + bias

### 2016-12-08 paper/media/vae_conceptual.tex
added conceptual vae figure

\documentclass{standalone}
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,outer sep=3,inner sep=3,minimum size=10,line width=1, very thick, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\begin{tikzpicture}

\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0,0) {$x$};
\node[fill=black!20] (lbl_input) at (0,1) {Input};

% output / reconstruction
\node[gb] (output) at ($(input) + (5.5cm,0)$) {$\tilde x$};
\node[fill=black!20] (lbl_rec) at ($(output) + (0,1cm)$) {Reconstruction};

% latent space
\node[fill=black!20] (lbl_lat) at ($(input) + (2.7cm,1cm)$) {Latent space};

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  xtick=\empty,
  xlabel={$(\mu,\sigma)$},
  width=3cm,
  at={($(input) + (2cm,-0.5cm)$)}
  ]
  \addplot [very thick,cyan!50!black, domain=0:6] {gauss(3,1)};
\end{axis}

% input -> latent
\draw[arr] (input) -- node[label=above:$q_\phi(z|x)$] {}($(input) + (2cm,0)$) {};

% latent -> output (reconstruction)
\draw[arr] ($(input) + (3.5cm,0)$) -- node[label=above:$p_\theta(x|z)$] {}(output) {};

\end{scope}

\end{tikzpicture}
\end{document}

### 2016-12-12 code/Network/GAN.py
added tf+pt implementation for simple GAN

@@ -1,0 +1,177 @@
+#!/usr/bin/env python3
+from datetime import datetime
+import numpy as np
+import tensorflow as tf
+import prettytensor as pt
+import sys
+import os
+from numpy.random import random,uniform
+from progressbar import ETA, Bar, Percentage, ProgressBar
+
+
+class GAN(object):
+    """Generative adversarial networks implementation in TensorFlow/PrettyTensor
+
+    Just for educational purposes, generator + discriminator implemented using MLP.
+    """
+
+    HYPERPARAMS = {
+            "batch_size": 128,
+            "learning_rate": 1e-2,
+            "beta1": 0.5, # for AdamOptimizer
+            "z_dim": 1, # dimension of noise prior
+            "h_dim": 5, # dim of hidden layer in D and G
+            "k": 1, # updates of discriminator
+            "data_dim": 1, # dimensionality of generated data
+            "nonlin": tf.nn.relu, # TODO LeakyReLU
+            "squash": tf.nn.tanh,
+            }
+
+    def __init__(self,
+            hyper_params = {}, # update to HYPERPARAMS
+            models_dir = "models/",
+            ):
+        if not os.path.isdir(models_dir):
+            os.makedirs(models_dir)
+        self.models_dir = models_dir
+
+        tf.reset_default_graph()
+
+        # insert updated hyper parameters into class
+        self.__dict__.update(GAN.HYPERPARAMS, **hyper_params)
+
+        print('Building tensorflow graph')
+        self.sess = tf.Session()
+        self._build_graph()
+
+        self.constructed = datetime.now().strftime("%y%m%d_%H%M")
+        logs_path = os.path.join("logs", "run_"+self.constructed)
+        self.writer = tf.train.SummaryWriter(logs_path, self.sess.graph)
+        self._build_summaries()
+
+        self.sess.run(tf.initialize_all_variables())
+
+    def _build_summaries(self):
+            summary_d = [
+                # tf.scalar_summary('loss_d_real', self.loss_d_real),
+                # tf.scalar_summary('loss_d_fake', self.loss_d_fake),
+                tf.scalar_summary('loss_d', self.loss_d),
+                ]
+            summary_g = [
+                tf.scalar_summary('loss_g', self.loss_g)
+            ]
+            self.merged_d = tf.merge_summary(summary_d)
+            self.merged_g = tf.merge_summary(summary_g)
+
+    def _build_graph(self):
+        """Build tensorflow graph (discriminator + generator)"""
[…]

### 2016-12-12 code/Network/GAN.py
sample from random_sample in [-1,1], not uniform

@@ -5,7 +5,7 @@ import tensorflow as tf
 import prettytensor as pt
 import sys
 import os
-from numpy.random import random,uniform
+from numpy.random import uniform,random_sample
 from progressbar import ETA, Bar, Percentage, ProgressBar
 
 
@@ -116,14 +116,19 @@ class GAN(object):
                    .fully_connected(self.data_dim)).tensor
 
     def _gen_prior(self):
-        low = -2
-        high = 4
-        return uniform(low, high, (self.batch_size,self.z_dim))
+        """return z ~ prior
 
-    def decode(self):
+        normalized to [-1,1]
+        """
+        return random_sample((self.batch_size, self.z_dim)) * 2 -1
+        # low = -2
+        # high = 4
+        # return uniform(low, high, (self.batch_size,self.z_dim))
+
+    def decode(self, zs = None):
         """Sample x ~ G(z)
         """
-        prior = self._gen_prior()
+        prior = self._gen_prior() if zs is None else zs
         return self.sess.run(self.output_g, {self.noise_tensor: prior})
     
 

### 2016-12-12 code/Network/GAN.py
added summaries to merged ones

@@ -53,8 +53,8 @@ class GAN(object):
 
     def _build_summaries(self):
             summary_d = [
-                # tf.scalar_summary('loss_d_real', self.loss_d_real),
-                # tf.scalar_summary('loss_d_fake', self.loss_d_fake),
+                tf.scalar_summary('loss_d_real', self.loss_d_real),
+                tf.scalar_summary('loss_d_fake', self.loss_d_fake),
                 tf.scalar_summary('loss_d', self.loss_d),
                 ]
             summary_g = [

### 2016-12-12 code/Network/GAN.py
undo transformation

@@ -120,7 +120,7 @@ class GAN(object):
 
         normalized to [-1,1]
         """
-        return random_sample((self.batch_size, self.z_dim)) * 2 -1
+        return random_sample((self.batch_size, self.z_dim))
         # low = -2
         # high = 4
         # return uniform(low, high, (self.batch_size,self.z_dim))

### 2016-12-12 paper/chapters/gan.tex
added feature matching short text

@@
 \label{ssub:gan_feature_matching}
+In computer vision, features are often used to identify relevant pieces of information in images for recognition and classification.
+They may refer to a set of points or edges in an image, which exactly is not clearly defined.
+With the recent advances in classification tasks performed by deep convolutional networks, automatically learned features have been a large part of that success. (citation!)
+Feature matching is an technique designed to improve the learning procedure of GANs where not directly $\mathcal{L}_g$ is maximized but instead also learn the discriminator on intermediate layer activation outputs.
+Using this modified procedure the modified objective function aids the generator to output data matching the statistics of the input data.
+With $f(x)$ denoting an intermediate layer in $D$, the new objective function becomes:
+$$
+|| \mathbb{E}_{x \sim p_{data}}f(x) - \mathbb{E}_{z \sim p(z)}f(G(z)) || ^{2}_{2}
+$$
 

### 2016-12-12 paper/chapters/gan.tex
modified some text for introduction of GANs

@@
 
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
-GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
-The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
+%%% when introduced, game theory approach
+Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al based on a game theoretic approach to generative modelling.
+%which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
+
+The idea is to have two neural networks playing against each other.
+A generator $G(z)$ and a discriminator $D$ where the generator tries to produce data similar to those encountered in a training set.
+$D$ receives samples from either the training set or the generated data, without knowing where it came from.
+Both $G$ and $D$ are usually implemented as neural networks with $G$ being trained to output samples fooling the discriminator and $D$ distinguishing generated samples from training examples.
+
+%GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
+%The goal of $G$ is to produce similar data to the given training set while $D$ is tasked with discriminating(?) the generated samples from the training data.
 Both $G$ and $D$ are learning function approximators, for example Multi-Layer Peceptron (MLP) networks as proposed in the original GAN paper.
 % --> declare distribution p_data, p_model, x, z
-During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
-In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
-We don't have access to that distribution.
+%During learning, both networks are updated in parallel according to the gradient of their respective loss function which we will explore now.\\
+%In the following $p_{data}$ is the true data generating distribution from which the training samples have been generated.
+%We don't have access to that distribution.
 
@@
 \label{sub:gan_arch}
+Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
 
@@
 \subsubsection{Generator Network}
-The generator network in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
+The generator $G$ in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
 This process can be formally written as minimizing the following cost function.
 $$
-\mathcal{C}_g = \log(1 - D(G(z)))
+\mathcal{L}_g = \min \log(1 - D(G(z)))
 $$
-$z$ is produced by sampling from the noise prior $p_g(z)$.\\
+Where $z$ is sampling from the noise prior $p(z)$.\\
+
+Another common form of the generator loss is $\mathcal{L'}_g$.
+$$
+\mathcal{L'}_g = \max \log D(G(z))
+$$
+This formulation has the benefit of the same optimum fixed point, but has a better gradient flow early in training when $G$ produces samples which can be distinguished easily from real data by $D$.
 
@@
 The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{C}_d$.
+This process can be formalized to maximizing the cost function $\mathcal{L}_d$.
 $$
-\mathcal{C}_d = \log\big(\log D(x) + \log (1 - D(G(z))\big)
+\mathcal{L}_d = \max \log\big(\log D(x) + \log (1 - D(G(z))\big)
 $$
@@
 
-Combining the cost functions of generator and discriminator yield us the following objective with value function $V(G,D)$.
+Combining the loss functions of both generator and discriminator yields us the following objective with value function $V(G,D)$.
 $$

### 2016-12-12 paper/chapters/gan.tex
added stuff about (V)BN

@@
 \label{ssub:gan_batch_norm}
-Using batch normalization \cite{batch_norm:2015} helped stabilizing learning DCGAN \cite{dcgan:2015}.
+% TODO explain internal covariate shift?
+Batch normalization (BN) \cite{batch_norm:2015} is a recent technique to boost performance in deep neural networks addressing the \emph{internal covariate shift} by normalizing minibatches.
+Essentially, BN works by normalizing each sample from a minibatch to the statistics of the whole batch.
+In a paper \cite{improved_gan:2016} by Salimans et al on improving stability of the GAN training procedure, a variant of BN has been proposed called virtual batch normalization.
+
+%Using batch normalization \cite{batch_norm:2015} helped stabilizing learning DCGAN \cite{dcgan:2015}.
 \cite{improved_gan:2016} introduced virtual batch normalization where each input $x$ is normalized on one reference batch chosen at the start of learning instead of the other inputs in the mini-batch.

### 2016-12-12 paper/chapters/vae.tex
small fixes not worth an own commit
add iwae,bn,cvae papers
added conceptual vae figure
added stuff about (V)BN

@@
-\section{Variational Autoencoders}
+\section{Variational Autoencoders \cite{vae:2013}}
 \label{sec:vae}
@@
 
+\paragraph{Variational Inference} is a method for approximating intractable posterior distributions over a set of hidden varibales given observations.
+More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are hidden and $x$ observed variables into an optimization problem.
+The term \emph{variational} derives from the calculus of variations, which is a field in mathematical analysis dealing with choosing the best function $q$ from a set of functions $Q$.
+In the case of variational inference, this $Q$ denotes a class of distributions and the variational parameters are the parameters which are used to define $q \in Q$.
+VI is a alternative to markov chain monte carlo (MCMC) methods like gibbs sampling, due to lower computational cost.
+However, there is no guarantee that VI methods will asymptotically reach the optimimum in contrast to MCMC.
+
 Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling in practice(?).
+
+% TODO: add clarification that q/p are deep nets!
+\begin{figure}[htb]
+\centering
+\includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
+  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
+\end{figure}
 
@@
   \caption{MNIST 2-dimensional manifold}\label{fig:vae_manifold}
+  \medskip
+  \small
+  Manifold learned by a VAE using a 2-dimensional latent space for $z \in [-3,3]^2$.
 \end{figure}
@@
 
-\subsubsection{Deep Recurrent Attention Writer}
-\label{ssub:vae_deep_recurrent_attention_writer}
+\subsubsection{Conditional VAE \cite{cvae:2015}}
+\label{ssub:vae_conditional_vae}
+\input{chapters/vae/cvae}
 
@@
 \label{ssub:vae_importance_weighted_autoencoder}
+\input{chapters/vae/iwae}
 
-\subsubsection{Conditional VAE}
-\label{ssub:vae_conditional_vae}
+\subsubsection{Deep Recurrent Attention Writer \cite{draw:2015}}
+\label{ssub:vae_deep_recurrent_attention_writer}
+\input{chapters/vae/draw}

### 2016-12-12 paper/chapters/vae.tex
clarified relation to other autoencoders

@@
 
-\paragraph{Variational Inference} is a method for approximating intractable posterior distributions over a set of hidden varibales given observations.
+\paragraph{Variational Inference (VI)} is a method for approximating intractable posterior distributions over a set of hidden varibales given observations.
 More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are hidden and $x$ observed variables into an optimization problem.
@@
 \paragraph{Relation to Auto-Encoders}
-The Variational autoencoder has the same overall architecture than other autoencoders, for example sparse or denoising autoencoders.
-Similar to other autoencoder frameworks VAE provides an encoder as well as a decoder model, but in contrast to other autoencoders both networks are probabilistic.
+The variational autoencoder has the same encoding-decoding architecture as other autoencoders, for example sparse or denoising autoencoders.
+But in contrast to these autoencoders, the VAE enforces a probability distribution on the learned latent space $z$.
+In particular, the framework encourages the model to learn a representation that is close to $p(z)$ (which usually is an isotropic multivariate gaussian distribution) by including the KL divergence in the objective function.
+%Similar to other autoencoder frameworks VAE provides an encoder as well as a decoder model, but in contrast to other autoencoders both networks are probabilistic.
 This means that the model described by the VAE framework can be seen as a jointly trained probabilistic encoder and probabilistic decoder.

### 2016-12-12 paper/chapters/vae.tex
added argument for regularizer for VAE
added figure for reparameterization trick

@@
 \paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative Kullback-Leibler (KL) divergence between the learned variational distribution $q_\theta(z|x)$ and the latent space prior distribution $p(z)$.
+The regularizer also prevents the model from collapsing into a single point.
 
@@
 In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
+
+\begin{figure}
+  \centering
+  \begin{tabular}{c:c}
+    \subcaptionbox{without reparameterization}{\includestandalone[height=0.4\linewidth]{media/reparameterization_trick_without}}
+      \quad \quad
+    &
+      \quad
+
+    \subcaptionbox{with reparameterization applied}{\includestandalone[height=0.4\linewidth]{media/reparameterization_trick_with}}\\
+  \end{tabular}
+  \caption{reparameterization trick}
+  \medskip
+  \small
+  Function $f_\theta(x)$ with inputs $x$ and (variational) parameters $\theta$ shown as directed graphical model.\\
+  Left without and on the right with the reparameterization trick applied.\\
+  Solid arrows show the forward pass and dashed arrows the error backpropagation.
+  Rectangular nodes represent deterministic functions while circles indicate stochastic variables.
+  Note that backpropagation of gradients through stochastic nodes is problematic and adds bias (verify!)
+\end{figure}
 

### 2016-12-12 paper/chapters/gan.tex
clarified some GAN stuff

@@
 Both $G$ and $D$ are usually implemented as neural networks with $G$ being trained to output samples fooling the discriminator and $D$ distinguishing generated samples from training examples.
+
+% use of words: forgery
+To understand the inuition behind the approach to GANs, it may be helpful to view(?) analogy of counterfeiters and the police.
+Counterfeiters try to produce fake money imitating legit money while the police tries to distinguish fake from real money.
+This procedure of generating and distinguishing will go back-and-forth as the counterfeiters will get better at producing authentic looking money and the police is hard-pressed to differentiate between fake and real.
+Eventually, this procedure will result in a nash equlibrium where the police or discriminator will have 50\% chance of seperating authentic from real.
+
+Ideally, we want to find this equilibrium for the GAN model as well.
 
@@
 \label{sub:gan_training}
-The GAN framework can be fully trained with stochastic backpropagation due to both $D$ and $G$ being differentiable.
-
 Combining the loss functions of both generator and discriminator yields us the following objective with value function $V(G,D)$.
@@
 $$
-In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
-\begin{algorithm}
-  \caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
-  \label{alg:gan_online}
-  \begin{algorithmic}[1]
-    \Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
-    \Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{number of training iterations}
-      \Let{$z_d$}{sample from noise prior $p_g(z)$}
-      \Let{$z_g$}{sample from noise prior $p_g(z)$}
-      \Let{$x$}{example from training set}
-      \Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
-      \Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
-    \EndFor
-  \end{algorithmic}
-\end{algorithm}
 
-In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
-Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
-Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
-Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{C}_d$ but minimizing $\mathcal{C}_g$ (correct?).
+The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable.
+Training will oscillate between training the generator and discriminator, where the other party is fixed respectively.
+In order to avoid a too strong discriminator (or generator), the ratio of learning steps will vary even though there are no hard rules for this.
+
+\paragraph{Theoretical Foundation}
+Using a model with infinite training time and capacity in both generator and discriminator allows theoretical analysis of the GAN framework \cite{gan:2013}.
+For any given generator $G$, the optimal discriminator $D$ is given by $D_G^*(x)$:
+$$
+D_G^*(x) = \frac{p_{data}(x)}{p_{data}(x) + p_{model}(x)}
+$$
+
+%In the original GAN paper an minibatch-based algorithm for training GAN is proposed, but we will take a look at an the online learning version below:\\
+%\begin{algorithm}
+  %\caption{Online learning of generative adversarial networks $-$ simple version ($k=1$)}
+  %\label{alg:gan_online}
+  %\begin{algorithmic}[1]
+    %\Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
+    %\Let{$\theta_d, \theta_g$}{Parameters of function approximators}
+    %\For{number of training iterations}
+      %\Let{$z_d$}{sample from noise prior $p_g(z)$}
+      %\Let{$z_g$}{sample from noise prior $p_g(z)$}
+      %\Let{$x$}{example from training set}
+      %\Let{$\theta_d$}{$\theta_d + \eta_d \nabla_{\theta_d} \bigg[\log D(x) + \log \big(1 - D(G(z_d))\big)\bigg]$}
+      %\Let{$\theta_g$}{$\theta_g - \eta_g \nabla_{\theta_g} \bigg[\log\big(1 - D(G(z_d))\big)\bigg]$}
+    %\EndFor
+  %\end{algorithmic}
+%\end{algorithm}
+
+%In algorithm \ref{alg:gan_online} the original algorithm has been modified for single pass (online) learning.
+%Sampling minibatches from the noise prior and $p_{data}(x)$ result in the minibatch version.
+%Additionally, it is often required for practical reasons to let the discriminator run multiple backward passes while updating the generator once (why?).
+%Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{L}_d$ but minimizing $\mathcal{L}_g$ (correct?).
 

### 2016-12-12 paper/chapters/introduction.tex
more on introduction

@@
 \label{sec:introduction}
-Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
-More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
-This assumption is reasonable given the following examples:\\
 
+"What I cannot create, I do not understand." -- Richard Feynman \\\\\\
+
+% on building reasonable representations
+One of the key motivations to generative modelling is the assumption that having a interpretable representation of the real world is important to building intelligent agents.
+Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
+Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
+Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\
+
+With this assumption in mind we can take a look at some examples of input data.
 \begin{itemize}
-  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler.
+  \item \textbf{Images} commonly have thousands of dimensions, but the underlying structure is often far simpler.
 For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
@@
 
-After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two promising frameworks, the variational autoencoder (VAE) and the generative adversarial network (GAN)
+To formalize this setting, it is often beneficial to view the structure as a probability distribution $p$ on the input data $x$ we build the structure 
+%( another word!)
+from: $p(x)$.
+
+The assumption that $p(x)$ is dependent on some underlying structure can now be expressed as having the conditional probability $p(x|z)$ where $z$ are the causal factors.
+We call $z$ also the latent variables.\\
+
+%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.
+
+After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two recent promising frameworks, the variational autoencoder (VAE \ref{sec:vae}) and the generative adversarial network (GAN \ref{sec:gan}).
+
+%Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
+%More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
+%This assumption is reasonable given the following examples:\\
 

### 2016-12-12 paper/chapters/vae.tex
remove theta from prior distribution
sbn + VI + draw

@@
 We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
-The prior distribution $p_\theta(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
+The prior distribution $p(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
 The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2013} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
@@
 \begin{equation}
-  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p_\theta(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
+  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
 \end{equation}
@@
 \label{sub:vae_inference}
+In bayesian analysis, statistical inference is the process of evaluating $p(z|x)$ for a given conditional probability $p(x|z)$.
+% more info on marginal distribution, posterior inference etc!
+There are two approaches to performing this task, Markov chain Monte Carlo (MCMC) methods using Gibbs sampling and variational inference.
+MCMC methods have the advantage of being asympotically exact and unbiased while variational methods are computationally faster but biased and not guaranteed to approaching the distribution.
+In the case of VAE, variational inference is also used in order to backpropagate through the layers (using the reparameterization trick).
+
 \begin{figure}[htb]

### 2016-12-12 paper/chapters/vae/draw.tex
sbn + VI + draw

The \emph{deep recurrent attention writer}, in short DRAW is an advanced model uses the selective attention mechanism in addition to a VAE with both recurrent encoder and recurrent decoder.
DRAW allows to basically the pixels of an image sequentially very human-like.

### 2016-12-12 paper/chapters/vae/iwae.tex
sbn + VI + draw

@@
+% write about multiple stochastic layers
+Until now, the presented model includes only one stochastic layer.
+\cite{dlgm:2014} have presented a model for learning through multiple stochastic layers which can be easily applied to the VAE framework as shown in \ref{fig:vae_multiple_layers}.
+
+\begin{figure}[htb]
+\centering
+\includestandalone[width=\linewidth,mode=buildnew]{media/vae_multiple_layers}
+  \caption{VAE with multiple stochastic layers }\label{fig:vae_multiple_layers}
+  \medskip
+  \small
+  The left upward arrows indicate the recognition networks $q(h_1|x)$ and $q(h_2|h_1)$.
+  Note that the recognition networks can have multiple deterministic layers (including non-linearities).\\
+
+  $p(h_2)$ denotes the probability distribution over the highest(?) latent space.\\
+
+  On the right, the two conditional distributions $p(h_1|h_2)$ and $p(x|h_2)$ are representing two networks used for the generative process.
+\end{figure}
+% write about k passes (just averages)
+
+% write about IWAE, extends these k passes with importance weights
 Importance Weighted Autoencoders (IWAE) extends on the VAE framework using $k$-sample importance weighting of the log-likelihood.
+% add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
+Using importance weighting can be informally interpreted as putting more weight on data with high likelihood.
+
 The authors of the papers argue that using more samples will only tighten the lower variational bound.

### 2016-12-12 code/Network/VAE.py
removed dropout, refactored to pt

@@ -2,6 +2,7 @@
 from datetime import datetime
 import numpy as np
 import tensorflow as tf
+import prettytensor as pt
 import sys
 import os
 import logging
@@ -14,14 +15,12 @@ from progressbar import ETA, Bar, Percentage, ProgressBar
 import matplotlib.pyplot as plt
 from IPython import display
 
-
 class VAE(object):
     """ Variational Autoencoder implementation in TensorFlow"""
 
     HYPERPARAMS = {
             "batch_size": 128,
             "learning_rate": 1e-3,
-            "dropout": 1.0,
             "nonlin": tf.nn.relu, # TODO relu
             "squash": tf.nn.sigmoid, # TODO explore (softplus?)
             }
@@ -29,121 +28,116 @@ class VAE(object):
     def __init__(self,
             network_architecture = [], # list of nodes per layer
             hyper_params = {}, # update to HYPERPARAMS
-            models_dir = "models/"
+            models_dir = "models/",
+            k = 1, # number of forward passes
+            **kwargs
             ):
+        self.k_passes = k
+        self.arch = network_architecture
         if not os.path.isdir(models_dir):
             os.makedirs(models_dir)
         self.models_dir = models_dir
         # insert updated hyper parameters into class
         tf.reset_default_graph()
-        self.__dict__.update(VAE.HYPERPARAMS, **hyper_params)
+        self.__dict__.update(VAE.HYPERPARAMS, **kwargs)
         self.sess = tf.Session()
 
-        print('Building tensorflow graph')
+        print('Building tensorflow graph with the following hyper parameters:')
+        for k in VAE.HYPERPARAMS:
+            print('\t{}: {}'. format(k, self.__dict__[k]))
+        print('With the following network architecture: [{}]'.format(",".join(map(str,self.arch))))
 
-        # build tf graph according to network_architecture and hyperparams
-        handles = self._build_graph(network_architecture)
+
+        # build tf graph according to hyperparams
+        self._build_graph()
 
         self.constructed = datetime.now().strftime("%y%m%d_%H%M")
         logs_path = os.path.join("logs", "run_"+self.constructed)
-        self.summary_writer = tf.train.SummaryWriter(logs_path, self.sess.graph)
-        self.writer = tf.merge_all_summaries()
+        self.summary_writer = tf.summary.FileWriter(logs_path, self.sess.graph)
[…]

### 2016-12-12 code/Network/GAN.py
noise tensor is now initialized randomly by default

@@ -66,7 +66,11 @@ class GAN(object):
     def _build_graph(self):
         """Build tensorflow graph (discriminator + generator)"""
         self.data_tensor = tf.placeholder(np.float32, shape=(self.batch_size, self.data_dim))
-        self.noise_tensor = tf.placeholder(np.float32, shape=(self.batch_size, self.z_dim))
+        minval = -1
+        maxval = 1
+        self.noise_tensor = tf.get_variable("z", (self.batch_size, self.data_dim),
+                                initializer=tf.random_uniform_initializer(minval, maxval),
+                                trainable=False)
 
 
         self.output_g = self._build_generator(self.noise_tensor)

### 2016-12-12 code/Network/GAN.py
moved count to class, clarified noise tensor stuff

@@ -14,6 +14,7 @@ class GAN(object):
 
     Just for educational purposes, generator + discriminator implemented using MLP.
     """
+    count = 0
 
     HYPERPARAMS = {
             "batch_size": 128,
@@ -125,23 +126,17 @@ class GAN(object):
         normalized to [-1,1]
         """
         return random_sample((self.batch_size, self.z_dim))
-        # low = -2
-        # high = 4
-        # return uniform(low, high, (self.batch_size,self.z_dim))
 
     def decode(self, zs = None):
-        """Sample x ~ G(z)
-        """
+        """Sample x ~ G(z)"""
         prior = self._gen_prior() if zs is None else zs
         return self.sess.run(self.output_g, {self.noise_tensor: prior})
-    
 
 
     def train(self, X, num_epochs = 2000, modulo = 100, output = True):
         now = datetime.now().isoformat()[11:]
         if output:
             print("[**] Begin training: {}".format(now))
-        count = 0
         l_d = 0
         summary_d = None
         n_k = self.k
@@ -158,7 +153,7 @@ class GAN(object):
             # update generator once
             prior = self._gen_prior()
             l_g,summary_g, _ = self.sess.run([self.loss_g, self.merged_g, self.opt_g], {
-                self.noise_tensor: prior
+                    self.noise_tensor: prior
                 })
 
 
@@ -172,13 +167,13 @@ class GAN(object):
                             self.merged_g
                         ], {
                             self.data_tensor: X,
-                            self.noise_tensor: prior,
+                            # self.noise_tensor: prior,
                         })
-                self.writer.add_summary(summary_d, count)
-                self.writer.add_summary(summary_g, count)
+                self.writer.add_summary(summary_d, self.count)
+                self.writer.add_summary(summary_g, self.count)
                 if output:
-                    print('{:>5}@{:<2} loss_d: {:.3}\tloss_g: {:.3}'.format(count, n_k, l_d, l_g))
-                count += 1
+                    print('{:>5}@{:<2} loss_d: {:.3}\tloss_g: {:.3}'.format(self.count, n_k, l_d, l_g))
+                self.count += 1
 
 
         now = datetime.now().strftime("%y%m%d_%H%M")

### 2016-12-12 code/Network/VAE.py
minor changes to plotting, TODO

@@ -235,14 +235,21 @@ class VAE(object):
             plt.close('all')
         print("done")
 
-    def plot_manifold(self, range_x=(-4,4), range_y=(-4,4), num_rowdigits = 20, output="manifold.pdf"):
+    def plot_manifold(self, range_x=(-4,4), range_y=(-4,4), axis = (1,2), single_axis = False, num_rowdigits = 20, output="manifold.pdf"):
         x1 = np.linspace(range_x[0], range_x[1], num = num_rowdigits)
-        x2 = np.linspace(range_y[0], range_y[1], num = num_rowdigits)
+        if not single_axis:
+            x2 = np.linspace(range_y[0], range_y[1], num = num_rowdigits)
+        else:
+            x2 = np.zeros((1, num_rowdigits))
 
         manifold = np.zeros(shape=(len(x1)*28, len(x2)*28))
         for i_x, x in enumerate(x1):
             for i_y, y in enumerate(x2):
-                img = self.decode([[x,y]])
+                input = np.zeros(shape=(1, self.arch[-1]))
+                input[0][axis[0]] = x
+                if not single_axis:
+                    input[0][axis[1]] = y
+                img = self.decode(input)
                 manifold[i_x*28:(i_x+1)*28,i_y*28:(i_y+1)*28] = img.reshape(28,28)
         plt.imshow(manifold, cmap=plt.cm.gray)
         plt.axis('off')
@@ -251,11 +258,17 @@ class VAE(object):
     def plot_z_mean(self, data, num_samples = 5000):
         x, y = data.test.next_batch(num_samples)
         z_mean, _ = self.encode(x)
-        plt.scatter(z_mean[:,0], z_mean[:,1], c=np.argmax(y, 1), alpha=1, edgecolors='black')
-        plt.show()
+        if self.arch[-1] == 2:
+            plt.scatter(z_mean[:,0], z_mean[:,1], c=np.argmax(y, 1), alpha=1, edgecolors='black')
+            plt.show()
+        elif self.arch[-1] == 3:
+            fig = plt.figure()
+            ax = fig.add_subplot(111, projection='3d')
+            ax.scatter(z_mean[:,0], z_mean[:,1], z_mean[:,2], c=np.argmax(y,1), s = 40, marker='.')
+            plt.show()
 
 
-    
+
     def train(self, X, num_epochs = 75, plot_manifold = False):
         avg_train_error = 0
         now = datetime.now().isoformat()[11:]

### 2016-12-12 code/Utils.py
added utils, lrelu, BN, conv/conv_t_2d

@@ -10,3 +10,82 @@ def compose_all(*args):
     """
     return partial(functools.reduce, compose)(*args)
 
+def lrelu(x, leak=0.2, name="lrelu"):
+    """Leaky ReLU"""
+    with tf.variable_scope(name):
+        f1 = 0.5 * (1 + leak)
+        f2 = 0.5 * (1 - leak)
+        return f1 * x + f2 * abs(x)
+
+
+class batch_norm(object):
+    """Code modification of http://stackoverflow.com/a/33950177"""
+    def __init__(self, epsilon=1e-5, momentum = 0.9, name="batch_norm"):
+        with tf.variable_scope(name):
+            self.epsilon = epsilon
+            self.momentum = momentum
+
+            self.ema = tf.train.ExponentialMovingAverage(decay=self.momentum)
+            self.name = name
+
+    def __call__(self, x, train=True):
+        shape = x.get_shape().as_list()
+
+        if train:
+            with tf.variable_scope(self.name) as scope:
+                self.beta = tf.get_variable("beta", [shape[-1]],
+                                    initializer=tf.constant_initializer(0.))
+                self.gamma = tf.get_variable("gamma", [shape[-1]],
+                                    initializer=tf.random_normal_initializer(1., 0.02))
+
+                batch_mean, batch_var = tf.nn.moments(x, [0, 1, 2], name='moments')
+                ema_apply_op = self.ema.apply([batch_mean, batch_var])
+                self.ema_mean, self.ema_var = self.ema.average(batch_mean), self.ema.average(batch_var)
+
+                with tf.control_dependencies([ema_apply_op]):
+                    mean, var = tf.identity(batch_mean), tf.identity(batch_var)
+        else:
+            mean, var = self.ema_mean, self.ema_var
+
+        normed = tf.nn.batch_norm_with_global_normalization(
+                x, mean, var, self.beta, self.gamma, self.epsilon, scale_after_normalization=True)
+
+        return normed
+
+
+# Conv2D Layer
+def conv2d(input_, out_channels, filter_h=5, filter_w=5, stride_vert=2, stride_horiz=2, stddev=0.02, name="conv2d"):
+    with tf.variable_scope(name):
+        # Get the number of input channels
+        in_channels = input_.get_shape()[-1]
+
+        # Construct filter
+        w = tf.get_variable('w', [filter_h, filter_w, in_channels, out_channels],
[…]

### 2016-12-12 paper/chapters/vae.tex
some minor clarifications (added link) to VAE

@@
 \subsubsection{Reparameterization Trick}
+
+% more math:
+% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
 To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
@@
 \label{sub:vae_learning}
-In practice, the usual choice for $q_\theta(z|x)$ is the multivariate gaussian distribution.
-$$
-q_\theta(z|x) = \mathcal{N}(z|\mu(X;\vartheta), \Sigma(X;\vartheta))
-$$
-Because we assume that $P(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
+The probabilistic encoder $q_\theta(z|x)$ is commonly implemented using a deep neural network which transforms the input data $x$ into a multivariate gaussian distribution with mean $\mu_0$ and diagonal variance $\Sigma_0$.
+To allow the computation of the KL divergence between $q_\theta(z|x)$ and the latent prior distribution $p(z)$ in closed form, the prior is an isotropic gaussian $\mathcal{N}(0, 1)$.
+
+% define X correctly or replace formula below!
+%$$
+%q_\theta(z|x) = \mathcal{N}(z|\mu(X;\theta), \Sigma(X;\theta))
+%$$
+
+%Because we assume that $p(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
 \begin{align*}
-  \mathcal{D}_{\mathrm{KL}}\big[Q(z) || P(z|x)\big] &= \mathbb{E}\big[\log \frac{p}{q}\big]\\
-  \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]
-  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)
+  \mathcal{D}_{\mathrm{KL}}\big[q_\theta(z|x) || p_\phi(z|x)\big] &= \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]\\
+  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)\\
 \end{align*}
@@
 
-\begin{figure}[htb]
+\begin{figure}[h!]
 \centering

### 2016-12-12 code/Network/VAE.py
added TODO for VAE

@@ -16,7 +16,13 @@ import matplotlib.pyplot as plt
 from IPython import display
 
 class VAE(object):
-    """ Variational Autoencoder implementation in TensorFlow"""
+    """Variational Autoencoder implementation in TensorFlow using PrettyTensor
+
+    TODO:
+    - batch normalization
+    - convolutional network
+    - inverse flow stuff
+    """
 
     HYPERPARAMS = {
             "batch_size": 128,

### 2016-12-14 code/Network/VAE.py
added some TODO items, removed model saver

@@ -21,7 +21,8 @@ class VAE(object):
     TODO:
     - batch normalization
     - convolutional network
-    - inverse flow stuff
+    - inverse normalizing flow stuff
+    - dropout
     """
 
     HYPERPARAMS = {
@@ -61,7 +62,7 @@ class VAE(object):
         logs_path = os.path.join("logs", "run_"+self.constructed)
         self.summary_writer = tf.summary.FileWriter(logs_path, self.sess.graph)
         self.writer = tf.summary.merge_all()
-        self.saver = tf.train.Saver()
+        # self.saver = tf.train.Saver()
 
         self.sess.run(tf.global_variables_initializer())
 
@@ -304,8 +305,8 @@ class VAE(object):
 
                 # write summary every batch
                 self.summary_writer.add_summary(summary, epoch)
-            model_path = os.path.join(self.models_dir, "model_{}_epoch_{:03d}.cpkt".format(self.constructed, epoch))
-            save_path = self.saver.save(self.sess, model_path)
+            # model_path = os.path.join(self.models_dir, "model_{}_epoch_{:03d}.cpkt".format(self.constructed, epoch))
+            # save_path = self.saver.save(self.sess, model_path)
             print("epoch {}: avg cost: {}".format(epoch, avg_error_train))
             if plot_manifold:
                 self.plot_manifold(range_x=20, range_y=20, output = "manifold_{}.pdf".format(epoch))

### 2016-12-18 code/Network/VAE.py
fixed range for manifold plotting

@@ -242,7 +242,7 @@ class VAE(object):
             plt.close('all')
         print("done")
 
-    def plot_manifold(self, range_x=(-4,4), range_y=(-4,4), axis = (1,2), single_axis = False, num_rowdigits = 20, output="manifold.pdf"):
+    def plot_manifold(self, range_x=(-4,4), range_y=(-4,4), axis = (0,1), single_axis = False, num_rowdigits = 20, output="manifold.pdf"):
         x1 = np.linspace(range_x[0], range_x[1], num = num_rowdigits)
         if not single_axis:
             x2 = np.linspace(range_y[0], range_y[1], num = num_rowdigits)
@@ -309,7 +309,7 @@ class VAE(object):
             # save_path = self.saver.save(self.sess, model_path)
             print("epoch {}: avg cost: {}".format(epoch, avg_error_train))
             if plot_manifold:
-                self.plot_manifold(range_x=20, range_y=20, output = "manifold_{}.pdf".format(epoch))
+                self.plot_manifold(range_x=(-1,1), range_y=(-1,1), output = "manifold_{:03}.pdf".format(epoch))
                 print("done plotting manifold")
 
         now = datetime.now().strftime("%y%m%d_%H%M")

### 2016-12-18 paper/chapters/overview.tex
small fixes & added citations
sbn + VI + draw
VAE & GAN -> differentiable generator nets

--- 2016-12-01 paper/chapters/overview.tex
@@
 
-\paragraph{Variational Autoencoders (VAE) \cite{vae:2013}\cite{rezende:2014}}
-\label{par:overview_vae}
-Variational Autoencoders (VAE) use two networks which are trained jointly using approximated learned inference.
-The objective function is aimed at building a reasonable simple latent space approximating the intractable posterior distribution using variational inference.
-It consists of the recognition and inference network which can be understood in the autoencoder jargon as probabilistic encoder and decoder.
-Using variational bayesian methods the intractable posterior distribution can be approximated and learned using backpropagation and gradient-based methods.
+\paragraph{Sigmoid Belief Nets (SBN)}
+\label{par:overview_sbn}
+SBNs are one of the earliest generative models, using a simple architecture in which sampling from is easy, but learning and performing inference is not.
+
+\paragraph{Differentiable Generator Networks}
+Due to the success of gradient-based methods for learning deep networks, a large part of recent models are using a differentiable generator network which allows for error backpropagation.\\
+Most prominent examples of models with differentiable generator nets are the variational autoencoder (VAE) \cite{vae:2014} and generative adversarial networks (GAN) \cite{gan:2014}.
+These models include a generator networks representing a family of parametrized non-linear functions $g$ implemented usually implemented as neural networks.\\
+Using $g$ as distribution transformation allows samples $z$ to be transformed into $x$ where the distribution over $x$ may be more complicated than $z$.
+The parameters of $g$ can be inferred using training data and an appropiate loss function.\\
+However, instead of maximizing the log-likelihood of $p(x)$ directly we choose $g$ to emit not samples of $x$ but parameters of a conditional distribution over $x$.
+
+\paragraph{Auto-regressive networks}
+
+%\paragraph{Variational Autoencoders (VAE) \cite{vae:2013}\cite{rezende:2014}}
+%\label{par:overview_vae}
+%Variational Autoencoders (VAE) use two networks which are trained jointly using approximated learned inference.
+%The objective function is aimed at building a reasonable simple latent space and approximating the intractable posterior distribution using variational inference.
+%It consists of the recognition and inference network which can be understood in the autoencoder jargon as probabilistic encoder and decoder.
+%Using variational bayesian methods the intractable posterior distribution can be approximated and learned using backpropagation and gradient-based methods.
 
@@
 
-\paragraph{Generative Adversarial Networks}
-\label{par:overview_gan}
-GANs are a game theoretic approach to generative modelling
-in which two networks compete against each other.
-The discriminator $D$ is trying to distinguish the generated samples
-by the generator $G$ from real world data.
-The goal for $G$ however is to produce data as realistic as possible.\\
-During the early phase of training, $G$ produces random noise which
-is easy to differentiate from groundtruth data.
-Both networks are trained in parallel, where both networks try to get better
-at their respective objective.
-In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
-case the discriminator will only be able to guess correctly with a 50\% chance.
-This plateau is called Nash equilibrium (citation needed).
+%\paragraph{Generative Adversarial Networks\cite{gan:2014}}
+%\label{par:overview_gan}
+%GANs are a game theoretic approach to generative modelling
+%in which two networks compete against each other.
+%The discriminator $D$ is trying to distinguish the generated samples
+%by the generator $G$ from real world data.
+%The goal for $G$ however is to produce data as realistic as possible.\\
+%During the early phase of training, $G$ produces random noise which
+%is easy to differentiate from groundtruth data.
+%Both networks are trained in parallel, where both networks try to get better
+%at their respective objective.
+%In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
+%case the discriminator will only be able to guess correctly with a 50\% chance.
+%This plateau is called Nash equilibrium (citation needed).

### 2016-12-18 paper/chapters/vae.tex
added reconstruction and samples for MNIST@VAE

@@
 
-\begin{figure}[h!]
+\begin{figure*}[t!]
     \centering
-%\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
-\includegraphics{media/manifold.pdf}
-  \caption{MNIST 2-dimensional manifold}\label{fig:vae_manifold}
-  \medskip
-  \small
-  Manifold learned by a VAE using a 2-dimensional latent space for $z \in [-3,3]^2$.
-\end{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=7cm]{media/vae_manifold-crop.pdf}
+        \caption{Learned manifold using 2-dimensional latent space}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=7cm]{media/vae_samples-crop.pdf}
+        \caption{generated images by sampling from latent space and applying $p(x|z)$}
+    \end{subfigure}
+    \caption{Variational autoencoder learning performance}
+\end{figure*}
 

### 2016-12-19 paper/chapters/abstract.tex
updated

@@
 \label{sec:abstract}
-Deep Learning has advanced at an incredible speed the last few years and has achieved human-level performance in some discrimination and classification tasks (object recognition, speech recognition).
-These supervised methods have the huge drawback of requiring a vast amount of labelled data which is often expensive or unfeasible to obtain.
-Learning hidden structure from unlabelled data known as unsupervised learning has had less of an impact recently.
-However, it can be argued that this method is most similar to human learning, in particular inferring information from observations and building an appropriate model from it.
-We will discuss and explore some models for performing this inference as well as generating new data from these representations.
-Building reasonable representation of the world from unlabelled data is arguably one of the key aspects of machine intelligence.
+%One of the key aspects of human intelligence is the ability to construct accurate and generalized internal representations by observing high-dimensional data.
+%We present several deep directed generative models aiming to learn rich and hierarchical latent structures, either implicitly or explicitly, as well as transforming them back into input space.
+
+Deep directed generative models leverage the possibilities of deep learning to learn rich, hierarchical structures from datasets in a possibly unsupervised context.
+These models can be used to understand high-dimensional, real-world data and transform them into compact representations.
+Similar to how humans learn to generalize from sensory input, we discuss several models which are able to learn compact, internal representational structures.
+In this article we will focus on two recently proposed architectures, variational autoencoders and generative adversarial networks.
+
+% maybe include:
+% * unsupervised models are in high demand due to the massive increase in the amount of unlabelled data due to web, videos, internat in general, laboratory measurements, etc.
+
+%Deep Learning has advanced at an incredible speed the last few years and has achieved human-level performance in some discrimination and classification tasks (object recognition, speech recognition).
+%These supervised methods have the huge drawback of requiring a vast amount of labelled data which is often expensive or unfeasible to obtain.
+%Learning hidden structure from unlabelled data known as unsupervised learning has had less of an impact recently.
+%However, it can be argued that this method is most similar to human learning, in particular inferring information from observations and building an appropriate model from it.
+%We will discuss and explore some models for performing this inference as well as generating new data from these representations.
+%Building reasonable representation of the world from unlabelled data is arguably one of the key aspects of machine intelligence.
 %However one of the key aspects to machine intelligence is learning a rich (?) model about the world and draw conclusions from it.

### 2016-12-19 paper/chapters/gan.tex
updated

@@
 
+\newpage
+
 %GANs can be understood as a two-player game where one player is called generator $G$ playing against the discriminator $D$.
@@
 %We don't have access to that distribution.
-
-\subsection{Architecture}
-\label{sub:gan_arch}
-Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
-
-% http://www.slideshare.net/xavigiro/deep-learning-for-computer-vision-generative-models-and-adversarial-training-upc-2016
-% TODO: recreate in tikz
 
@@
 
+\newpage
+
+\subsection{Architecture}
+\label{sub:gan_arch}
+Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
+
+% http://www.slideshare.net/xavigiro/deep-learning-for-computer-vision-generative-models-and-adversarial-training-upc-2016
+% TODO: recreate in tikz
+
 %\begin{figure}[htb]
@@
 This process can be formally written as minimizing the following cost function.
+%Where $z$ is sampling from the noise prior $p(z)$.\\
+
+%Another common form of the generator loss is $\mathcal{L'}_g$.
+%$$
 $$
-\mathcal{L}_g = \min \log(1 - D(G(z)))
-$$
-Where $z$ is sampling from the noise prior $p(z)$.\\
-
-Another common form of the generator loss is $\mathcal{L'}_g$.
-$$
+\mathcal{L}_g = \min \log(1 - D(G(z)))\\
 \mathcal{L'}_g = \max \log D(G(z))
@@
 
+\newpage
+
 \subsection{Training}
@@
 
+\vspace{5cm}
 \paragraph{Theoretical Foundation}
-Using a model with infinite training time and capacity in both generator and discriminator allows theoretical analysis of the GAN framework \cite{gan:2013}.
+Using a model with infinite training time and capacity in both generator and discriminator allows theoretical analysis of the GAN framework \cite{gan:2014}.
 For any given generator $G$, the optimal discriminator $D$ is given by $D_G^*(x)$:
@@
 %Note that the discriminator will ascend while the generator descend its gradient due to maximizing $\mathcal{L}_d$ but minimizing $\mathcal{L}_g$ (correct?).
+
+\newpage
 
@@
 
+\vspace{10cm}
+
 \subsubsection{Historical Averaging}
@@
 
-\subsection{Application}
-\label{sub:gan_application}
-GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
+\newpage
+
+%\subsection{Application}
+%\label{sub:gan_application}
+%GAN have been applied to many different areas which include generating images (GAN, DCGAN, LAPGAN), sequences (SeqGAN), videos (\cite{gan_video:2016}), 3D objects (\cite{gan_3d:2016}), text-to-image synthesis (\cite{gan_t2i:2016})
 

### 2016-12-19 paper/chapters/introduction.tex
updated

--- 2016-12-06 paper/chapters/introduction.tex
@@
 \label{sec:introduction}
-Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
-More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
-This assumption is reasonable given the following examples:\\
 
+"What I cannot create, I do not understand." -- Richard Feynman \\\\\\
+
+Directed generative models are a wide class of models used for generating samples from an unknown distribution.
+Lets denote $p_{data}(x)$ as the real-world data distribution which is unknown, however we have access to samples from it: $x \sim p_{data}(x)$.
+In order to do so, we define an approximated distribution $p_{model}$ from which we are able to sample.
+Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\
+
+Generative modelling has several important application areas, which will be highlighted on below.\\
+Deep learning techniques have provided great success in applying gradient-based optimization methods to deep neural networks to achieve impressive results in discriminative models.
+These models require however large amounts of labelled data, which is often expensive or infeasible to obtain.
+Generative models can be used to learn \emph{representational} structures from rich, unlabelled data by having less parameters than the input data and are thus forced to capture only essential information.\\
+\newpage
+
+%Prior to discussion individual models, we need to declare a few assumptions and definitions.
+%Additionally we assume that samples $x$ are also conditionally dependent on some hidden structure $z$: $x \sim p(x|z)$. 
+%\begin{equation}
+  %z \sim p(z)\\
+  %x \sim p(x|z)
+%\end{equation}
+
+%\paragraph{Sampling}
+%means for any given $z$ to draw a sample $x$ from our model.
+%Most commonly, sampling is the easiest requirement to fulfil.
+
+%\paragraph{Learning}
+%of the model parameters through maximum likelihood estimation generally involves $p(x)$ or its gradient.\\
+%To get $p(x)$, we first need to marginalize our the latent variables.
+%\begin{equation}
+  %p(x) = \sum_h p(x|z)p(z)
+%\end{equation}
+%Unfortunately, in the case of deep learning with many layers this is intractable, as we need to evaluate all model configurations.
+%via maximum likelihood generally involves the gradients of $p(x)$
+%\paragraph{Inference}
+% motivation / what we want exactly and why
+%Building intelligent systems requires having accurate and generalized information about the environment in order to perform actions or reason about the world.
+%Inferring structured and meaning information from unlabelled data, a process called unsupervised learning, remains a challenging task.
+
+%However, due to the massive increases in availability of data emerging from the web as well as the ongoing progress in computational power there has been significant progress using this data.
+
+%Generative modelling aims to generate new data similar to previously seen input, in a way which is sound with the internal structure of the data.
+
+%% on building reasonable representations
+The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
+In order to understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real world data.
 \begin{itemize}
-  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler.
+  \item \textbf{Images} consisting of millions of pixels can be interpreted as data where each pixel represents an own dimension.
+  However, these dimensions are high correlated and structured, for example nearby pixels have with high probability similar values.
+  Looking at images of handwritten digits, the data can be reduced to a handful variables which may include the digit number, stroke width, position or size.
+  Statistical models help us discover these relations and reduce them to representations in lower dimensional space.
+  %\item commonly have thousands of dimensions, but the underlying structure is often far simpler.
 For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
@@
 We will take a look at this example later on in \ref{sec:vae}.
-  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
+  \item \textbf{Audio data} is omni-present in the real world, however most data sources can be transformed from raw audio samples to spoken words or musical notes without losing too much information for reconstruction.
   \item \textbf{...}
 \end{itemize}
+% cope with
+%To develop systems learning from real world data we use statistical models which are able to infer the underlying structure from
+%Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
+%Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
+%Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\
 
-After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two promising frameworks, the variational autoencoder (VAE) and the generative adversarial network (GAN)
+%With this assumption in mind we can take a look at some examples of input data.
+
+The assumption that $p(x)$ is dependent on some underlying structure can now be expressed as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.
+This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\
+
+\begin{figure}[htb]
+\centering
+\includegraphics{media/directed_graphical_model}
+  \caption{Directed graphical model}
+  \label{fig:dgm}
+  \medskip
+  \small
+  Arrows show stochastic dependency.
+\end{figure}
+
+%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.
+
+%After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two recent promising frameworks, the variational autoencoder (VAE \ref{sec:vae}) and the generative adversarial network (GAN \ref{sec:gan}).
+
+%Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
+%More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
+%This assumption is reasonable given the following examples:\\
 

### 2016-12-19 paper/chapters/overview.tex
updated

@@
+\newpage
+
 \section{Overview of Directed Generative Models}
@@
 
-\paragraph{Sigmoid Belief Nets (SBN)}
+\subsection{Sigmoid Belief Nets}
 \label{par:overview_sbn}
-SBNs are one of the earliest generative models, using a simple architecture in which sampling from is easy, but learning and performing inference is not.
+Sigmoid Belief Nets (SBN) are a specific type of bayesian networks consisting of an set of binary states with each state depending on activations of preceeding nodes.
 
-\paragraph{Differentiable Generator Networks}
-Due to the success of gradient-based methods for learning deep networks, a large part of recent models are using a differentiable generator network which allows for error backpropagation.\\
-Most prominent examples of models with differentiable generator nets are the variational autoencoder (VAE) \cite{vae:2014} and generative adversarial networks (GAN) \cite{gan:2014}.
-These models include a generator networks representing a family of parametrized non-linear functions $g$ implemented usually implemented as neural networks.\\
-Using $g$ as distribution transformation allows samples $z$ to be transformed into $x$ where the distribution over $x$ may be more complicated than $z$.
-The parameters of $g$ can be inferred using training data and an appropiate loss function.\\
-However, instead of maximizing the log-likelihood of $p(x)$ directly we choose $g$ to emit not samples of $x$ but parameters of a conditional distribution over $x$.
+Bayesian networks or belief networks are an probabilistic interpretation of an directed acyclic graphical model with a set of states $s$.
+As usual in the bayesian framework, each state can be interpreted as a random variable with their conditional probabilities.
+%SBNs are one of the earliest deep directed generative models, consisting of binary states $s$ most commonly divided into many layers.
 
-\paragraph{Auto-regressive networks}
+%where in the most common form the architecture consists of several layers where each layer $h_i$ is connected from the previous layer $h_{i-1}$.
+%Sampling is an easy task in this model, however inferring good structure for the first layer is a difficult task.
+%Usually, in the context of probabilistic generative models we are mostly interested in three different aspects.
+%First, sampling from the network. This means given 
+%We are mostly interested in three different things involving
+
+%\vspace{3cm}
+%using a simple architecture in which sampling from is easy, but learning and performing inference is not.
+
+\subsection{Differentiable Generator Networks}
+Most of recent success in deep learning can be attributed to the backpropagation algorithm, including several useful extensions like dropout (citation), or piecewise-linear activation units providing a well-behaving gradient.
+Most prominent examples of recent directed generative models leverage these powerful gradient-based optimization techniques by including a generator network.
+
+Most prominent examples of models with differentiable generator nets are the variational autoencoder (VAE) \cite{vae:2014} and generative adversarial networks (GAN) \cite{gan:2014}, which we will discuss in detail later in \ref{sec:vae} and \ref{sec:gan} respectively.
+
+%Due to the success of gradient-based methods for learning deep neural networks, a large part of recent models are using a differentiable generator network allowing error backpropagation.\\
+
+%Both of these models contain a generator network coupled with another network which enables the models to be learned end-to-end.
+Both models include generator networks representing a family of parametrized non-linear functions $g_\varphi$ parametrized by $\varphi$ where these parameters can be learned by gradient-based techniques.\\
+As neural networks are basically powerful function approximators, they can also be understood as a distribution transformation.
+Given samples $z$ from one distribution $p(z)$, applying $g$ is equivalent to sampling from the posterior distribution $p(x|z)$ which is generally more complicated than the prior $p(z)$.\\
+
+%which transforms samples $z$ from a distribution $p(z)$ into $x$ values from a different, possible more complicated distribution $p(x|z)$.
+The advantage of this approach is that powerful gradient-based optimization techniques can be used to learn the transformation $g$ without modifying the prior distribution $p(z)$.
+The parameters of $g$ can be learned using training data and an appropiate loss function.\\
+\newpage
+%However, instead of maximizing the log-likelihood of $p(x)$ directly we choose $g$ to emit not samples of $x$ but parameters of a conditional distribution over $x$.
+
+%\subsection{Auto-regressive networks}
 

### 2016-12-19 paper/chapters/sigmoid_belief_nets.tex
updated

@@
 \label{sec:sbn}
-Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a a specific type of bayesian networks \cite{pearl:1985}.
-They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
-
-These random variables are interpreted in a bayesian way where each one may be observed, hidden (or latent) or unknown representing different states of knowledge or belief \cite{definetti:1974}.
 \begin{figure}[htb]
@@
 \input{media/sbn_architecture}
-  \caption{SBN architecture}
+  \caption{Bayesian networks}
   \label{fig:sbn_arch}
@@
   \small
-  Sigmoid Belief Network described as a acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
+  Bayesian networks or belief nets can be described as an acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
   Arrows show stochastic dependency. The dashed arrow indicates that it is possible in the original definition, even though most commonly the hidden variables are divided into layers.
@@
 
+Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
+\ref{fig:sbn_arch} shows a bayesian network as a graphical model with each state influenced by its ancestors.
+In the case of SBNs however, the states are all binary and the activation function is the sigmoid function.
+They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
+%As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
+Each state $s_i$ is a random variables which may be observed, hidden or unknown representing different states of knowledge or belief. As such it can be interpreted with a probability distribution.
+\begin{equation}
+  \label{eq:sbn_node}
+p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
+\end{equation}
+In equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.
+
+In probabilistic networks, we are mostly interested in the aspects: sampling, learning and inferring.
+Even though sampling can be done efficiently, the other operations are intractable in almost all cases.
+\paragraph{Sampling}
+can be done using ancestral sampling through the hidden layers evaluating \ref{eq:sbn_node}
+
+%These random variables are interpreted in a bayesian way where each one may be observed, hidden (or latent) or unknown representing different states of knowledge or belief \cite{definetti:1974}.
+
+\newpage
 \subsection{Learning}
@@
 \end{figure}
+
 In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
-$$
-p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
-$$
 
@@
 There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
+
+\newpage

### 2016-12-19 paper/chapters/vae.tex
moved the cropped version
updated

@@
-\section{Variational Autoencoders \cite{vae:2013}}
+\section{Variational Autoencoders}
 \label{sec:vae}
-Variational Autoencoders (VAE) have been introduced in 2014 by Kingma and Welling which uses variational inference to perform efficient inference and learning in deep latent models.
-Deep latent models are based on the assumption that input data can be explained using non-linear transformations coming from simple distributions. \cite{rezende:2014} (check!?).
+Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma and Welling \cite{vae:2014} and Rezende et al\cite{dlgm:2014}.
+VAEs couple a generator network with an inference network in which efficient inference and learning can be performed, thus allowing the model to be trained purely using gradient-based methods.
+
+%which uses variational inference to perform efficient inference and learning in deep latent models.
+%Deep latent models are based on the assumption that input data can be explained using non-linear transformations coming from simple distributions. \cite{rezende:2014} (check!?).
 
@@
 \end{figure}
+
+\newpage
 
@@
 
+\vspace{10cm}
 % move to architecture:
@@
 $\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+
+%\begin{figure}[htb]
+%\centering
+%\includestandalone[width=\linewidth,mode=buildnew]{media/vae_multiple_layers}
+  %\caption{VAE with multiple stochastic layers }\label{fig:vae_multiple_layers}
+  %\medskip
+  %\small
+  %The left upward arrows indicate the recognition networks $q(h_1|x)$ and $q(h_2|h_1)$.
+  %Note that the recognition networks can have multiple deterministic layers (including non-linearities).\\
+
+  %$p(h_2)$ denotes the probability distribution over the highest(?) latent space.\\
+
+  %On the right, the two conditional distributions $p(h_1|h_2)$ and $p(x|h_2)$ are representing two networks used for the generative process.
+%\end{figure}
+
+\newpage
 
@@
 The prior distribution $p(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
-The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2013} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
+The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2014} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
 Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
@@
 \end{equation}
+
+\newpage
 
@@
         \centering
-        \includegraphics[height=7cm]{media/vae_manifold-crop.pdf}
+        \includegraphics[height=7cm]{media/vae_manifold.pdf}
         \caption{Learned manifold using 2-dimensional latent space}
@@
         \centering
-        \includegraphics[height=7cm]{media/vae_samples-crop.pdf}
+        \includegraphics[height=7cm]{media/vae_samples.pdf}
         \caption{generated images by sampling from latent space and applying $p(x|z)$}
@@
 
-\subsection{Performance}
-\label{sub:vae_performance}
+%\subsection{Performance}
+%\label{sub:vae_performance}
+
+\newpage
 
@@
 
+\newpage
+
 \subsubsection{Importance Weighted Autoencoder \cite{iwae:2015}}
@@
 \input{chapters/vae/draw}
+
+\newpage

### 2016-12-19 paper/chapters/vae/iwae.tex
updated

@@
 
-\begin{figure}[htb]
-\centering
-\includestandalone[width=\linewidth,mode=buildnew]{media/vae_multiple_layers}
-  \caption{VAE with multiple stochastic layers }\label{fig:vae_multiple_layers}
-  \medskip
-  \small
-  The left upward arrows indicate the recognition networks $q(h_1|x)$ and $q(h_2|h_1)$.
-  Note that the recognition networks can have multiple deterministic layers (including non-linearities).\\
-
-  $p(h_2)$ denotes the probability distribution over the highest(?) latent space.\\
-
-  On the right, the two conditional distributions $p(h_1|h_2)$ and $p(x|h_2)$ are representing two networks used for the generative process.
-\end{figure}
 % write about k passes (just averages)
@@
 Of course, using more samples will result in higher computational cost.
+
+\vspace{5cm}

### 2016-12-19 paper/chapters/sigmoid_belief_nets.tex
done some stuff

@@
 \input{media/sbn_architecture}
-  \caption{Bayesian networks}
+  \caption{Conceptual architecture of belief nets}
   \label{fig:sbn_arch}
@@
   \small
-  Bayesian networks or belief nets can be described as an acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
+  Bayesian networks, also called belief nets can be described as an acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
   Arrows show stochastic dependency. The dashed arrow indicates that it is possible in the original definition, even though most commonly the hidden variables are divided into layers.
@@
 Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
-\ref{fig:sbn_arch} shows a bayesian network as a graphical model with each state influenced by its ancestors.
+Figure \ref{fig:sbn_arch} shows a bayesian network as a graphical model with each state influenced by its ancestors.
 In the case of SBNs however, the states are all binary and the activation function is the sigmoid function.
@@
 %As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
-Each state $s_i$ is a random variables which may be observed, hidden or unknown representing different states of knowledge or belief. As such it can be interpreted with a probability distribution.
+Each state $s_i$ is a random variables in the bayesian framework which means it can be observed, hidden or unknown representing different states of knowledge.
+As such it can also be interpreted with a probability distribution.
 \begin{equation}
@@
 
-In probabilistic networks, we are mostly interested in the aspects: sampling, learning and inferring.
-Even though sampling can be done efficiently, the other operations are intractable in almost all cases.
-\paragraph{Sampling}
-can be done using ancestral sampling through the hidden layers evaluating \ref{eq:sbn_node}
+To draw samples from this network, it is possible to simply evaluate equation \ref{eq:sbn_node} for each visible node.
+However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
+Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
+Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\\\
+
+Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
+Especially in deep networks, this kind of computation is infeasible.\\
+
+% DONE
+
+%However in order to draw good samples from this network, the weights need to be adjusted to make it 
+%However most other operations involving this network are intractable for most cases, like learning and performing inference.
+
+%In probabilistic networks, we are mostly interested in the aspects: sampling, learning and inferring.
+%Even though sampling can be done efficiently, the other operations are intractable in almost all cases.
+%\paragraph{Sampling}
+%can be done using ancestral sampling through the hidden layers evaluating \ref{eq:sbn_node}
 
@@
 
-\newpage
-\subsection{Learning}
-\begin{figure}[htb]
-\centering
-\includestandalone[mode=buildnew]{media/sbn_node}
-  \caption{SBN node}
-  \label{fig:sbn_node}
-\end{figure}
+%\newpage
+%\subsection{Learning}
+%\begin{figure}[htb]
+%\centering
+%\includestandalone[mode=buildnew]{media/sbn_node}
+  %\caption{SBN node}
+  %\label{fig:sbn_node}
+%\end{figure}
 
-In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
+%In its most common form, SBNs are comprised by a number of layers where each node in a layer is a sigmoid function of its ancestors:
 
-While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
-% show why?
+%While sampling from these networks is very efficient, computing the inference $p(z|x)$ over the hidden units is intractable.
+%% show why?
 
-There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
+%There have been several attempts on solving the inference problem, using inference networks which have shown promising results.
 

### 2016-12-19 paper/chapters/vae.tex
done some stuff

@@
 \label{sec:vae}
-Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma and Welling \cite{vae:2014} and Rezende et al\cite{dlgm:2014}.
-VAEs couple a generator network with an inference network in which efficient inference and learning can be performed, thus allowing the model to be trained purely using gradient-based methods.
+Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
+The proposed framework includes in addition to an differentiable generator network an inference network in order to perform efficient inference.
+
+Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.
+
+The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
+Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\
+
+However it is not appearent how this model would provide sufficiently good results.
+In order to understand how well the variational autoencoder performs, it is important to understand the variational methods used to build the model.
+
+\begin{figure}[htb]
+\centering
+\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
+  \caption{Graphical model of the variational autoencoder \cite{vae:2014}}
+  \label{fig:vae_architecture}
+\end{figure}
+
+%VAEs couple a generator network with an inference network in which efficient inference and learning can be performed, thus allowing the model to be trained purely using gradient-based methods.
 
@@
 
-\paragraph{Variational Inference (VI)} is a method for approximating intractable posterior distributions over a set of hidden varibales given observations.
-More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are hidden and $x$ observed variables into an optimization problem.
-The term \emph{variational} derives from the calculus of variations, which is a field in mathematical analysis dealing with choosing the best function $q$ from a set of functions $Q$.
+\subsection{Variational Inference}
+The calculus of variations is used to perform optimization on a family of functions.
+As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems.
+
+These intractable posteriors unfortunately arise in almost all interesting models and there are mostly two different approaches to approximate the posterior distribution.
+% MCMC
+Using Markov chain Monte Carlo (MCMC) methods, which construct a Markov chain with the desired attributes.
+The advantage of this approach is that is guaranteed to asympotically approach the desired posterior, however it is difficult to reason about the convergence rate.
+For deep neural networks, running MCMC in the inner loop of optimization techniques is impractical.\\
+% variational inference
+The other approach which is using variational inference which picks a specific distribution restricted to a family of distributions matching the posterior most closely.
+Advantages using this approach rather than MCMC is the greater speed at which the computation is performed.
+However it is not guaranteed to even approximate the posterior accurately, if the true posterior can not be expressed as a distribution the family.
+
+More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are the hidden and $x$ the observed variables into an optimization problem, where we can use gradient-based methods.
+
+\subsubsection{Kullback-Leibler Divergence}
+In order to approximate one distribution $Q$ to another distribution $P$, a measurement of difference.
+One commonly used measurement is called Kullback-Leibler divergence (short KL-divergence) which enables the variational inference methods to choose appropiate distributions.
+%\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
+$\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
+For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} as:
+\begin{equation}
+  \mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
+\end{equation}
+
+Note that divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
+$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the variational inference.
+
+%The need for such optimizations techniques
+
+%is suited to approximate intractable posterior distributions over a set of hidden variables given observations.
+%Variational-based methods have become a popular choice for inference problems, 
+%Variational Inference (VI) is a method for approximating intractable posterior distributions over a set of hidden varibales given observations.
+
+%The term \emph{variational} derives from the calculus of variations, which is a field in mathematical analysis dealing with choosing the best function $q$ from a set of functions $Q$.
 In the case of variational inference, this $Q$ denotes a class of distributions and the variational parameters are the parameters which are used to define $q \in Q$.
-VI is a alternative to markov chain monte carlo (MCMC) methods like gibbs sampling, due to lower computational cost.
-However, there is no guarantee that VI methods will asymptotically reach the optimimum in contrast to MCMC.
+%VI is a alternative to markov chain monte carlo (MCMC) methods like gibbs sampling, due to lower computational cost.
+%However, there is no guarantee that VI methods will asymptotically reach the optimimum in contrast to MCMC.
 
-Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling in practice(?).
+%Given this assumption, building a good representation in the latent space is crucial for minimal reconstruction error as well as good generative modelling in practice(?).
 
 % TODO: add clarification that q/p are deep nets!
-\begin{figure}[htb]
-\centering
-\includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
-  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
-\end{figure}
 
-\newpage
+%\newpage
 
-\paragraph{Formal Setup}
-Formally, $p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.
-Inferring the posterior distribution $p(z|x)$ is particularly interesting, because it means to enable learning parameters for good latent space representation.
-$p(z|x)$ can be expanded using bayes rule to $\frac{p(x|z) p(z)}{p(x)}$.
-Computing the nominator is straightforward, $p(z)$ is the probability distribution we chose for the latent space, oftenmost simply a gaussian.
-$p(x|z)$ is easy to compute as well, ...? (forward pass?).\\
-Meanwhile is the denominator difficult to evaluate, because it requires to consider all possible input combinations.
-What we do instead in the variational autoencoders is to derive a lower-bound on $p(z|x)$ using a auxiliary distribution called $q_\theta(z|x)$.
-The subscript $\theta$ indicates that the distribution is parametrized by an variational term, so that for one $\theta$, $q_\theta(z|x) \approxeq p(z|x)$ holds ($q(z|x)$ doesn't need to depend on $x$).
 %Performing statistical inference means computing $p(z|x)$ which is intractable in all but very simple cases (integrating over it, show calc).
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
   \centering
-\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
-  \caption{VAE architecture (source: VAE paper)}\label{fig:vae_architecture}
+  \includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
+  \caption{simplified VAE architecture}
+  \label{fig:vae_architecture}
 \end{figure}
+
+$p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.
+Inferring the posterior distribution $p(z|x)$ is particularly interesting, because it means to enable learning parameters for good latent space representation.
+$p(z|x)$ can be expanded using bayes rule to $\frac{p(x|z) p(z)}{p(x)}$.
+Computing the nominator is straightforward, $p(z)$ is the probability distribution we chose for the latent space, oftenmost simply a gaussian.
+$p(x|z)$ is easy to compute as well, ...? (forward pass?).\\
+Meanwhile is the denominator difficult to evaluate, because it requires to consider all possible input combinations.
+What we do instead in the variational autoencoders is to derive a lower-bound on $p(z|x)$ using a auxiliary distribution called $q_\theta(z|x)$.
+The subscript $\theta$ indicates that the distribution is parametrized by an variational term, so that for one $\theta$, $q_\theta(z|x) \approxeq p(z|x)$ holds ($q(z|x)$ doesn't need to depend on $x$).
+
 VAE just like other autoencoders encode the input data into a latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
@@
 %One of the main differences is that the VAE framework enforces a specific prior distribution on the latent space, most of the time simply an isotropic gaussian.
-
-\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
-$\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
-The KL divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
-$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the VAE to approximate the true posterior distribution.
 
@@
 
-\vspace{10cm}
+%\vspace{10cm}
 % move to architecture:
@@
 %\end{figure}
-
-\newpage
 
@@
 
+\begin{algorithm}
+  \caption{Learning in the VAE model}
+  \label{alg:vae}
+  \begin{algorithmic}[1]
+    %\Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
+    %\Let{$\theta_d, \theta_g$}{Parameters of function approximators}
+    \For{Until convergence}
+      \Let{$x$}{sample from $p_{data}(x)$}
+      \Let{$\epsilon$}{Random samples from noise prior $p(\epsilon)$}
+      \Let{$g$}{Compute gradients $\nabla_{\theta,\phi}\mathcal{L}(\theta, \phi; x, \epsilon)$}
+      \Let{$\theta, \phi$}{Update according to gradients $g$}
+    \EndFor
+  \end{algorithmic}
+\end{algorithm}
+
 % define X correctly or replace formula below!
@@
 %Because we assume that $p(z)$ is also a multivariate Gaussian distribution, the KL-divergence can now be computed in closed form as follows\cite{derivations:2007}.
-\begin{align*}
-  \mathcal{D}_{\mathrm{KL}}\big[q_\theta(z|x) || p_\phi(z|x)\big] &= \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]\\
-  &= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)\\
-\end{align*}
+%\begin{align*}
+  %\mathcal{D}_{\mathrm{KL}}\big[q_\theta(z|x) || p_\phi(z|x)\big] &= \mathcal{D}_{\mathrm{KL}}\big[\mathcal{N}(\mu_0,\Sigma_0) || \mathcal{N}(\mu_1,\Sigma_1)\big]\\
+  %&= \frac{1}{2} \big(\mathrm{tr}\big(\Sigma_1^{-1}\Sigma_0\big) + \big(\big)\big)\\
+%\end{align*}
 

### 2016-12-20 paper/chapters/vae.tex
added

@@
 Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
-The proposed framework includes in addition to an differentiable generator network an inference network in order to perform efficient inference.
-
-Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.
-
-The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
-Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\
+The proposed framework includes in addition to a differentiable generator network an inference network with which the model is able to efficiently approximate the posterior distribution.
+Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.\\
+%The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
+%Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\
 
 However it is not appearent how this model would provide sufficiently good results.
-In order to understand how well the variational autoencoder performs, it is important to understand the variational methods used to build the model.
+In order to understand how well the variational autoencoder performs, it is important to first understand the variational methods used to build the model.
 
@@
 
+\paragraph{Inference problem}
+Given a dataset $X = { x^{(1)}, x^{(2)}, \dots, x^{(n)}}$ and a model with observations $x \in X$ and hidden variables $z$.
+The inference problem is posed as computing the posterior distribution $p(z|x)$
+\begin{equation}
+  \label{eq:intractable_posterior}
+  p(z|x) = \frac{p(x|z) p(z)}{p(x)}
+\end{equation}
+However, computing the right-hand side of equation \ref{eq:intractable_posterior} is intractable due to the exponentially many configurations of the hidden variables.
+As such it is often needed to use approximate inference such as variational inference.
+
 \subsection{Variational Inference}
-The calculus of variations is used to perform optimization on a family of functions.
-As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems.
+Variational inference (VI) can be used to turn the inference problem into an optmization problem which we are able to approximate using gradient-based methods.
+Approximating a posterior distribution 
+%The calculus of variations is used to perform optimization on a family of functions.
+%As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems as seen in \ref{eq:intractable_posterior}.
 

### 2016-12-20 paper/chapters/gan.tex
done some stuff
zZzZzzz almost done

@@
 %%% when introduced, game theory approach
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al based on a game theoretic approach to generative modelling.
+Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
 %which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
 
-The idea is to have two neural networks playing against each other.
-A generator $G(z)$ and a discriminator $D$ where the generator tries to produce data similar to those encountered in a training set.
-$D$ receives samples from either the training set or the generated data, without knowing where it came from.
-Both $G$ and $D$ are usually implemented as neural networks with $G$ being trained to output samples fooling the discriminator and $D$ distinguishing generated samples from training examples.
+The idea is to have two players playing a non-cooperative game against each other.
+One player is called the generator $G$ tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually originating from a dataset.
+It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.
+
+Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
+On the other side, $G$ tries to produce as many samples as possibly which got accepted as real by $D$.\\
+
+As the game progresses, both parties will get better at their respective objective and the generated samples will look more and more like the training examples.\\
+
+Using the analogy of counterfeiters and police can help understand the intuition behind GANs better.
+Counterfeiters will try to produce fake money looking to fool the police into thinking it is real money.
+The police however is able to distinguish badly made fake money from real money easily and as such rejects the fake money with high confidence.
+This rejection in turn triggers the counterfeiters to look at real money more closely and try to imitate it even better.
+Therefore they are able to improve the quality of the faked money.
+Improvements made to the fake money in this iteration will lower the confidence with which the police was able to reject previous samples.
+
+The game-theoretic solution to the emerging situation where neither counterfeiters as generator nor police as discriminator can improve its own success with respect to the outcome of the game gets called \emph{Nash equilibrium}\cite{game_theory:1994}.
+
+In the case of GANs, this situation means the generator produces samples which can only be rejected with probability $p=0.5$ by the discriminator, meaning it produces as good samples as there are in the dataset.\\\\
+
+To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.
+
+In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
+GANs uses neural network as function approximator for $G$ as well as $D$.
+The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
+where $\xi$ denotes the parameters learned by the neural network.\\
+Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.
 
 % use of words: forgery
-To understand the inuition behind the approach to GANs, it may be helpful to view(?) analogy of counterfeiters and the police.
-Counterfeiters try to produce fake money imitating legit money while the police tries to distinguish fake from real money.
-This procedure of generating and distinguishing will go back-and-forth as the counterfeiters will get better at producing authentic looking money and the police is hard-pressed to differentiate between fake and real.
-Eventually, this procedure will result in a nash equlibrium where the police or discriminator will have 50\% chance of seperating authentic from real.
+%To understand the inuition behind the approach to GANs, it may be helpful to view(?) analogy of counterfeiters and the police.
+%Counterfeiters try to produce fake money imitating legit money while the police tries to distinguish fake from real money.
+%This procedure of generating and distinguishing will go back-and-forth as the counterfeiters will get better at producing authentic looking money and the police is hard-pressed to differentiate between fake and real.
+%Eventually, this procedure will result in a nash equlibrium where the police or discriminator will have 50\% chance of seperating authentic from real.
 
-Ideally, we want to find this equilibrium for the GAN model as well.
-
-\newpage
+%Ideally, we want to find this equilibrium for the GAN model as well.
 
@@
 
-\subsubsection{Minibatch discrimination}
-\label{ssub:gan_minibatch_discrimination}
-
-\vspace{10cm}
-
-\subsubsection{Historical Averaging}
-\label{ssub:gan_historical_averaging}
-
-\newpage
-
 %\subsection{Application}

### 2016-12-20 paper/chapters/introduction.tex
done some stuff
zZzZzzz almost done

@@
 Directed generative models are a wide class of models used for generating samples from an unknown distribution.
-Lets denote $p_{data}(x)$ as the real-world data distribution which is unknown, however we have access to samples from it: $x \sim p_{data}(x)$.
+Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
 In order to do so, we define an approximated distribution $p_{model}$ from which we are able to sample.
@@
 %% on building reasonable representations
-The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
+%The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
+Models using deep learning often assume that observed data is stochastically dependent on some hidden structure which we wish to explain.
 In order to understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real world data.
@@
   \item \textbf{Audio data} is omni-present in the real world, however most data sources can be transformed from raw audio samples to spoken words or musical notes without losing too much information for reconstruction.
-  \item \textbf{...}
+  %\item \textbf{...}
 \end{itemize}
@@
 
-The assumption that $p(x)$ is dependent on some underlying structure can now be expressed as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.
+The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in figure \ref{fig:dgm}.
+%as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.
 This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\
@@
 
+\newpage
 %Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.

### 2016-12-20 paper/chapters/sigmoid_belief_nets.tex
zZzZzzz almost done

@@
 Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
-Figure \ref{fig:sbn_arch} shows a bayesian network as a graphical model with each state influenced by its ancestors.
-In the case of SBNs however, the states are all binary and the activation function is the sigmoid function.
+Figure \ref{fig:sbn_arch} shows a generic belief network as a graphical model with each state influenced by its ancestors.
+In the case of SBNs however, all states are binary and the activation function is the sigmoid function.
 They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
 %As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
-Each state $s_i$ is a random variables in the bayesian framework which means it can be observed, hidden or unknown representing different states of knowledge.
+Each state $s_i$ is a random variable in the bayesian framework which means it can be observed, hidden or unknown representing different states of knowledge.
 As such it can also be interpreted with a probability distribution.
@@
 Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
-Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\\\
+Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\
 
 Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
-Especially in deep networks, this kind of computation is infeasible.\\
+Especially in deep networks, this kind of computation is infeasible as it requires evaluating exponentially many configurations.\\
 

### 2016-12-20 paper/chapters/vae.tex
zZzZzzz almost done

@@
 \label{sec:vae}
-Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
-The proposed framework includes in addition to a differentiable generator network an inference network with which the model is able to efficiently approximate the posterior distribution.
-Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.\\
+Variational autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
+%The proposed framework resembles the encoder-decoder architecture, using a probabilistic encoder and decoder.
+The general idea of VAEs is to have an encoder-decoder architecture where inputs $x$ gets mapped to latent variables $z$ and from there to the reconstruction $~x$.
+
+%\begin{figure}[htb]
+%\centering
+%\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
+  %\caption{Graphical model of the variational autoencoder \cite{vae:2014}}
+  %\label{fig:vae_architecture}
+%\end{figure}
+
+However, in addition to this autoencoder structure an additional regularizer forces the latent space to have a specific simple form.
+To achieve this, variational-based methods are used to output a distribution on the latent space instead of directly emitting samples.
+
+The probabilistic encoder generally implemented as feed-forward network receives input from the dataset and outputs parameters of a distribution.
+For example the parameters for a gaussian distribution with diagonal covariance, in which case the networks emits $(\mu,\Sigma)$.\\
+
+Using gradient-based optimization techniques enables the network to be trained to be similar to the true posterior $p(z|x)$.
+In addition to this inference network there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.
+
+One useful aspect of this approach is the possibility to train both networks using error backpropagation due to the use of neural networks as function approximators.\\\\
+
+Formally, the described probabilitic encoder emits variational parameters $\phi$ used to define the probability distribution $q_\phi(z|x)$.
+Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.\\
+
+\begin{figure}[h]
+  \centering
+  \includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
+  \caption{simplified VAE architecture using a gaussian prior on $z$}
+  \label{fig:vae_architecture}
+\end{figure}
+
+The key insight on how the variational autoencoder is able to optimize $q_\phi(z|x)$ to approach the true data posterior distribution $p(z|x)$ is in defining a variational lower bound which is used to push the approximated $q_\phi(z|x)$ towards $p(z|x)$.
+We will discuss this lower bound $\mathcal{L}$ in the following section in detail.
+
+\newpage
+
+%The approach taken by VAE is instead of sampling directly to a latent space $z$ it uses a inference network to output parameters of an distribution on $z$.
+
+%One of the useful aspects using this approach is that the inference problem of computing $p(z|x)$ is posed as optimizing the inference network to output good values for distributions approximating $p(z|x)$.
+%The model maps the input $x$ using an inference network to a distribution $q_\theta(z|x)$ while a generative network maps samples from the latent space $z$ to to the data space $~x \
+
+%Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.\\
 %The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
 %Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\
-
-However it is not appearent how this model would provide sufficiently good results.
-In order to understand how well the variational autoencoder performs, it is important to first understand the variational methods used to build the model.
-
-\begin{figure}[htb]
-\centering
-\resizebox{5cm}{!}{\input{media/vae_architecture.tikz}}
-  \caption{Graphical model of the variational autoencoder \cite{vae:2014}}
-  \label{fig:vae_architecture}
-\end{figure}
 
@@
 
-\paragraph{Inference problem}
-Given a dataset $X = { x^{(1)}, x^{(2)}, \dots, x^{(n)}}$ and a model with observations $x \in X$ and hidden variables $z$.
+%\paragraph{Inference problem}
+%The problem variational-based methods can solve is approximating intractable posterior distributions.\\
+
+%Given a dataset $X = { x^{(1)}, x^{(2)}, \dots, x^{(n)}}$ and a model with observations $x \in X$ and hidden variables $z$.
 The inference problem is posed as computing the posterior distribution $p(z|x)$
-\begin{equation}
-  \label{eq:intractable_posterior}
-  p(z|x) = \frac{p(x|z) p(z)}{p(x)}
-\end{equation}
-However, computing the right-hand side of equation \ref{eq:intractable_posterior} is intractable due to the exponentially many configurations of the hidden variables.
+%\begin{equation}
+  %\label{eq:intractable_posterior}
+  %p(z|x) = \frac{p(x|z) p(z)}{p(x)}
+%\end{equation}
+%However, computing the right-hand side of equation \ref{eq:intractable_posterior} is intractable due to the exponentially many configurations of the hidden variables.
 As such it is often needed to use approximate inference such as variational inference.
@@
 Variational inference (VI) can be used to turn the inference problem into an optmization problem which we are able to approximate using gradient-based methods.
-Approximating a posterior distribution 
 %The calculus of variations is used to perform optimization on a family of functions.
@@
 $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the variational inference.
+
+\newpage
 
@@
 
-\subsection{Architecture}
-\label{sub:vae_architecture}
+%\subsection{Architecture}
+%\label{sub:vae_architecture}
 
-\begin{figure}[t]
-  \centering
-  \includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
-  \caption{simplified VAE architecture}
-  \label{fig:vae_architecture}
-\end{figure}
-
-$p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.
-Inferring the posterior distribution $p(z|x)$ is particularly interesting, because it means to enable learning parameters for good latent space representation.
-$p(z|x)$ can be expanded using bayes rule to $\frac{p(x|z) p(z)}{p(x)}$.
-Computing the nominator is straightforward, $p(z)$ is the probability distribution we chose for the latent space, oftenmost simply a gaussian.
-$p(x|z)$ is easy to compute as well, ...? (forward pass?).\\
-Meanwhile is the denominator difficult to evaluate, because it requires to consider all possible input combinations.
-What we do instead in the variational autoencoders is to derive a lower-bound on $p(z|x)$ using a auxiliary distribution called $q_\theta(z|x)$.
+%$p(x,z)$ is the joint probability distribution over both input and latent variables while $p(z|x)$ is the conditional probability of the latent variables given input data.
+%Inferring the posterior distribution $p(z|x)$ is particularly interesting, because it means to enable learning parameters for good latent space representation.
+%$p(z|x)$ can be expanded using bayes rule to $\frac{p(x|z) p(z)}{p(x)}$.
+%Computing the nominator is straightforward, $p(z)$ is the probability distribution we chose for the latent space, oftenmost simply a gaussian.
+%$p(x|z)$ is easy to compute as well, ...? (forward pass?).\\
+%Meanwhile is the denominator difficult to evaluate, because it requires to consider all possible input combinations.
+%What we do instead in the variational autoencoders is to derive a lower-bound on $p(z|x)$ using a auxiliary distribution called $q_\theta(z|x)$.
 The subscript $\theta$ indicates that the distribution is parametrized by an variational term, so that for one $\theta$, $q_\theta(z|x) \approxeq p(z|x)$ holds ($q(z|x)$ doesn't need to depend on $x$).
 
-VAE just like other autoencoders encode the input data into a latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
-But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.
-This allows to sample from this distribution and generate output which will look similar to the data on which the VAE has been trained.
-\ref{fig:vae_architecture}.
+%VAE just like other autoencoders encode the input data into a latent space similar to compression of data and is able to decode a vector of latent variables into output while trying to match the output to the input.
+%But in contrast to other autoencoders (sparse, denoising), we enforce a specific distribution on the latent space.
+%This allows to sample from this distribution and generate output which will look similar to the data on which the VAE has been trained.
+%\ref{fig:vae_architecture}.
 
-\paragraph{Relation to Auto-Encoders}
-The variational autoencoder has the same encoding-decoding architecture as other autoencoders, for example sparse or denoising autoencoders.
-But in contrast to these autoencoders, the VAE enforces a probability distribution on the learned latent space $z$.
-In particular, the framework encourages the model to learn a representation that is close to $p(z)$ (which usually is an isotropic multivariate gaussian distribution) by including the KL divergence in the objective function.
+%\paragraph{Relation to Auto-Encoders}
+%The variational autoencoder has the same encoding-decoding architecture as other autoencoders, for example sparse or denoising autoencoders.
+%But in contrast to these autoencoders, the VAE enforces a probability distribution on the learned latent space $z$.
+%In particular, the framework encourages the model to learn a representation that is close to $p(z)$ (which usually is an isotropic multivariate gaussian distribution) by including the KL divergence in the objective function.
 %Similar to other autoencoder frameworks VAE provides an encoder as well as a decoder model, but in contrast to other autoencoders both networks are probabilistic.
@@
 \subsection{Objective Function}
-The objective function of the VAE consists of the reconstruction error and the regularizer.
+%The objective function of the VAE consists of the reconstruction error and the regularizer.
 %$\mathcal{L}$ learns both the encoder as well as the decoder with their respective parameters $\phi$ and $\theta$.
@@
 % probabilistic decoder $p_\theta(x|z)$
-We will discuss both terms in detail below following the derivation of the lower variational bound $\mathcal{L}$ and the rewritten objective function.
+%We will discuss both terms in detail below following the derivation of the lower variational bound $\mathcal{L}$ and the rewritten objective function.
+Following the thoughts earlier, we wish to derivate an objective function which can be used to perform gradient-based optimization in the model.\\
+We construct the approximate posterior $q_\phi(z|x)$ which we want to approach the true data posterior $p(z|x)$.
+When using the KL divergence as measurement for difference between two distributions the goal for which we want to optimize becomes the following.
+\begin{equation}
+  q_\phi^{*}(z|x) = \arg\min_{\phi} \mathcal{D}_{\mathrm{KL}} \big(q_\phi(z|x) || p(z|x)\big)
+\end{equation}
 
-\paragraph{Derivation of lower bound}
+However we are not able to optimize for that goal directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
+The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
+\begin{equation}
+  \label{eq:integral}
+p(x) = \int p(x,z) p(z) dz
+\end{equation}
+
+Computing the integral in equation \ref{eq:integral} requires the evaluation of every possible configuration over the hidden variables $z$ and therefore intractable.\\\\
+
+Instead of directly optimizing for the true posterior, we construct an evidence lower bound (ELBO) $\mathcal{L}$ which we can maximize which is equivalent to minimizing the KL divergence between $q_\phi(z|x)$ and $p(z|x)$.\\
+
+\paragraph{Derivation of the evidence lower bound}
 \input{chapters/vae/derivation_loss}
 
-\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative Kullback-Leibler (KL) divergence between the learned variational distribution $q_\theta(z|x)$ and the latent space prior distribution $p(z)$.
-The regularizer also prevents the model from collapsing into a single point.
+\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative KL divergence between the learned variational distribution $q_\phi(z|x)$ and the latent space prior distribution $p(z)$.
 
-%\vspace{10cm}
-% move to architecture:
-%In the original VAE paper, the unit gaussian distribution with diagonal variance was proposed for $p(z)$ and although there have been multiple proposals (citations, multi-modal etc) for different distributions we will stick with $\mathcal{N}(0,I)$ for the sake of simplicity.
-% ---
+%The regularizer is also important to the VAE as it ensures that the model learns generalized representations in latent space.
 
-%For two probability distributions $P$ and $Q$, the KL divergence $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can informally be described as the amount of information which is lost when using $Q$ to represent $P$.
+With the example using handwritten digits it appears why this regularizer is important for the model to learn generalized representations in latent space.
+Suppose we have two data points from the dataset which both represent the same handwritten digit.
+Ideally, our model would learn the representations in latent space to be fairly close to ensure the model has learned to generalize the semantics.
+Without $\mathcal{L}_{reg}$ the model would assign each data example a different representation in latent space with a distribution collapsing to that single point.
+This behaviour is undesired and is as such penalitized.
+
+%Suppose two data points with similar input
+%Without $\mathcal{L}_{reg}$ the encoder would encode data points we want to have similar representations, fo
+
+%The regularizer term is also important to prevent the model from collapsing into a single point.
+
+\vspace{2cm}
 
 \paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
-This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\phi(z | x_i)$.
+This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\phi(z | x_i)$.\\
 $\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+
+\newpage
 
@@
 \subsubsection{Reparameterization Trick}
-
-% more math:
-% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
-To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
-Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
-Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
-\begin{equation}
-  \label{eq:rep_trick}
-  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
-\end{equation}
-This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
-
-The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
-In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
 
@@
   Rectangular nodes represent deterministic functions while circles indicate stochastic variables.
-  Note that backpropagation of gradients through stochastic nodes is problematic and adds bias (verify!)
+  %Note that backpropagation of gradients through stochastic nodes is problematic and adds bias (verify!)
 \end{figure}
+
+% more math:
+% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
+To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
+Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
+Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
+\begin{equation}
+  \label{eq:rep_trick}
+  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
+\end{equation}
+This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
+
+The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
+In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
 
@@
 \label{sub:vae_learning}
-The probabilistic encoder $q_\theta(z|x)$ is commonly implemented using a deep neural network which transforms the input data $x$ into a multivariate gaussian distribution with mean $\mu_0$ and diagonal variance $\Sigma_0$.
-To allow the computation of the KL divergence between $q_\theta(z|x)$ and the latent prior distribution $p(z)$ in closed form, the prior is an isotropic gaussian $\mathcal{N}(0, 1)$.
+The probabilistic encoder $q_\phi(z|x)$ and decoder $p_\theta(x|z)$ are commonly implemented using a deep neural network.
+To allows analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.
+
+%To allow the computation of the KL divergence between $q_\theta(z|x)$ and the latent prior distribution $p(z)$ in closed form, the prior is an isotropic gaussian $\mathcal{N}(0, 1)$.
 
@@
     %\Let{$\theta_d, \theta_g$}{Parameters of function approximators}
-    \For{Until convergence}
+    \While{has not reached convergence}
       \Let{$x$}{sample from $p_{data}(x)$}
@@
       \Let{$\theta, \phi$}{Update according to gradients $g$}
-    \EndFor
+    \EndWhile
   \end{algorithmic}
 \end{algorithm}
+
+\subsection{Experiments}
+In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}.
+Results of this experiments are shown in fig.
+\ref{fig:vae_experiments}.
+%The model has been trained on the MNIST dataset \cite{mnist} with satisfactory generative performance as shown in figure \ref{fig:vae_experiments}
 
@@
 
-\subsection{Inference}
-\label{sub:vae_inference}
-In bayesian analysis, statistical inference is the process of evaluating $p(z|x)$ for a given conditional probability $p(x|z)$.
-% more info on marginal distribution, posterior inference etc!
-There are two approaches to performing this task, Markov chain Monte Carlo (MCMC) methods using Gibbs sampling and variational inference.
-MCMC methods have the advantage of being asympotically exact and unbiased while variational methods are computationally faster but biased and not guaranteed to approaching the distribution.
-In the case of VAE, variational inference is also used in order to backpropagate through the layers (using the reparameterization trick).
+%\subsection{Inference}
+%\label{sub:vae_inference}
+%In bayesian analysis, statistical inference is the process of evaluating $p(z|x)$ for a given conditional probability $p(x|z)$.
+%% more info on marginal distribution, posterior inference etc!
+%There are two approaches to performing this task, Markov chain Monte Carlo (MCMC) methods using Gibbs sampling and variational inference.
+%MCMC methods have the advantage of being asympotically exact and unbiased while variational methods are computationally faster but biased and not guaranteed to approaching the distribution.
+%In the case of VAE, variational inference is also used in order to backpropagate through the layers (using the reparameterization trick).
 
@@
         \includegraphics[height=7cm]{media/vae_samples.pdf}
-        \caption{generated images by sampling from latent space and applying $p(x|z)$}
+        \caption{Generated images by sampling from latent space}
     \end{subfigure}
     \caption{Variational autoencoder learning performance}
+    \label{fig:vae_experiments}
+    \medskip
+    \small
+    (a) Evaluation of a variational autoencoder with 2 hidden layers with each 500 nodes and a two-dimensional latent space.
+    The shown manifold has been plotted using samples uniformly samples in the range $[-3,-3]^2$ and then reconstructed using the decoder network $p_\theta(x|z)$.
+The results have shown that the model has learned to arrange the latent space in a sensible way by placing digits with high similarity near each other.\\
+    (b) generative performance qualitively evaluated using samples in the data space by random sampling $z \sim p(z)$ and decoding $z$ using the decoder network.
+For this experiment an 50-dimensional latent space has been used.
 \end{figure*}
@@
 
-\subsubsection{Conditional VAE \cite{cvae:2015}}
+\subsubsection{Conditional VAE \cite{cvae:2014} \cite{cvae_2:2015}}
 \label{ssub:vae_conditional_vae}
@@
 \newpage
+
+\subsubsection{Deep Latent Gaussian Model}
+%\begin{figure*}[t!]
+    %\centering
+    %\includegraphics[height=7cm]{media/dlgm_computational_graph-crop.pdf}
+    %\caption{VAE with multiple stochastic layers as a deep latent gaussian model \cite{dlgm:2014}}
+    %\label{fig:dlgm_model}
+    %\medskip
+    %\small
+    %This figure shows the computational graph in a deep latent gaussian model. Red arrows indicate the forward pass and black arrows the backward pass. Meanwhile represent j
+%\end{figure*}
+Up until now the presented variational autoencoder only has one stochastic latent layer.
+However in the original paper by Rezende et al.
+\cite{dlgm:2014} the authors have shown a variational autoencoder using multiple stochastic layers with gaussian priors in each layer.
+Note that the deterministic non-linear transformations between each layer still can be represented using deep neural networks.
+This enables the model to learn richer, more hierarchical structures captured in each latent layer.
+
+%Up until now, the variational autoencoder has only included one stochastic layer. It is however possible to extend the architecture to support multiple latent spaces.
+
+%\cite{dlgm:2014} have presented a model for learning through multiple stochastic layers which can be easily applied to the VAE framework as shown in \ref{fig:vae_multiple_layers}.
 

### 2016-12-20 paper/chapters/vae/cvae.tex
zZzZzzz almost done

@@
-Instead of randomly generated data, generation of output conditioned on some prior information is needed.
-A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks
+Instead of using VAEs in a completely unsupervised manner, oftentimes learning from a dataset where a small subset of datapoints has corresponding labels.
+Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly.\\
 
-Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.
+To adjust the VAE model to the semi-supervised case, we have to add an random variable $y$ corresponding to the corresponding label if available.
+Thus we can extend the distributions as shown below.
 
-To support conditional distributions, the following relations have to be rewritten:
+%The modified setup of the semi-supervised case for the variational autoencoder can be formalized as follows.
+%Dataset $X$ of size $N$ with observations $x_i$ with $i<N$ where for some small subset of observations $x_i$ the corresponding labels $y_i$ are known.
+%We denote the labelled subset as $~p_l(x,y)$ and the unlabelled dataset as $~p_u(x,y)$.
+%\begin{equation}
+
+%\end{equation}
+%Instead of randomly generated data, generation of output conditioned on some prior information is needed.
+%A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks
+
+%Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.
+
+%To support conditional distributions, the following relations have to be rewritten:
 \begin{itemize}
@@
 \end{itemize}
+
+These modifications allows us to formalize the new objective functionas shown in equations \ref{eq:cvae_labels} and \ref{eq:cvae_unknown}.
+
+For datapoints $x$ where labels $y$ are available.
+\begin{align}
+  \label{eq:cvae_labels}
+  \log p_\theta(x) &\geq \mathbb{E}_{q_\phi(z|x,y)} \bigg[log p_\theta(x|y,z) + \log p_\theta(y) + \log p(z) - \log q_\phi(z|x,y)\bigg]
+\end{align}
+
+And for the unsupervised case where the label $y$ corresponds to an unknown entity, we treat $y$ as an additional latent variable.
+\begin{align}
+  \label{eq:cvae_unknown}
+  \log p_\theta(x) &\geq \mathbb{E}_{q_\phi(y,z|x)} \bigg[\log p_\theta(x|y,z) + \log p_\theta(y) + \log p(z) - \log q_\phi(y,z|x)\bigg]
+\end{align}
+
+The final objective function for whole dataset is the summation between right-hand terms of equations \ref{eq:cvae_labels} for the labelled subset and \ref{eq:cvae_unknown} for the unlabelled subset.

### 2016-12-20 paper/chapters/vae/draw.tex
zZzZzzz almost done

@@
-The \emph{deep recurrent attention writer}, in short DRAW is an advanced model uses the selective attention mechanism in addition to a VAE with both recurrent encoder and recurrent decoder.
-DRAW allows to basically the pixels of an image sequentially very human-like.
+The \emph{deep recurrent attention writer}, in short DRAW \cite{draw:2015} is an advanced extension to the variational autoencoder.
+It uses recurrent neural networks as encoder as well as decoder and includes a selective attention mechanism.
+These advancements allow the model to observe an image through a series of \emph{glimpses} and draw to an output region determined by the attention mechanism.
+Without going into much detail, this allows the model to sequentially draw images in a way which resembles much of human-like drawing or sketching.
+
+%It includes recurrent neural networks and a selective attention mechanism to
+
+%uses the selective attention mechanism in addition to a VAE with both recurrent encoder and recurrent decoder.
+%DRAW allows to basically the pixels of an image sequentially very human-like.

### 2016-12-20 paper/chapters/gan.tex
more stuff on GANS

@@
 %%% when introduced, game theory approach
-Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
+Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
 %which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
 
-The idea is to have two players playing a non-cooperative game against each other.
-One player is called the generator $G$ tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually originating from a dataset.
+GANs circumvent the issues arising the inference problem by playing a two-player game.
+Both players play non-cooperatively against each other, which means they only try to maximize their respective success or profit.
+One player is called the generator $G$ which tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually real data.
 It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.
@@
 Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
-On the other side, $G$ tries to produce as many samples as possibly which got accepted as real by $D$.\\
+On the other side, $G$ tries to produce as many real-looking samples as possible to get accepted by $D$.\\
 
@@
 
-To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.
-
-In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
-GANs uses neural network as function approximator for $G$ as well as $D$.
-The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
-where $\xi$ denotes the parameters learned by the neural network.\\
-Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.
+Another formulation of the GAN framework is to define the zero-sum game between the generative model and its adversary.
+The payoff of the discriminator can be described as the value function $v(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-v(G,D)$.\\
+This formulation allows us to define the convergence as equation \ref{zerosum_convergence}.
+\begin{equation}
+  \label{eq:zerosum_convergence}
+  g^* = \arg \min_{G} \max_{D} v(g,d)
+\end{equation}
 
@@
 \centering
-\includestandalone[width=\linewidth,mode=buildnew]{media/gan_conceptual}
-  \caption{Conceptual figure of generative adversarial network}
+\includestandalone[width=\textwidth,mode=buildnew]{media/gan_conceptual}
+  \caption{Architecture used for generative adversarial networks}
   \label{fig:gan_a}
@@
   \small
-  The above figure is illustrating the information and gradient flow in the GAN framework.\\
-  \emph{data space} denotes the real-world data sampled from $p_{data}$, which we aim to model.
-  Dashed arrows denote the gradient flow.
+  The above figure illustrates the computational graph and gradient flow in the GAN framework.\\
+  $X$ denotes the dataset containing real-world samples from $p_{data}$, which we aim to model.\\
+  Solid arrows indicate the forward pass in the model while dashed arrows denote the gradient flow during a backward pass.
 \end{figure}
@@
 \label{sub:gan_arch}
-Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
+To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.
+
+In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
+GANs uses neural network as function approximator for $G$ as well as $D$.
+The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
+where $\xi$ denotes the parameters learned by the neural network representing $G$.\\
+Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.\\
+
+Figure \ref{fig:gan_a} summarizes the described architecture.\\
+%Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
 
@@
 %$$
-$$
-\mathcal{L}_g = \min \log(1 - D(G(z)))\\
+\begin{equation}
+  \label{eq:l_g_1}
+  \mathcal{L}_g = \min \log(1 - D(G(z)))
+\end{equation}
+
+However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning where the discriminator is able to reject samples from the generator with high confidence.\\
+Therefore is in practice often equation \ref{eq:l_g_2} instead of equation \ref{eq:l_g_1} to optimize its objective.\\
+\begin{equation}
+  \label{eq:l_g_2}
   \mathcal{L'}_g = \max \log D(G(z))
-$$
-This formulation has the benefit of the same optimum fixed point, but has a better gradient flow early in training when $G$ produces samples which can be distinguished easily from real data by $D$.
+\end{equation}
 
@@
 
-\newpage
-
 \subsection{Training}
 \label{sub:gan_training}
-Combining the loss functions of both generator and discriminator yields us the following objective with value function $V(G,D)$.
+\begin{figure}[htb]
+\centering
+  \includegraphics[width=\linewidth]{media/gan_training-crop}
+  \caption{Architecture used for generative adversarial networks}
+  \label{fig:gan_training}
+  \medskip
+  \small
+\end{figure}
+
+Combining the objective functions of both generator and discriminator in equation \ref{eq:l_g_2} and \ref{eq:l_d}, respectively yields us the following objective with value function $V(G,D)$.
 $$

### 2016-12-20 paper/chapters/gan.tex
more gan

@@
 The payoff of the discriminator can be described as the value function $v(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-v(G,D)$.\\
-This formulation allows us to define the convergence as equation \ref{zerosum_convergence}.
+This formulation allows us to define the convergence as equation \ref{eq:zerosum_convergence}.
 \begin{equation}
   \label{eq:zerosum_convergence}
-  g^* = \arg \min_{G} \max_{D} v(g,d)
+  G^* = \arg \min_{G} \max_{D} V(G,D)
 \end{equation}
@@
 This process can be formalized to maximizing the cost function $\mathcal{L}_d$.
-$$
+\begin{equation}
+  \label{eq:l_d}
   \mathcal{L}_d = \max \log\big(\log D(x) + \log (1 - D(G(z))\big)
-$$
+\end{equation}
 
@@
 \label{sub:gan_training}
+
 \begin{figure}[htb]
@@
 
-Combining the objective functions of both generator and discriminator in equation \ref{eq:l_g_2} and \ref{eq:l_d}, respectively yields us the following objective with value function $V(G,D)$.
-$$
+Combining the two objectives into one value function yields equation \ref{eq:minimax_value}.
+\begin{equation}
+\label{eq:minimax_value}
 \min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
-$$
+\end{equation}
 
-The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable.
+The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable with $G$ and $D$ implemented as deep neural nets.
+Figure \ref{fig:gan_training} illustrates the ideal training procedure for learning of an 1D gaussian.
 Training will oscillate between training the generator and discriminator, where the other party is fixed respectively.

### 2017-01-06 paper/chapters/abstract.tex
more updates

@@
 These models can be used to understand high-dimensional, real-world data and transform them into compact representations.
-Similar to how humans learn to generalize from sensory input, we discuss several models which are able to learn compact, internal representational structures.
-In this article we will focus on two recently proposed architectures, variational autoencoders and generative adversarial networks.
+Similar to how humans learn to generalize from sensory input, we discuss several models which can learn compact, internal representational structures.
+In this article, we will focus on two recently proposed architectures, variational autoencoders, and generative adversarial networks.
 

### 2017-01-06 paper/chapters/conclusion.tex
more updates

@@
-\section{Conclusion}
-\label{sec:conclusion}
+\section{Summary}
+\label{sec:summary}
 
-Generative models have shown great promise to tackle interesting problems.
-Academic work is moving incredibly fast and many new ideas and combinations
-are being proposed almost daily (citation needed, meta paper?).
-For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
-which produced way larger images than GANs alone.
+In this article we have taken a look at different directed models which perform generative modeling using deep neural networks.
+Variational autoencoders as well as generative adversarial networks have shown to produce good results using large datasets with unsupervised learning.
+Deep generative models have shown to learn rich internal representation of high-dimensional data in order to generate samples.
+Going further, as intelligent agents will need to have accurate and generalized internal representations inferred through sensory input to reason about the world, generative models are are likely to be the cornerstones of more advanced intelligent systems in the future.
 
-% not relevant, but wanted to write about it :p
-As mentioned in the motivation, further advances to generative architectures are
-likely to solve current issues with depending on large labelled data.
-Another large and increasing area of research is semi-supervised learning
-(combining learning from labelled data as well as unlabelled data)
-and also one-shot learning, which is where humans excel and machines fail currently.
-\\\\
-Exciting times :)
+Both models explained in detail have been proposed in recent years with many proposed extensions which were not covered in this article, thus further improvements are likely to happen.
+
+Especially interesting is the case of unsupervised learning, due to vast amount of freely available data on the web and in images, video and audio.
+Deep learning has been mostly focussed on supervised algorithms, however these methods require most labelled datasets which are often not feasible to obtain.
+
+%Generative models have shown great promise to tackle interesting problems.
+%Academic work is moving incredibly fast and many new ideas and combinations
+%are being proposed almost daily (citation needed, meta paper?).
+%For example, VAEs and GANs have been combined (https://github.com/skaae/vaeblog)
+%which produced way larger images than GANs alone.
+
+%% not relevant, but wanted to write about it :p
+%As mentioned in the motivation, further advances to generative architectures are
+%likely to solve current issues with depending on large labelled data.
+%Another large and increasing area of research is semi-supervised learning
+%(combining learning from labelled data as well as unlabelled data)
+%and also one-shot learning, which is where humans excel and machines fail currently.
+%\\\\
+%Exciting times :)

### 2017-01-06 paper/chapters/gan.tex
more updates

@@
 %%% when introduced, game theory approach
-Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
+Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modeling.
 %which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
 
-GANs circumvent the issues arising the inference problem by playing a two-player game.
-Both players play non-cooperatively against each other, which means they only try to maximize their respective success or profit.
+GANs circumvent the issues arising around the inference problem by playing a two-player game.
+Both players play non-cooperatively against each other, which means they only try to maximize their respective success.
 One player is called the generator $G$ which tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually real data.
@@
 
-Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
-On the other side, $G$ tries to produce as many real-looking samples as possible to get accepted by $D$.\\
+As each party plays to maximize its own success, the discriminator will try to maximize the number of samples which got discriminated correctly.
+On the other side, $G$ tries to produce samples which fool the $D$ often.
 
-As the game progresses, both parties will get better at their respective objective and the generated samples will look more and more like the training examples.\\
+As the game progresses, both parties will get better at their respective objective and the generated samples will look more and more like the training examples which is the ultimative goal is generative modeling.\\
 
-Using the analogy of counterfeiters and police can help understand the intuition behind GANs better.
+Using the analogy of counterfeiters and police can help understand the intuition behind game between $G$ and $D$ better.
 Counterfeiters will try to produce fake money looking to fool the police into thinking it is real money.
 The police however is able to distinguish badly made fake money from real money easily and as such rejects the fake money with high confidence.
-This rejection in turn triggers the counterfeiters to look at real money more closely and try to imitate it even better.
-Therefore they are able to improve the quality of the faked money.
-Improvements made to the fake money in this iteration will lower the confidence with which the police was able to reject previous samples.
+This rejection in turn triggers the counterfeiters to improve their forgery process try to imitate real money even better.
+Thus they are able to improve the quality of the faked money.
+Improvements made to the fake money in this iteration will lead to lower confidence with which the police is able to reject forgeries.
 
@@
 
-In the case of GANs, this situation means the generator produces samples which can only be rejected with probability $p=0.5$ by the discriminator, meaning it produces as good samples as there are in the dataset.\\\\
+In the case of GANs, this situation arises when the generator produces samples which get rejected with probability $p=0.5$ by the discriminator.
+In the case of such equilibrium, the generator produces good enough samples to fool the generator in about half cases.
 
-Another formulation of the GAN framework is to define the zero-sum game between the generative model and its adversary.
-The payoff of the discriminator can be described as the value function $v(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-v(G,D)$.\\
-This formulation allows us to define the convergence as equation \ref{eq:zerosum_convergence}.
+Another formulation of the same game is to define the zero-sum game between the generative model and its adversary, the discriminator.
+The payoff of the discriminator can be described as the value function $V(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-V(G,D)$.\\
+This formulation allows us to define the reached nash equlibium in equation \ref{eq:zerosum_convergence}.
 \begin{equation}
@@
   The above figure illustrates the computational graph and gradient flow in the GAN framework.\\
-  $X$ denotes the dataset containing real-world samples from $p_{data}$, which we aim to model.\\
+  $X$ denotes the dataset containing real-world data sampled from $p_{data}$, which we aim to model.\\
   Solid arrows indicate the forward pass in the model while dashed arrows denote the gradient flow during a backward pass.
@@
 
-\newpage
+%\newpage
 
@@
 
-In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
+In the case of generative modeling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
 GANs uses neural network as function approximator for $G$ as well as $D$.
-The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
-where $\xi$ denotes the parameters learned by the neural network representing $G$.\\
+The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples $z$ from random noise as $x = G(z)$.\\
 Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.\\
-
-Figure \ref{fig:gan_a} summarizes the described architecture.\\
+Figure \ref{fig:gan_a} summarizes the described architecture with the according gradient flow used to update $G$ and $D$.\\
 %Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
@@
 \subsubsection{Generator Network}
-The generator $G$ in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from the training data.
-This process can be formally written as minimizing the following cost function.
+The generator $G$ in the GAN framework tries to fool the discriminator $D$ by producing samples which are indistinguishable from training data.
+This process can formally be written as minimizing $\mathcal{L}_g$ in equation \ref{eq:l_g_1}.
 %Where $z$ is sampling from the noise prior $p(z)$.\\
@@
   \label{eq:l_g_1}
-  \mathcal{L}_g = \min \log(1 - D(G(z)))
+  \mathcal{L}_g = \min \log\bigg(1 - D(G(z))\bigg)
 \end{equation}
@@
 However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning where the discriminator is able to reject samples from the generator with high confidence.\\
-Therefore is in practice often equation \ref{eq:l_g_2} instead of equation \ref{eq:l_g_1} to optimize its objective.\\
+Therefore is in practice often equation \ref{eq:l_g_2} instead of equation \ref{eq:l_g_1} used to optimize the generator.\\
+
 \begin{equation}
@@
 The objective of the discriminator network $D$ is to distinguish the generated samples produced by $G$ from the training set samples.
-This process can be formalized to maximizing the cost function $\mathcal{L}_d$.
+This process can be formalized to maximizing the cost function $\mathcal{L}_d$ in equation \ref{eq:l_d}.
 \begin{equation}
@@
   \includegraphics[width=\linewidth]{media/gan_training-crop}
-  \caption{Architecture used for generative adversarial networks}
+  \caption{Learning a GAN on a one-dimensional distribution from the original paper\cite{gan:2014}}
   \label{fig:gan_training}
@@
   \small
+  The line named $z$ denoted the noise prior, while the arrows to the line indicated by $x$ is the transformation learned by the generator network.
+Three different distributions are showed, the solid line represents the generated sample in each step, while the dots are sampled from the data distribution $p_{data}(x)$.
+The dashed line shows the decision boundary with which probability the discriminator can distinguish between real and fake data.\\
+  (a) Prior to training when both $D$ and $G$ are poor and $p_{model}$ does not match $p_{data}$.\\
+  (b) After a step of learning $D$ with fixed $G$, the discriminator is able to distinguish fake from real samples easily as indicated by the decision boundary.\\
+  (c) After a step of learning $G$ with fixed $D$ has the generator learned to produce samples more similar to $p_{data}(x)$.\\
+  (d) The model has reached the nash equilibrium where $p_{model} = p_{data}$ and the discriminator can only distinguish between real and fake sample with $p=0.5$.
 \end{figure}
 
-Combining the two objectives into one value function yields equation \ref{eq:minimax_value}.
+Combining the two objectives into one value function yields equation \ref{eq:minimax_value} which is used to define minmax-game played by the two networks.
 \begin{equation}
@@
 
-The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable with $G$ and $D$ implemented as deep neural nets.
+The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable with $G$ and $D$ implemented as (deep) neural nets.
 Figure \ref{fig:gan_training} illustrates the ideal training procedure for learning of an 1D gaussian.
-Training will oscillate between training the generator and discriminator, where the other party is fixed respectively.
-In order to avoid a too strong discriminator (or generator), the ratio of learning steps will vary even though there are no hard rules for this.
+Training will switch between training the generator and discriminator, where the other party is fixed respectively.
 
-\vspace{5cm}
-\paragraph{Theoretical Foundation}
-Using a model with infinite training time and capacity in both generator and discriminator allows theoretical analysis of the GAN framework \cite{gan:2014}.
-For any given generator $G$, the optimal discriminator $D$ is given by $D_G^*(x)$:
-$$
+%\paragraph{Theoretical Foundation}
+Using a model with infinite training time as well as infinite capacity in both generator and discriminator allows to formulate the nash equilibrium in the generative adversarial network.
+For any given generator $G$ and data points $x \sim p_{data}(x)$, the optimal discriminator $D$ is given by equation \ref{eq:opt_d}.
+\begin{equation}
+  \label{eq:opt_d}
   D_G^*(x) = \frac{p_{data}(x)}{p_{data}(x) + p_{model}(x)}
-$$
+\end{equation}
 
@@
 \label{sub:gan_stability}
-In the practical case where both function approximators $G$ and $D$ are represented using a neural network,
-GANs have been shown to be difficult to train using gradient descent.
-These issues have been credited to the problem of finding the nash equilibrium in the minimax-game with continious, high-dimensional parameters whereas current gradient-based methods are [designed] to reach an minimum.
-Additionally to these hypothetical thoughts, \cite{gan_distinguish_crit:2014} have shown that these methods may fail to converge.
-Acknoledging these problems, Salimans et al \cite{improved_gan:2016} has studied the stability and performance of GAN in detail resulting in a number of practical methods improving sample generation and improved semi-supervised learning performance.
-We will take a look at a selection of these methods in the following, focussing on neural networks as approximators.
+In the common case where both function approximators $G$ and $D$ are implemented using deep neural networks, GANs have been shown to be difficult to train using gradient descent.
+In fact, there is no guarantee that GANs will converge in the case of using non-convex functions to represent $G$ or $D$.
+
+%These issues have been credited to the problem of finding the nash equilibrium in the minimax-game with continious, high-dimensional parameters instead of more traditional optimization problems.
+%TheseAdditionally to these hypothetical thoughts, \cite{gan_distinguish_crit:2014} have shown that these methods may fail to converge.
+Acknowledging these problems, Salimans et al.\cite{improved_gan:2016} has studied the stability and performance of GAN in detail, resulting in a number of practical methods improving performance of the training procedure of GANs.
+We will take a look at a selection of these methods in the following, focussing on deep neural networks as approximators in both $D$ and $G$.
 
@@
 \label{ssub:gan_freeze_learning}
-One of the main issues prominent during early GAN research is the disparity between $G$ and $D$.
-Commonly during training one of these networks will outpace the other resulting in degraded gradients where the other player is unable to learn from.
-Freeze learning is one of the natural counter to this problem, just freezing either $G$ or $D$ until the other network has catched up (formulation!).
-In practice the generator is far slower to train than the discriminator in most cases (citation).
+One of the main issues prominent during GAN training is the disparity in performance between $G$ and $D$.
+Most of the time during training will one network outperform the other player by such great margin that the gradients from the stronger player will vanish resulting in even slower training for the weaker network.
+Thus one of the common practices\cite{improved_gan:2016}\cite{dcgan:2015} for learning GANs is to freeze one network and prevent it from learning until the other player has catched up.
+%In practice the generator is far slower to train than the discriminator in most cases (citation).
 
-\subsubsection{Label Smoothing}
+\subsubsection{Label Smoothing \cite{improved_gan:2016}}
 \label{ssub:gan_label_smoothing}
-Another basic idea to improve gradient flow is to replace the binary classification values $0/1$ with smoothed values $0.1/0.9$ or $\epsilon/1-\epsilon$.
-This guarantees to provide the generator with an informative gradient to improve learning performance
+Another rather basic idea to improve gradient flow is to replace the binary classification from the discriminator with smoothed values $0.1/0.9$ or $\epsilon/1-\epsilon$.
+This helps the generator with an informative gradient to improve learning performance.
 Additionally, smoothed labels have been shown to reduce the attack surface of neural networks to adversarial examples \cite{adv_examples:2016}.
@@
 Essentially, BN works by normalizing each sample from a minibatch to the statistics of the whole batch.
-In a paper \cite{improved_gan:2016} by Salimans et al on improving stability of the GAN training procedure, a variant of BN has been proposed called virtual batch normalization.
+To improve training performance of GANs, a modified version of batch normalization has been proposed \cite{improved_gan:2016} for training of the generator.
+Instead of normalizing each input $x$ to their corresponding minibatch, one reference batch is chosen prior to training which will be used to normalize all further input.
+
+%In a paper \cite{improved_gan:2016} by Salimans et al on improving stability of the GAN training procedure, a variant of BN has been proposed called virtual batch normalization.
 
 %Using batch normalization \cite{batch_norm:2015} helped stabilizing learning DCGAN \cite{dcgan:2015}.
-\cite{improved_gan:2016} introduced virtual batch normalization where each input $x$ is normalized on one reference batch chosen at the start of learning instead of the other inputs in the mini-batch.
+%\cite{improved_gan:2016} introduced virtual batch normalization where each input $x$ is normalized on one reference batch chosen at the start of learning instead of the other inputs in the mini-batch.
 
@@
 They may refer to a set of points or edges in an image, which exactly is not clearly defined.
-With the recent advances in classification tasks performed by deep convolutional networks, automatically learned features have been a large part of that success. (citation!)
+With the recent advances in classification tasks performed by deep convolutional networks, automatically learned features have been a large part of that success.
 Feature matching is an technique designed to improve the learning procedure of GANs where not directly $\mathcal{L}_g$ is maximized but instead also learn the discriminator on intermediate layer activation outputs.
-Using this modified procedure the modified objective function aids the generator to output data matching the statistics of the input data.
-With $f(x)$ denoting an intermediate layer in $D$, the new objective function becomes:
-$$
-|| \mathbb{E}_{x \sim p_{data}}f(x) - \mathbb{E}_{z \sim p(z)}f(G(z)) || ^{2}_{2}
-$$
+Using this modified procedure aids the generator to output data matching the statistics of the input data more closely.
+%With $f(x)$ denoting an intermediate layer in $D$, the new objective function becomes:
+%$$
+%|| \mathbb{E}_{x \sim p_{data}}f(x) - \mathbb{E}_{z \sim p(z)}f(G(z)) || ^{2}_{2}
+%$$
 
@@
 
-\subsubsection{DCGAN \cite{dcgan:2015}}
+\subsubsection{Deep Convolutional GAN}
 \label{ssub:dcgan}
-Early implementations of the GAN framework have been implemented using fully-connected layers,
-which have proven to be difficult and slow to train on real-world data (citation, original convnet papers?).
-Using deep convolutional networks has helped reach human-level performance in object recognition and classification problems on images (citations!).
-As the discriminator in the GAN model implements a binary classification task and the generator performance is heavily dependent on the discriminator, using convnets to train the model should improve performance of image generation immensely.
-\emph{Deep convolutional generative adversarial network} (DCGAN) has applied convolutional networks to both generator and discriminator, resulting in quantitively evualuated better results.
-DCGAN uses no fully connected or pooling layers, but instead fractionally strided convolutions (often wrongfully called deconvolutions).
+Deep convolutional networks\cite{lecun-89e} have yielded impressive results for image classification \cite{imagenet:2012} and have arguably been one of the main reasons for the recent success of deep learning.
+These techniques however have mostly been applied in a supervised learning context, not with unsupervised learning.
 
-\subsubsection{LAPGAN \cite{lapgan:2015}}
+The discriminator in the GAN model implements a binary classification task and the performance of the generator is heavily dependent on the discriminator, as the training procedure uses the discriminator in the objective function $\mathcal{L}_{g}$ in equation \ref{eq:l_g_1} and \ref{eq:l_g_2}.\\
+This leads to the suggestion that using deep convolutional networks helps boost the model performance.
+In fact is the deep convolution GAN proposed by Radford et al.\cite{dcgan:2015} one of the first GAN variants which results in visually pleasing results.
+
+In order to use convolutional networks in the generator and discriminator, fractionally strided convolutions have been used to basically upsample the image.
+The exact details of the architecture is beyond the scope of this article, however in addition to the impressive generative performance the authors have shown that the model has learned to produce representional structures in latent space.
+Furthermore, as shown in figure \ref{fig:dcgan_vector} they were able to perform vector artithmetic in the latent space with results indicating that their model was able to arrange the latent space in an useful way.
+
+\begin{figure}[t]
+\centering
+  \includegraphics[width=\linewidth]{media/dcgan_vector}
+  \caption{Vector arithmetic performed in latent space $z$, from DCGAN paper\cite{dcgan:2015}}
+  \label{fig:dcgan_vector}
+  \medskip
+  \small
+\end{figure}
+
+\newpage
+\subsubsection{Laplacian Generative Adversarial Networks}
 \label{ssub:lapgan}
-LAPGAN is an extension of GAN applied to images using laplacian pyramids \cite{laplacian:1983}.
-Instead of using a single generator network, LAPGAN uses a series of deep convolutional networks as generators.
+%Applied to unsupervised learning in the case of GANs, deep convolutional GANs (DCGAN\cite{dcgan:2015}) has shown great success in producing images with
+
+%Early implementations of the GAN framework have been implemented using fully-connected layers,
+%which have proven to be difficult and slow to train on real-world data (citation, original convnet papers?).
+%Using deep convolutional networks has helped reach human-level performance in object recognition and classification problems on images (citations!).
+%\emph{Deep convolutional generative adversarial network} (DCGAN) has applied convolutional networks to both generator and discriminator, resulting in quantitively evualuated better results.
+%DCGAN uses no fully connected or pooling layers, but instead fractionally strided convolutions (often wrongfully called deconvolutions).
+
+\begin{figure}[t]
+\centering
+  \includegraphics[width=\linewidth]{media/bedroom_scenes-crop}
+  \caption{Bedroom scenes generated by the DCGAN model, from the original paper\cite{lapgan:2015}}
+  \label{fig:lapgan}
+  \medskip
+  \small
+\end{figure}
+Laplacian Generative Adversarial Networks (LAPGAN\cite{lapgan:2015}) is an extension to GANs applied to images using laplacian pyramids \cite{laplacian:1983}.
+Instead of using a single generator network, LAPGAN uses a series of deep convolutional networks as generators where each generator receives the output of the previous generator.
 
@@
 \paragraph{Results}
-As GANs have no direct measurement of the quality of generations due to their lack of objective function, human evaluation of the authenticity of generated images is used for quantitive results.
-LAPGAN has pushed the state of the art significantly, with generated images which are high-resolution in comparison to earlier methods and arguebly high variety of generations.
-Furthermore, humans have evaluated the generated images for real images 40\% of the time.
+As GANs have no direct measurement of the quality of generations due to their lack of unified objective function, human evaluation of the authenticity of generated images is often used for qualitively evaluation.
+LAPGAN has pushed the state of the art significantly, with generated images which are high-resolution in comparison to earlier methods and arguably high variety of generated images as shown in figure \ref{fig:lapgan}.
+Furthermore, humans have mistaken the generated images for real images 40\% of the time.

### 2017-01-06 paper/chapters/introduction.tex
more updates

@@
 
-"What I cannot create, I do not understand." -- Richard Feynman \\\\\\
+"What I cannot create, I do not understand." -- Richard Feynman \\\\
 
-Directed generative models are a wide class of models used for generating samples from an unknown distribution.
-Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
-In order to do so, we define an approximated distribution $p_{model}$ from which we are able to sample.
+Directed generative models are a wide class of models used for generating samples from an unknown distribution $p_{data}$.
+%Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
+The presented models in this article will approximate $p_{data}$ with $p_{model}$ which enables us to draw samples from the model.
 Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\
 
-Generative modelling has several important application areas, which will be highlighted on below.\\
-Deep learning techniques have provided great success in applying gradient-based optimization methods to deep neural networks to achieve impressive results in discriminative models.
-These models require however large amounts of labelled data, which is often expensive or infeasible to obtain.
-Generative models can be used to learn \emph{representational} structures from rich, unlabelled data by having less parameters than the input data and are thus forced to capture only essential information.\\
+\paragraph{Short-term applications}
+\begin{itemize}
+  \item \textbf{Data augmentation} is a technique to enrich an existing dataset by adding more derived information to provide more in-depth insight.
+  \item \textbf{Semi-supervised learning} on datasets with a small amount of unlabelled data and a large amount of labeled data to improve model performance.
+Examples include the conditional variational autoencoder discussed in \ref{ssub:cvae}.
+  \item \textbf{Data exploration and understanding} is often needed especially for high-dimensional and complex data.
+By using models which reduce the dimensions by selecting only important features for generative modeling can help to improve data understanding.
+Figure \ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
+\end{itemize}
+
+In addition to those short-term applications, there exist possibilities for more long-term aspects of generative modeling.
+For example with unsupervised representation learning, where statistical models are able to learn robust, accurate and generalized internal representations from high-dimensional sensory input can help power future agents to help understand real-world data, allowing these agents to take actions and reason about the real world.
+
 \newpage
+%Deep learning techniques have provided great success in applying gradient-based optimization methods to deep neural networks to achieve impressive results in discriminative models.
+%These models require however large amounts of labelled data, which is often expensive or infeasible to obtain.
+%Generative models can be used to learn \emph{representational} structures from rich, unlabelled data by having less parameters than the input data and are thus forced to capture only essential information.\\
+%\newpage
 
@@
 %The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
-Models using deep learning often assume that observed data is stochastically dependent on some hidden structure which we wish to explain.
-In order to understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real world data.
+Models using deep learning often assume that observed data $x$ is stochastically dependent on some hidden structure $z$.
+To understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real-world data.
+
 \begin{itemize}
-  \item \textbf{Images} consisting of millions of pixels can be interpreted as data where each pixel represents an own dimension.
+  \item \textbf{Images} consisting of millions of pixels can be interpreted as data vectors with each pixel representing an own dimension.
   However, these dimensions are high correlated and structured, for example nearby pixels have with high probability similar values.
   Looking at images of handwritten digits, the data can be reduced to a handful variables which may include the digit number, stroke width, position or size.
-  Statistical models help us discover these relations and reduce them to representations in lower dimensional space.
+  Statistical models help us discover these relations and reduce them to representations in lower-dimensional space.
   %\item commonly have thousands of dimensions, but the underlying structure is often far simpler.
@@
   \small
-  Arrows show stochastic dependency.
+  The latent variable $z$ is a stochastic dependency of the observed data $x$.
 \end{figure}

### 2017-01-06 paper/chapters/more.tex
added GAN architecture image
updated
more updates

@@
+\newpage
 \section{More Generative Directed Models}
@@
 
+There are several more deep directed generative models than shown here, however in order to discuss some models more in detail only two recent promising models have been presented.
+
 \subsection{Generative Moment Matching Networks}
 \label{sub:more_mmn}
+Generative moment matching networks (GMMN) present a similar idea to GANs, however instead of posing the generative modelling problem as game it is optimized using \emph{maximum mean discrepancy} (MMD).
+
+\vspace{10cm}
 
@@
 
-See \cite{gan_openai:2016} and \cite{gan_training:2016}.
+%See \cite{gan_openai:2016} and \cite{gan_training:2016}.

### 2017-01-06 paper/chapters/sigmoid_belief_nets.tex
more updates

@@
   \small
-  Bayesian networks, also called belief nets can be described as an acyclic directed graph where blue nodes indicate hidden variables while red nodes represent observed variables.
-  Arrows show stochastic dependency. The dashed arrow indicates that it is possible in the original definition, even though most commonly the hidden variables are divided into layers.
+  Bayesian networks, also called belief nets are an acyclic directed graph where blue nodes indicate hidden variables and red nodes represent observed variables.
+  Arrows indicate stochastic dependency.
+  Even though most belief nets are divided into many layers in the context of deep learning, the original definition does not specify that and thus indicates the dashed arrow the possibility of inter-layer dependencies.
 \end{figure}
 
-Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
-Figure \ref{fig:sbn_arch} shows a generic belief network as a graphical model with each state influenced by its ancestors.
+Sigmoid Belief Networks (SBN) were proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985} shown in figure \ref{fig:sbn_arch}
 In the case of SBNs however, all states are binary and the activation function is the sigmoid function.
-They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
+They are one of the earliest neural networks used for generative modeling, predating all other presented models in this article.
 %As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
@@
 Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
-Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\
+Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the states in the visible layer.\\
 
 Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
-Especially in deep networks, this kind of computation is infeasible as it requires evaluating exponentially many configurations.\\
+Especially in deep networks, this kind of computation is infeasible as it requires computing exponentially many configurations.\\
 

### 2017-01-06 paper/chapters/vae.tex
more updates

@@
 \label{sec:vae}
-Variational autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
+Variational autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al.\cite{vae:2014} and Rezende et al.\cite{dlgm:2014}.
 %The proposed framework resembles the encoder-decoder architecture, using a probabilistic encoder and decoder.
-The general idea of VAEs is to have an encoder-decoder architecture where inputs $x$ gets mapped to latent variables $z$ and from there to the reconstruction $~x$.
+The general idea of VAEs is to have an encoder-decoder architecture where input $x$ gets mapped to a specific distribution defined in the latent space $z$ from which we sample in order to reconstruct the input.
 
@@
 
-However, in addition to this autoencoder structure an additional regularizer forces the latent space to have a specific simple form.
-To achieve this, variational-based methods are used to output a distribution on the latent space instead of directly emitting samples.
+%However, in addition to this autoencoder structure an additional regularizer forces the latent space to have a specific simple form.
+%To achieve this, variational-based methods are used to output a distribution on the latent space instead of directly emitting samples.
 
 The probabilistic encoder generally implemented as feed-forward network receives input from the dataset and outputs parameters of a distribution.
-For example the parameters for a gaussian distribution with diagonal covariance, in which case the networks emits $(\mu,\Sigma)$.\\
+For example the parameters for a gaussian distribution with diagonal covariance, in which case the networks emits the mean $\mu$ and $\sigma$ used to define the gaussian distribution $\mathcal{N}(\mu,\sigma \times I)$, where $I$ is the identity matrix.\\
 
-Using gradient-based optimization techniques enables the network to be trained to be similar to the true posterior $p(z|x)$.
+Using gradient-based optimization techniques enables the encoder network to be trained to output distributions similar to the true posterior $p(z|x)$.
 In addition to this inference network there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.
@@
 
-Formally, the described probabilitic encoder emits variational parameters $\phi$ used to define the probability distribution $q_\phi(z|x)$.
-Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.\\
+Properly defined emits the described probabilitic encoder variational parameters $\phi$ used to define the probability distribution $q_\phi(z|x)$.
+Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
+This architecture is shown in figure \ref{fig:vae_architecture} with a one-dimensonal gaussian prior.
+\\
 
@@
 The key insight on how the variational autoencoder is able to optimize $q_\phi(z|x)$ to approach the true data posterior distribution $p(z|x)$ is in defining a variational lower bound which is used to push the approximated $q_\phi(z|x)$ towards $p(z|x)$.
-We will discuss this lower bound $\mathcal{L}$ in the following section in detail.
+We will discuss this lower bound $\mathcal{L}$ and its derivation in detail in the following sections.
 
@@
 % MCMC
-Using Markov chain Monte Carlo (MCMC) methods, which construct a Markov chain with the desired attributes.
+Using Markov chain Monte Carlo (MCMC) methods, which construct a Markov chain with the desired attributes matching those of the posterior.
 The advantage of this approach is that is guaranteed to asympotically approach the desired posterior, however it is difficult to reason about the convergence rate.
@@
 % variational inference
-The other approach which is using variational inference which picks a specific distribution restricted to a family of distributions matching the posterior most closely.
+Using variational inference is the second approach, used in the variational autoencoder, which picks a specific distribution restricted to a family of distributions matching the posterior most closely.
 Advantages using this approach rather than MCMC is the greater speed at which the computation is performed.
-However it is not guaranteed to even approximate the posterior accurately, if the true posterior can not be expressed as a distribution the family.
+However it is not guaranteed to approximate the posterior accurately, if the true posterior can not be approximately expressed as a distribution of the family.
 
@@
 \subsubsection{Kullback-Leibler Divergence}
-In order to approximate one distribution $Q$ to another distribution $P$, a measurement of difference.
-One commonly used measurement is called Kullback-Leibler divergence (short KL-divergence) which enables the variational inference methods to choose appropiate distributions.
+In order to approximate one distribution $Q$ to another distribution $P$, there needs to be a measurement how close $Q$ is to $P$.
+One commonly used measurement is called Kullback-Leibler divergence (short KL divergence) which enables the variational inference methods to choose appropiate distributions.
 %\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
 $\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
-For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} as:
+For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
 \begin{equation}
+  \label{eq:kl_div_disc}
   \mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
+\end{equation}
+For continuous distributions $P$ and $Q$ the KL divergence is defined similarly in equation \ref{eq:kl_div_cont}.
+\begin{equation}
+  \label{eq:kl_div_cont}
+  \mathcal{D}_{\mathrm{KL}}(P || Q) = \int P(i) \log \frac{P(i)}{Q(i)} di
 \end{equation}
@@
 \begin{equation}
+  \label{eq:vae_obj}
   q_\phi^{*}(z|x) = \arg\min_{\phi} \mathcal{D}_{\mathrm{KL}} \big(q_\phi(z|x) || p(z|x)\big)
@@
 
-However we are not able to optimize for that goal directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
+However we are not able to optimize equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
 The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
@@
 
-Computing the integral in equation \ref{eq:integral} requires the evaluation of every possible configuration over the hidden variables $z$ and therefore intractable.\\\\
+Computing the integral in equation \ref{eq:integral} requires the evaluation of every possible configuration over the hidden variables $z$ and is therefore intractable.\\\\
 
-Instead of directly optimizing for the true posterior, we construct an evidence lower bound (ELBO) $\mathcal{L}$ which we can maximize which is equivalent to minimizing the KL divergence between $q_\phi(z|x)$ and $p(z|x)$.\\
+Instead of directly optimizing for the true posterior, we construct an evidence lower bound (ELBO) $\mathcal{L}$ which we can maximize which is equivalent to minimizing the KL divergence between $q_\phi(z|x)$ and $p(z|x)$ without explicity evaluating the evidence.\\
 
@@
 
-With the example using handwritten digits it appears why this regularizer is important for the model to learn generalized representations in latent space.
+Using the example of handwritten digits it appears why this regularizer term $\mathcal{L}_{reg}$ is important for the model to learn generalized representations in latent space.
 Suppose we have two data points from the dataset which both represent the same handwritten digit.
-Ideally, our model would learn the representations in latent space to be fairly close to ensure the model has learned to generalize the semantics.
+Ideally, our model would learn representations in latent space to be fairly close to ensure the model has learned to generalize the semantics.
 Without $\mathcal{L}_{reg}$ the model would assign each data example a different representation in latent space with a distribution collapsing to that single point.
@@
 \paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
-This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\phi(z | x_i)$.\\
+This error is measured using the negative log-likelihood of the conditional probability distribution $p_\theta(x|z)$ where $z$ is sampled from $q_\phi(z | x)$.\\
 $\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
@@
   \caption{reparameterization trick}
+  \label{fig:rep_trick}
   \medskip
   \small
-  Function $f_\theta(x)$ with inputs $x$ and (variational) parameters $\theta$ shown as directed graphical model.\\
+  Function $f_\theta(x)$ with inputs $x$ and (variational) parameters $\theta$ as computational graph.
   Left without and on the right with the reparameterization trick applied.\\
-  Solid arrows show the forward pass and dashed arrows the error backpropagation.
+  Solid arrows show the forward pass and dashed arrows the error backpropagation using gradients.
   Rectangular nodes represent deterministic functions while circles indicate stochastic variables.
@@
 To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
-Because we add noise to the encoding, the gradients can't be computed directly. In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
-Instead of drawing $z ~ \mathcal{N}(\mu(x), \Sigma(x))$, we sample an auxiliary variable $\epsilon$ from $\mathcal{N}(0, I)$ which we then transform with equation \ref{eq:rep_trick} into $z$.
+However, because we only sample from the constructed distribution $q_\phi(z|x)$, we cannot backpropagate through the node due to the stochastic nature.
+In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
+Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample first an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.
+Then we apply equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
 \begin{equation}
@@
 \end{equation}
-This allows us to compute the gradient of the loss function and backpropagate through the entire model and only have the stochastic variable $\epsilon$ as input.\\\\
+Note that this transformation is entirely deterministic and as such allows for backpropagation through the transformation into the variational parameters $\phi$.
 
-The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
-In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
+Figure \ref{fig:rep_trick} expresses the reparameterization trick applied to an arbitrary function $f$ with the parameters $\theta$.\\
+Using this trick, the whole VAE model can be trained using gradient-based optimization techniques.\\\\
 
-We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
-The prior distribution $p(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
-The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2014} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
-Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
-\begin{equation}
-  \mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
-\end{equation}
+Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different probability distributions on $z$.
+%The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
+%In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).
+
+%We use VAEs when we have a complicated distribution $p_\theta(x,z)$ with unknown latent variables $z$.
+%The prior distribution $p(z)$ over the latent structure is a centered isotropic(?) multivariate Gaussian denoted by $\mathcal{N}(z;0, I)$.
+%The constructed architecture uses an probabilistic encoder $q_\theta(z|x)$ and a probabilistic decoder $p_\theta(x|z)$ in a form of neural networks, in the original paper MLP are used\cite{vae:2014} but there are various extensions (see \ref{sub:vae_extensions}, TODO).
+%Because the posterior is intractable ($p_\theta(x)$), we instead approximately maximize the lower variational bound $L(\theta,\phi;x)$.\\
+%\begin{equation}
+  %\mathcal{L}(\theta,\phi;x) = -D_{KL}(q_\theta(z|x)||p(z)) + \mathbb{E}_{q_\theta(z|x)}[log_{p_\theta}(x|z)]
+%\end{equation}
 
@@
 \label{sub:vae_learning}
-The probabilistic encoder $q_\phi(z|x)$ and decoder $p_\theta(x|z)$ are commonly implemented using a deep neural network.
-To allows analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.
+The probabilistic encoder $q_\phi(z|x)$ and decoder $p_\theta(x|z)$ are commonly implemented using deep neural networks.
+To allow analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.
 
@@
 \subsubsection{Conditional VAE \cite{cvae:2014} \cite{cvae_2:2015}}
-\label{ssub:vae_conditional_vae}
+\label{ssub:cvae}
 \input{chapters/vae/cvae}
@@
 %\end{figure*}
-Up until now the presented variational autoencoder only has one stochastic latent layer.
+Up until now the presented variational autoencoder framework has only considered one stochastic latent layer.
 However in the original paper by Rezende et al.
 \cite{dlgm:2014} the authors have shown a variational autoencoder using multiple stochastic layers with gaussian priors in each layer.
-Note that the deterministic non-linear transformations between each layer still can be represented using deep neural networks.
-This enables the model to learn richer, more hierarchical structures captured in each latent layer.
+They called this model the deep latent gaussian model (DLGM) which can be understood as a generalization of the variational autoencoder.
+Note that the deterministic non-linear transformations between each layer still can be represented using deep neural networks.\\
+Using multiple stochastic layers enables the model to learn more hierarchical structured dependencies from the dataset.
 
@@
 
-\subsubsection{Importance Weighted Autoencoder \cite{iwae:2015}}
+\subsubsection{Importance Weighted Autoencoder}
 \label{ssub:vae_importance_weighted_autoencoder}
@@
 
-\subsubsection{Deep Recurrent Attention Writer \cite{draw:2015}}
+\subsubsection{Deep Recurrent Attention Writer}
 \label{ssub:vae_deep_recurrent_attention_writer}

### 2017-01-06 paper/chapters/vae/cvae.tex
more updates

@@
-Instead of using VAEs in a completely unsupervised manner, oftentimes learning from a dataset where a small subset of datapoints has corresponding labels.
+Instead of using VAEs in a completely unsupervised manner on a unlabeled dataset, it is also possible to learn from datasets where a small subset of datapoints has corresponding labels while most datapoints are still unlabeled.
 Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly.\\
@@
   \label{eq:cvae_labels}
-  \log p_\theta(x) &\geq \mathbb{E}_{q_\phi(z|x,y)} \bigg[log p_\theta(x|y,z) + \log p_\theta(y) + \log p(z) - \log q_\phi(z|x,y)\bigg]
+  \log p_\theta(x) &\geq \mathbb{E}_{q_\phi(z|x,y)} \bigg[\log p_\theta(x|y,z) + \log p_\theta(y) + \log p(z) - \log q_\phi(z|x,y)\bigg]
 \end{align}

### 2017-01-06 paper/chapters/vae/iwae.tex
zZzZzzz almost done
more updates

@@
 % write about multiple stochastic layers
-Until now, the presented model includes only one stochastic layer.
-\cite{dlgm:2014} have presented a model for learning through multiple stochastic layers which can be easily applied to the VAE framework as shown in \ref{fig:vae_multiple_layers}.
 
@@
 % write about IWAE, extends these k passes with importance weights
-Importance Weighted Autoencoders (IWAE) extends on the VAE framework using $k$-sample importance weighting of the log-likelihood.
+Importance Weighted Autoencoders (IWAE\cite{iwae:2015}) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
 % add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
@@
 
-The authors of the papers argue that using more samples will only tighten the lower variational bound.
-$$
+Improving on the theoretical results in the VAE framework, using IWAE with more samples will tighten the lower variational bound and therefore improve the overall performance of the model.
+\begin{equation*}
 \log p(x) \geq \mathcal{L}_{k+1}(x) \geq \mathcal{L}_{k}(x)
-$$
+\end{equation*}
 
-Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum:
-$$
+Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum as $k$ approaches infinity in equation \ref{eq:iwae}
+\begin{equation}
+\label{eq:iwae}
 \lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
-$$
+\end{equation}
 
-Of course, using more samples will result in higher computational cost.
+However, using more samples will result in higher computational cost and thus should be carefully evaluated.
 
-\vspace{5cm}
+%\vspace{5cm}

### 2017-01-09 slides/figures/vae_conceptual.tex
synced

\documentclass{standalone}

\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{fit,backgrounds}

\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,outer sep=0.1cm,inner sep=0.3cm,minimum size=10,line width=0.5, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{tikzpicture}

%\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0cm,0cm) {$x$};
\node[fill=black!20] (lbl_input) at (0,-1) {Input};

% output / reconstruction
\node[gb] (output) at ($(input) + (3.0cm,0.0cm)$) {$\tilde x$};
\node[fill=black!20] (lbl_rec) at ($(output) + (0,-1cm)$) {Reconstruction};

% latent space
\node[fill=black!20] (lbl_lat) at ($(input) + (2.7cm,5.2cm)$) {$z \sim \mathcal{N}(\mu, \sigma^2 \odot I)$};

    %\begin{pgfonlayer}{background}

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  width=3.5cm,
  xtick=\empty,
  xlabel={$(\mu,\sigma)$},
  at={($(input) + (-0.95cm,4.5cm)$)}
  ]
  \addplot [thick,cyan!50!black, domain=0:6] {gauss(3,1)};
  \node[] (left_bot) at (axis cs:0.0,-0.15) {};
  \node[] (left_top_aux) at (axis cs:0.0,0.2) {};
\end{axis}

  \node [above right=of left_top_aux] (left_top) {Latent space};

    %\end{pgfonlayer}
\begin{scope}[on background layer]
  \node (outer) [dashed,draw,fill=blue!10,fit=(left_bot) (left_top) (lbl_lat)] {};
\end{scope}

% input -> latent
\draw[arr] (input) -- node[label=left:$q_\phi(z|x)$] {}($(input) + (0.0cm,3.8cm)$) {};

% latent -> output (reconstruction)
\draw[arr] ($(input) + (3.0cm,4.5cm)$) -- node[label=right:$p_\theta(x|z)$] {}(output) {};

%\end{scope}

\end{tikzpicture}
\end{document}

### 2017-01-09 slides/figures/vae_conceptual.tex.preview.tex
synced

--- 2017-01-09 slides/figures/vae_conceptual.tex
@@
 \documentclass{standalone}
-
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.14}
 \usetikzlibrary{fit,backgrounds}
 
+\pgfplotsset{compat=1.14}
 \usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}
@@
 \node[gb] (output) at ($(input) + (3.0cm,0.0cm)$) {$\tilde x$};
-\node[fill=black!20] (lbl_rec) at ($(output) + (0,-1cm)$) {Reconstruction};
+%\node[fill=black!20] (lbl_rec) at ($(output) + (0,1cm)$) {Reconstruction};
 
@@
 
+%\begin{scope}
     %\begin{pgfonlayer}{background}
+%\begin{scope}[on background layer]
+%  \node (outer) [draw,fill=blue,fit=($(input) + (-0.95cm,4.5cm) (output)] {};%
+%\end{scope}
+%\node (ax)  [above of=input,label=above:$s\le 3$] {};
 
@@
   ytick=\empty,
-  width=3.5cm,
   xtick=\empty,
   xlabel={$(\mu,\sigma)$},
+  width=3.5cm,
   at={($(input) + (-0.95cm,4.5cm)$)}
   ]
-  \addplot [thick,cyan!50!black, domain=0:6] {gauss(3,1)};
-  \node[] (left_bot) at (axis cs:0.0,-0.15) {};
-  \node[] (left_top_aux) at (axis cs:0.0,0.2) {};
+  \addplot [very thick,cyan!50!black, domain=0:6] {gauss(3,1)};
+
 \end{axis}
+    %\end{pgfonlayer}
 
-  \node [above right=of left_top_aux] (left_top) {Latent space};
-
-    %\end{pgfonlayer}
-\begin{scope}[on background layer]
-  \node (outer) [dashed,draw,fill=blue!10,fit=(left_bot) (left_top) (lbl_lat)] {};
-\end{scope}
+%\end{scope}
 
 % input -> latent
-\draw[arr] (input) -- node[label=left:$q_\phi(z|x)$] {}($(input) + (0.0cm,3.8cm)$) {};
+\draw[arr] (input) -- node[label=left:$q_\phi(z|x)$] {}($(input) + (0.0cm,4.0cm)$) {};
 
@@
 
+\usetikzlibrary{calc}
+\pgftransformreset
+\node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt] at (current bounding box) {};
+%add border to avoid cropping by pdflibnet
+\foreach \border in {0.1}
+  \useasboundingbox (current bounding box.south west)+(-\border,-\border) rectangle (current bounding box.north east)+(\border,\border);
+\newwrite\metadatafile
+\immediate\openout\metadatafile=\jobname_BB.txt
+\path
+  let
+    \p1=(current bounding box.south west),
+    \p2=(current bounding box.north east)
+  in
+  node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt,draw=white] at (current bounding box) {
+\immediate\write\metadatafile{\p1,\p2}
+};
+\immediate\closeout\metadatafile
 \end{tikzpicture}

### 2017-01-09 slides/figures/vae_with_rep.tex
synced

\documentclass{standalone}

\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{fit,backgrounds}

\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,fill=blue!10,outer sep=0.1cm,minimum height=4em, minimum width=4em]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{tikzpicture}

%\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0cm,0cm) {\Huge$x$};
%\node[fill=black!20] (lbl_input) at (0,-1) {Input};

    %\begin{pgfonlayer}{background}
\node[above=3cm of input] (lat_start) {};
\node[above right=1.25cm and 0.85cm of lat_start] (lat_dist) {};
%\node[draw,right=3cm of lat_start] (lat_end) {};

% latent space
\node[] (lbl_lat) at ($(lat_dist) + (8.5cm,0.0cm)$) {\huge $z = \mu + \epsilon \odot \sigma$};

\node[above=of lbl_lat] (lbl_eps) {\huge$\epsilon \sim \mathcal{N}(0, I)$};

% output / reconstruction
\node[gb] (output) at ($(lbl_lat) + (0.0cm,-5.43cm)$) {\huge$\tilde x$};
%\node[fill=black!20] (lbl_rec) at ($(output) + (0,-1cm)$) {Reconstruction};

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  width=3.5cm,
  xtick=\empty,
  xlabel={\huge$(\mu,\sigma)$},
  at={($(lat_start) + (-0.95cm,0.8cm)$)}
  ]
  \addplot [thick,cyan!50!black, domain=0:6] {gauss(3,1)};
  %\node[] (left_bot) at (axis cs:0.0,-0.15) {};
  %\node[] (left_top_aux) at (axis cs:0.0,0.2) {};
\end{axis}

  %\node [above right=of left_top_aux] (left_top) {Latent space};

    %\end{pgfonlayer}
%\begin{scope}[on background layer]
%  \node (outer) [dashed,draw,fill=blue!10,fit=(left_bot) (left_top) (lbl_lat)] {};
%\end{scope}

% input -> latent
\draw[arr] (input) -- node[label=left:\huge$q_\phi(z|x)$] {}(lat_start) {};

% latent -> latent
\draw[arr] (lat_dist) -- (lbl_lat) {};

\draw[arr] (lbl_eps) -- (lbl_lat);

%\node [label={[xshift=1.0cm, yshift=0.3cm]Label}] {Node};
% latent -> output (reconstruction)
\draw[arr] (lbl_lat) -- node[label=right:\huge$p_\theta(x|z)$] {}(output) {};

%\end{scope}

%\draw[dashed] (4.5,7.5) -- (4.5,-2);
\end{tikzpicture}
\end{document}

### 2017-01-20 slides/figures/gan_conceptual.tex
fixed font size in GAN

--- 2016-12-07 paper/media/gan_conceptual.tex
@@
 
+\usetikzlibrary{arrows.meta}
+
 \begin{document}
-\usetikzlibrary{arrows.meta}
 % why does the figure seems to have lots of whitespace to the left when shapes[.symbols] is included??
@@
 % latent space
-\node[circle,gb] (latent_space) {Latent space $z$};
+\node[circle,gb] (latent_space) {\large Random Noise};
 
 % nodes
-\node [gb] (generator) at ([shift={(5,0)}] latent_space) {Generator $G$};
-\node [cloud] (data) at ([shift={(0,4)}] generator) {data space};
-\node [gb] (discriminator) at ([shift={(5,2)}] generator) {Discriminator $D$};
-  \node[gb,dashed] (loss) at ([shift={(5,0)}] discriminator) {Loss $\mathcal{L}$};
+\node [gb] (generator) at ([shift={(5,0)}] latent_space) {\large Generator $G$};
+  \node [cloud] (data) at ([shift={(0,4)}] generator) {\Large $X$};
+
+\node [gb] (discriminator) at ([shift={(5,2)}] generator) {\large Discriminator $D$};
+  \node[gb,dashed] (loss) at ([shift={(5,0)}] discriminator) {\large Loss $\mathcal{L}$};
 
 % flow
-\draw[arr] (latent_space) -- (generator) node[midway,label=below:sample] {};
-\draw[arr] (generator) -- (discriminator) node[midway,label=below:generate] {};
-\draw[arr] (data) -- (discriminator) node[midway,pos=0.4,label=below:sample] {};
-\draw[arr] (discriminator) -- (loss) node[midway,label=below:Real/Fake] {};
+\draw[arr] (latent_space) -- (generator) node[midway,label=below:\large sample] {};
+  \draw[arr] (generator) -- (discriminator) node[midway,pos=0.7,label=below:\large generate] {};
+\draw[arr] (data) -- (discriminator) node[midway,pos=0.4,label=below:\large sample] {};
+\draw[arr] (discriminator) -- (loss) node[midway,label=below:\large real/fake] {};
 
 % gradient flow
-\path[dashed,arr] (loss) edge[bend right] node [label=$\nabla \mathcal{L}$ with regard to $D$] {} (discriminator);
-\path[dashed,arr] (loss) edge[bend left] node [label=below:$\nabla \mathcal{L}$ with regard to $G$] {} (generator);
+\path[dashed,arr] (loss) edge[bend right] node [label=\large $\nabla \mathcal{L}$ with regard to $D$] {} (discriminator);
+\path[dashed,arr] (loss) edge[bend left] node [label=below:\large $\nabla \mathcal{L}$ with regard to $G$] {} (generator);
 

### 2017-01-20 code/meta_train_plot.py
added meta training script for hyper-param search

@@ -1,0 +1,163 @@
+#!/usr/bin/env python3
+from Network import VAE
+import numpy as np
+import matplotlib.pyplot as plt
+from random import choice,choices
+import os
+import input_data
+import seaborn as sns
+from matplotlib.ticker import FormatStrFormatter
+
+mnist = input_data.read_data_sets('MNIST_data', one_hot=True)
+
+
+
+
+class Model:
+    def __init__(self, arch):
+        self.arch = arch
+        self.latent_dim = arch[-1]
+        self.model_name = "-".join(map(str,arch))
+        results_dir = "results"
+        self.results_dir = os.path.join(results_dir, self.model_name)
+        print('Using results directory: ' + self.results_dir)
+
+        # create results directory if needed
+        if not os.path.isdir(self.results_dir):
+            os.makedirs(self.results_dir)
+
+        self.v = VAE.VAE(arch=arch, name = self.model_name)
+
+    def train(self, num_epochs):
+        loss = np.array(list(self.v.train(mnist, num_epochs = num_epochs)))
+        return loss
+
+    def get_pdf(self, name):
+        return os.path.join(self.results_dir, name) + ".pdf"
+
+    def plot_loss(self, loss):
+        sns.set_style("darkgrid")
+        train_loss, = plt.plot(loss[:,0], loss[:,1], label = 'Train loss')
+        test_loss, = plt.plot(loss[:,0], loss[:,2], label = 'Test loss')
+        plt.xlabel('Epochs')
+        plt.ylabel('Negative log-likelihood')
+        plt.legend(handles=[train_loss, test_loss])
+        formatter = FormatStrFormatter('%d')
+        plt.gca().get_xaxis().set_major_formatter(formatter)
+        plt.savefig(self.get_pdf('loss'))
+        sns.set_style(None)
+        # also save loss to file
+        np.savetxt(os.path.join(self.results_dir, 'loss.np'), loss)
+        plt.close()
+
+    def plot_manifold(self, axes = (0,1)):
+        """Plots a manifold along the specified `axes`
+        Please only do this if you have a two-dimensional latent space.
+        To get good manifold in higher dimensions, you probably want to
+        sample from a n-dimensional z-sphere."""
+        # plot 2D manifold
+        range_x1 = (-3, 3)
+        range_x2 = (-3, 3)
+        num_rowdigits = 20
[…]

### 2017-01-20 code/Network/VAE.py
update model, add some saver code, remove plot

@@ -8,12 +8,8 @@ import os
 import logging
 from Layer.Dense import Dense
 from Utils import compose_all
-import matplotlib.gridspec as gridspec
+import pickle
 
-
-from progressbar import ETA, Bar, Percentage, ProgressBar
-import matplotlib.pyplot as plt
-from IPython import display
 
 class VAE(object):
     """Variational Autoencoder implementation in TensorFlow using PrettyTensor
@@ -30,17 +26,18 @@ class VAE(object):
             "learning_rate": 1e-3,
             "nonlin": tf.nn.relu, # TODO relu
             "squash": tf.nn.sigmoid, # TODO explore (softplus?)
+            "arch": [784,512,512,50],
             }
 
     def __init__(self,
-            network_architecture = [], # list of nodes per layer
             hyper_params = {}, # update to HYPERPARAMS
-            models_dir = "models/",
+            models_dir = "models",
             k = 1, # number of forward passes
+            name = None, # model name
             **kwargs
             ):
         self.k_passes = k
-        self.arch = network_architecture
+        # self.arch = network_architecture
         if not os.path.isdir(models_dir):
             os.makedirs(models_dir)
         self.models_dir = models_dir
@@ -48,21 +45,40 @@ class VAE(object):
         tf.reset_default_graph()
         self.__dict__.update(VAE.HYPERPARAMS, **kwargs)
         self.sess = tf.Session()
+        self.saver = None
+        self.name = name
+        
+
+        # Saves to $models_dir/$name/$name-.*.{meta,index,...}
+        # if name:
+            # self.meta_graph_path = os.path.join(self.models_dir, self.name, "{}.meta".format(name))
+            # if not os.path.isdir(os.path.join(models_dir, name)):
+                # os.makedirs(os.path.join(models_dir, name))
+            # else:
+                # print('Trying to restore model "{}"'.format(name))
+                # if os.path.isfile(self.meta_graph_path):
+                    # saver = tf.train.import_meta_graph(self.meta_graph_path)
+                    # saver.restore(self.sess, tf.train.latest_checkpoint('{}/{}/'))
+                    # restored_vars = tf.trainable_variables()
+                    # train_vars = tf.trainable_variables()
+                    # print('Restored {} trainable variables:'.format(len(train_vars)))
+                    # for v in train_vars:
+                        # print("\t{}".format(v.name))
+                    # return
 
[…]

### 2017-01-20 slides/figures/latent_model.tex
updated & added figures

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta,fit}
\begin{document}
\tikzstyle{hidden} = [draw,circle,outer sep=0.05cm,inner sep=0.3cm,line width=1, thick, draw=black!55]
\tikzstyle{observed} = [draw,circle,outer sep=0.05cm,inner sep=0.3cm,line width=1, thick, draw=black!55, fill=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=5mm,width=3mm]}]

\begin{tikzpicture}
\node[observed] (node_observed) {\Huge $x$};
\node[hidden,above=1.5cm of node_observed] (node_hidden) {\Huge $z$};

\draw[arr] (node_hidden) -- (node_observed) {};

\node[fit=(node_hidden)(node_observed), draw, rounded corners] {};
\end{tikzpicture}
\end{document}

### 2017-01-20 slides/figures/vae_architecture.tex
updated & added figures

\documentclass{standalone}
\usepackage{tikz}
\definecolor{cdfdfdf}{RGB}{223,223,223}
\definecolor{cffffff}{RGB}{255,255,255}

\begin{document}
\usetikzlibrary{arrows.meta}

\begin{tikzpicture}[y=0.80pt, x=0.80pt, yscale=-1.000000, xscale=1.000000, inner sep=0pt, outer sep=0pt]
\begin{scope}[cm={{1.25,0.0,0.0,-1.25,(0.0,1052.3622)}}]
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,fill=cdfdfdf,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (9.9628,0.0000) .. controls
      (9.9628,5.5023) and (5.5023,9.9628) .. (0.0000,9.9628) .. controls
      (-5.5023,9.9628) and (-9.9628,5.5023) .. (-9.9628,0.0000) .. controls
      (-9.9628,-5.5023) and (-5.5024,-9.9628) .. (0.0000,-9.9628) .. controls
      (5.5023,-9.9628) and (9.9628,-5.5024) .. (9.9628,0.0000) -- cycle;
  \end{scope}
    \path[cm={{1.0,0.0,0.0,-1.0,(303.468,695.373)}},fill=black,nonzero rule]
      (0.0000,0.0000) node[above right] (text4171) {$x$};
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,fill=cffffff,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (9.9628,48.6708) .. controls
      (9.9628,54.1732) and (5.5024,58.6336) .. (0.0000,58.6336) .. controls
      (-5.5023,58.6336) and (-9.9628,54.1732) .. (-9.9628,48.6708) .. controls
      (-9.9628,43.1685) and (-5.5024,38.7081) .. (0.0000,38.7081) .. controls
      (5.5023,38.7081) and (9.9628,43.1685) .. (9.9628,48.6708) -- cycle;
  \end{scope}
    \path[cm={{1.0,0.0,0.0,-1.0,(303.945,744.043)}},fill=black,nonzero rule]
      (0.0000,0.0000) node[above right] (text4181) {$z$};
    \path[cm={{1.0,0.0,0.0,-1.0,(260.848,741.767)}},fill=black,nonzero rule]
      (0.0000,0.0000) node[above right] (text4187) {$\phi$};
    \path[cm={{1.0,0.0,0.0,-1.0,(345.199,742.798)}},fill=black,nonzero rule]
      (0.0000,0.0000) node[above right] (text4193) {$\theta$};
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,dash pattern=on 2.39pt off 2.39pt,line join=miter,line
      cap=butt,miter limit=10.00,line width=0.319pt] (-38.5088,48.6708) --
      (-10.9710,48.6708);
  \end{scope}
  \begin{scope}[shift={(295.52,746.258)}]
    \path[draw=black,fill=black,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (-5.2035,2.3356) --
      (0.2989,0.0000) -- (-5.2035,-2.3356) -- (-5.2035,2.3356) -- cycle;
  \end{scope}
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,line join=miter,line cap=butt,miter limit=10.00,line
      width=0.319pt] (38.5088,48.6708) -- (10.9710,48.6708);
  \end{scope}
  \begin{scope}[cm={{-1.0,0.0,0.0,-1.0,(317.462,746.258)}}]
    \path[draw=black,fill=black,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (-5.2035,2.3356) --
      (0.2989,0.0000) -- (-5.2035,-2.3356) -- (-5.2035,2.3356) -- cycle;
  \end{scope}
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,line join=miter,line cap=butt,miter limit=10.00,line
      width=0.319pt] (38.5088,45.5084) -- (7.1087,8.4005);
  \end{scope}
  \begin{scope}[cm={{-0.64792,-0.7657,0.7657,-0.64792,(313.6,705.988)}}]
    \path[draw=black,fill=black,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (-5.2035,2.3356) --
      (0.2989,0.0000) -- (-5.2035,-2.3356) -- (-5.2035,2.3356) -- cycle;
  \end{scope}
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,dash pattern=on 2.39pt off 2.39pt,line join=miter,line
      cap=butt,miter limit=10.00,line width=0.319pt] (-7.1856,7.1856) .. controls
      (-16.6444,16.6438) and (-16.6444,32.0271) .. (-7.7582,40.9132);
  \end{scope}
  \begin{scope}[cm={{0.7071,0.7071,-0.7071,0.7071,(298.733,738.5)}}]
    \path[draw=black,fill=black,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (-5.2035,2.3356) --
      (0.2989,0.0000) -- (-5.2035,-2.3356) -- (-5.2035,2.3356) -- cycle;
  \end{scope}
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,line join=miter,line cap=butt,miter limit=10.00,line
      width=0.319pt] (0.0000,38.5088) -- (0.0000,10.9710);
  \end{scope}
  \begin{scope}[cm={{0.0,-1.0,1.0,0.0,(306.491,708.558)}}]
    \path[draw=black,fill=black,line join=miter,line cap=butt,miter
      limit=10.00,nonzero rule,line width=0.319pt] (-5.2035,2.3356) --
      (0.2989,0.0000) -- (-5.2035,-2.3356) -- (-5.2035,2.3356) -- cycle;
  \end{scope}
    \path[cm={{1.0,0.0,0.0,-1.0,(308.318,675.918)}},fill=black,nonzero rule]
      (0.0000,0.0000) node[above right] (text4239) {N};
  \begin{scope}[shift={(306.491,697.587)}]
    \path[draw=black,line join=miter,line cap=butt,miter limit=10.00,line
      width=0.319pt] (16.5376,62.3527) -- (-16.5376,62.3527) .. controls
      (-18.7386,62.3527) and (-20.5228,60.5685) .. (-20.5228,58.3676) --
      (-20.5228,-21.2037) .. controls (-20.5228,-23.4046) and (-18.7386,-25.1888) ..
      (-16.5376,-25.1888) -- (16.5376,-25.1888) .. controls (18.7386,-25.1888) and
      (20.5228,-23.4046) .. (20.5228,-21.2037) -- (20.5228,58.3676) .. controls
      (20.5228,60.5685) and (18.7386,62.3527) .. (16.5376,62.3527) -- cycle;
  \end{scope}
\end{scope}

\end{tikzpicture}
\end{document}

### 2017-02-03 paper/chapters/introduction.tex
minor corrections (fig/eq uppercase)

@@
   \item \textbf{Images} consisting of millions of pixels can be interpreted as data vectors with each pixel representing an own dimension.
-  However, these dimensions are high correlated and structured, for example nearby pixels have with high probability similar values.
-  Looking at images of handwritten digits, the data can be reduced to a handful variables which may include the digit number, stroke width, position or size.
+  However, these dimensions are highly correlated and structured, for example nearby pixels have with high probability similar values.
+  Looking at images of handwritten digits, the data can be reduced to a handful of variables which may include the digit number, stroke width, position or size.
   Statistical models help us discover these relations and reduce them to representations in lower-dimensional space.
@@
 
-The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in figure \ref{fig:dgm}.
+The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure \ref{fig:dgm}.
 %as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.

### 2017-02-03 paper/chapters/sigmoid_belief_nets.tex
minor corrections (fig/eq uppercase)

@@
 \end{equation}
-In equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.
+In Equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.
 
-To draw samples from this network, it is possible to simply evaluate equation \ref{eq:sbn_node} for each visible node.
+To draw samples from this network, it is possible to simply evaluate Equation \ref{eq:sbn_node} for each visible node.
 However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\

### 2017-02-03 paper/chapters/vae.tex
minor corrections (fig/eq uppercase)

@@
 The probabilistic encoder generally implemented as feed-forward network receives input from the dataset and outputs parameters of a distribution.
-For example the parameters for a gaussian distribution with diagonal covariance, in which case the networks emits the mean $\mu$ and $\sigma$ used to define the gaussian distribution $\mathcal{N}(\mu,\sigma \times I)$, where $I$ is the identity matrix.\\
+For example the parameters for a gaussian distribution with diagonal covariance, in which case the network emits the mean $\mu$ and $\sigma$ used to define the gaussian distribution $\mathcal{N}(\mu,\sigma \times I)$, where $I$ is the identity matrix.\\
 
@@
 Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
-This architecture is shown in figure \ref{fig:vae_architecture} with a one-dimensonal gaussian prior.
+This architecture is shown in Figure \ref{fig:vae_architecture} with a one-dimensonal gaussian prior.
 \\
@@
 Using Markov chain Monte Carlo (MCMC) methods, which construct a Markov chain with the desired attributes matching those of the posterior.
-The advantage of this approach is that is guaranteed to asympotically approach the desired posterior, however it is difficult to reason about the convergence rate.
+The advantage of this approach is that it is guaranteed to asympotically approach the desired posterior, however it is difficult to reason about the convergence rate.
 For deep neural networks, running MCMC in the inner loop of optimization techniques is impractical.\\
@@
 \subsubsection{Kullback-Leibler Divergence}
-In order to approximate one distribution $Q$ to another distribution $P$, there needs to be a measurement how close $Q$ is to $P$.
-One commonly used measurement is called Kullback-Leibler divergence (short KL divergence) which enables the variational inference methods to choose appropiate distributions.
+In order to approximate one distribution $Q$ to another distribution $P$, there needs to be a measurement of how close $Q$ is to $P$.
+One commonly used measurement is called Kullback-Leibler (KL) divergence\cite{kl_div:1951} which enables the variational inference methods to choose appropiate distributions.
 %\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
 $\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
-For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
+The KL divergence from $Q$ to $P$ is defined in Equation \ref{eq:kl_div}.
 \begin{equation}
@@
 
-However we are not able to optimize equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
+However we are not able to optimize Equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
 The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
@@
 
-Computing the integral in equation \ref{eq:integral} requires the evaluation of every possible configuration over the hidden variables $z$ and is therefore intractable.\\\\
+Computing the integral in Equation \ref{eq:integral} requires the evaluation of every possible configuration over the hidden variables $z$ and is therefore intractable.\\\\
 
@@
 In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
-Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample first an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.
-Then we apply equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
+Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.
+Then we apply Equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
 \begin{equation}
@@
 In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}.
-Results of this experiments are shown in fig.
-\ref{fig:vae_experiments}.
+Results of this experiments are shown in Figure \ref{fig:vae_experiments}.
 %The model has been trained on the MNIST dataset \cite{mnist} with satisfactory generative performance as shown in figure \ref{fig:vae_experiments}

### 2017-02-03 paper/chapters/gan.tex
todo: gan theoretical foundation and minor corrections

@@
 As each party plays to maximize its own success, the discriminator will try to maximize the number of samples which got discriminated correctly.
-On the other side, $G$ tries to produce samples which fool the $D$ often.
+On the other side, the generator tries to produce samples aiming to fool the discriminator.
 
@@
 
-Using the analogy of counterfeiters and police can help understand the intuition behind game between $G$ and $D$ better.
-Counterfeiters will try to produce fake money looking to fool the police into thinking it is real money.
-The police however is able to distinguish badly made fake money from real money easily and as such rejects the fake money with high confidence.
-This rejection in turn triggers the counterfeiters to improve their forgery process try to imitate real money even better.
+%Using the analogy of counterfeiters and police can help understand the intuition behind game between $G$ and $D$ better.
+%Counterfeiters will try to produce fake money looking to fool the police into thinking it is real money.
+%The police however is able to distinguish badly made fake money from real money easily and as such rejects the fake money with high confidence.
+%This rejection in turn triggers the counterfeiters to improve their forgery process try to imitate real money even better.
 Thus they are able to improve the quality of the faked money.
@@
 
-The game-theoretic solution to the emerging situation where neither counterfeiters as generator nor police as discriminator can improve its own success with respect to the outcome of the game gets called \emph{Nash equilibrium}\cite{game_theory:1994}.
-
+\paragraph{Solution}
+The game-theoretic solution to the emerging situation where neither $G$ nor $D$ can improve its own success with respect to the outcome of the game gets called \emph{Nash equilibrium}\cite{game_theory:1994}.
 In the case of GANs, this situation arises when the generator produces samples which get rejected with probability $p=0.5$ by the discriminator.
-In the case of such equilibrium, the generator produces good enough samples to fool the generator in about half cases.
+In the case of such equilibrium, the generator produces good enough samples to fool the generator in about half cases.\\\\
+\todo[inline]{Add some text showing the theoretical foundation of GANs}
 

### 2017-02-03 paper/chapters/vae/cvae.tex
improve iwae citation (in title, oops)

@@
+\todo[inline]{Add citations to \cite{cvae:2014} and \cite{cvae_2:2015}}
 Instead of using VAEs in a completely unsupervised manner on a unlabeled dataset, it is also possible to learn from datasets where a small subset of datapoints has corresponding labels while most datapoints are still unlabeled.
@@
 
-To adjust the VAE model to the semi-supervised case, we have to add an random variable $y$ corresponding to the corresponding label if available.
+To adjust the VAE model to the semi-supervised case, we have to add a random variable $y$ corresponding to the label if available.
 Thus we can extend the distributions as shown below.

### 2017-02-03 paper/chapters/gan.tex
added TODO for modified objective

@@
 
-However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning where the discriminator is able to reject samples from the generator with high confidence.\\
-Therefore is in practice often equation \ref{eq:l_g_2} instead of equation \ref{eq:l_g_1} used to optimize the generator.\\
+However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning \todo[inline]{early in training!
+show difference.} where the discriminator is able to reject samples from the generator with high confidence.\\
+Therefore is equation \ref{eq:l_g_2} more commonly used in practice \todo[inline]{citations! dcgan? improved gan?}.
 

### 2017-02-03 paper/chapters/gan.tex
further minor corrections

@@
 Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.\\
-Figure \ref{fig:gan_a} summarizes the described architecture with the according gradient flow used to update $G$ and $D$.\\
+Figure \ref{fig:gan_a} summarizes the described architecture with the respective gradient flow used to update $G$ and $D$.\\
 %Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).
@@
   \small
-  The line named $z$ denoted the noise prior, while the arrows to the line indicated by $x$ is the transformation learned by the generator network.
+  The line named $z$ denotes the noise prior, while the arrows to the line indicated by $x$ is the transformation learned by the generator network.
 Three different distributions are showed, the solid line represents the generated sample in each step, while the dots are sampled from the data distribution $p_{data}(x)$.
@@
 The GAN framework can be fully trained with stochastic backpropagation due to $V(G,D)$ being differentiable with $G$ and $D$ implemented as (deep) neural nets.
-Figure \ref{fig:gan_training} illustrates the ideal training procedure for learning of an 1D gaussian.
+Figure \ref{fig:gan_training} illustrates the ideal training procedure for learning of an one-dimensional distribution.
 Training will switch between training the generator and discriminator, where the other party is fixed respectively.

### 2017-02-03 paper/chapters/gan.tex
applied corrections to GAN

@@
 The exact details of the architecture is beyond the scope of this article, however in addition to the impressive generative performance the authors have shown that the model has learned to produce representional structures in latent space.
-Furthermore, as shown in figure \ref{fig:dcgan_vector} they were able to perform vector artithmetic in the latent space with results indicating that their model was able to arrange the latent space in an useful way.
+Furthermore, as shown in Figure \ref{fig:dcgan_vector} they were able to perform vector artithmetic in the latent space with results indicating that their model was able to arrange the latent space in a useful way.
+\todo[inline]{Elaborate on how DCGAN arranges latent space}
 
@@
 \end{figure}
-Laplacian Generative Adversarial Networks (LAPGAN\cite{lapgan:2015}) is an extension to GANs applied to images using laplacian pyramids \cite{laplacian:1983}.
+Laplacian Generative Adversarial Networks (LAPGAN) \cite{lapgan:2015} are an extension to GANs applied to images using laplacian pyramids \cite{laplacian:1983}.
 Instead of using a single generator network, LAPGAN uses a series of deep convolutional networks as generators where each generator receives the output of the previous generator.
@@
 These steps are repeated for other steps, downscaling the image.
-In the last stage, a traditional generator and discriminator is used for the $8x8$ image.
+In the last stage, a traditional generator and discriminator is used for the $8\times8$ image.
 \paragraph{Sampling} uses a conditional GAN to produce high-resolution images through different stages of deep generator networks.
@@
 As GANs have no direct measurement of the quality of generations due to their lack of unified objective function, human evaluation of the authenticity of generated images is often used for qualitively evaluation.
-LAPGAN has pushed the state of the art significantly, with generated images which are high-resolution in comparison to earlier methods and arguably high variety of generated images as shown in figure \ref{fig:lapgan}.
+LAPGAN has pushed the state of the art significantly, with generated images which are high-resolution in comparison to earlier methods and arguably high variety of generated images as shown in Figure \ref{fig:lapgan}.
 Furthermore, humans have mistaken the generated images for real images 40\% of the time.

### 2017-02-27 paper/chapters/introduction.tex
removed quote
added caption2img example, further applications (todo)

@@
 
-"What I cannot create, I do not understand." -- Richard Feynman \\\\
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{media/caption2img}
+  \caption[Caption to Image]{Images generated given a captioning text \cite{gan_t2i:2016}}
+  \label{fig:caption2img}
+\end{figure}
+While sampling from generative models has by itself interesting applications, such as producing images given a caption as seen in Figure \ref{fig:caption2img}, there are many applications covering various different areas.
+A subset of these applications include
+image denoising, % TODO: citation
+image impainting, % TODO: citation
+text-to-speech, % TODO: citation, e.g. WaveNet
+machine translation, % (neural?) TODO: e.g. NMT w/ VAE
+and semi-supervised learning, for example discussed in \ref{ssub:cvae}.
+\\\\
 
-Directed generative models are a wide class of models used for generating samples from an unknown distribution $p_{data}$.
-%Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
-The presented models in this article will approximate $p_{data}$ with $p_{model}$ which enables us to draw samples from the model.
-Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\
+%Directed generative models are a wide class of models used for generating samples from an unknown distribution $p_\mathrm{data}$.
+%The models presented in this article will approximate $p_\mathrm{data}$ with $p_\mathrm{model}$ which enables us to draw samples from the model.
+%Given this assumption, samples from a dataset following $p_\mathrm{data}$ should have high likelihood in $p_\mathrm{model}$ too.\\\\
+
+Generative models are in particular interesting as they
 

### 2017-02-27 paper/chapters/gan.tex
fixed captions (short caption)

@@
 \includestandalone[width=\textwidth,mode=buildnew]{media/gan_conceptual}
-  \caption{Architecture used for generative adversarial networks}
+  \caption[GAN Architecture]{Architecture used for generative adversarial networks}
   \label{fig:gan_a}
@@
   \includegraphics[width=\linewidth]{media/gan_training-crop}
-  \caption{Learning a GAN on a one-dimensional distribution from the original paper\cite{gan:2014}}
+  \caption[Learning Procedure in a GAN]{Learning a GAN on a one-dimensional distribution from the original paper\cite{gan:2014}}
   \label{fig:gan_training}
@@
   \includegraphics[width=\linewidth]{media/dcgan_vector}
-  \caption{Vector arithmetic performed in latent space $z$, from DCGAN paper\cite{dcgan:2015}}
+  \caption[Vector arithmetic in DCGAN]{Vector arithmetic performed in latent space $z$, from DCGAN paper\cite{dcgan:2015}}
   \label{fig:dcgan_vector}
@@
   \includegraphics[width=\linewidth]{media/bedroom_scenes-crop}
-  \caption{Bedroom scenes generated by the DCGAN model, from the original paper\cite{lapgan:2015}}
+  \caption[Generated Samples by DCGAN]{Bedroom scenes generated by the DCGAN model, from the original paper\cite{lapgan:2015}}
   \label{fig:lapgan}

### 2017-02-27 paper/chapters/vae.tex
replace explicit kl div eq by expectation
improve iwae citation (in title, oops)
fixed captions (short caption)

@@
   \includestandalone[width=\linewidth,mode=buildnew]{media/vae_conceptual}
-  \caption{simplified VAE architecture using a gaussian prior on $z$}
+  \caption[VAE architecture]{simplified VAE architecture using a gaussian prior on $z$}
   \label{fig:vae_architecture}
@@
 \begin{equation}
-  \label{eq:kl_div_disc}
-  \mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
+  \label{eq:kl_div}
+  \mathcal{D}_{\mathrm{KL}}(P || Q) = \mathbb{E}_{x \sim P(x)}\bigg[log \frac{P(x)}{Q(x)}\bigg]
 \end{equation}
-For continuous distributions $P$ and $Q$ the KL divergence is defined similarly in equation \ref{eq:kl_div_cont}.
-\begin{equation}
-  \label{eq:kl_div_cont}
-  \mathcal{D}_{\mathrm{KL}}(P || Q) = \int P(i) \log \frac{P(i)}{Q(i)} di
-\end{equation}
+\todo[inline]{Add some description to the equation above, logarithmic difference between P and Q, bla bla expectation}
+%For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
+%\begin{equation}
+  %\label{eq:kl_div_disc}
+  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
+%\end{equation}
+%For continuous distributions $P$ and $Q$ the KL divergence is defined similarly in equation \ref{eq:kl_div_cont}.
+%\begin{equation}
+  %\label{eq:kl_div_cont}
+  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \int P(i) \log \frac{P(i)}{Q(i)} di
+%\end{equation}
 
@@
   \end{tabular}
-  \caption{reparameterization trick}
+  \caption{Reparameterization Trick}
   \label{fig:rep_trick}
@@
 
-\subsubsection{Conditional VAE \cite{cvae:2014} \cite{cvae_2:2015}}
+\subsubsection{Conditional VAE}
 \label{ssub:cvae}

### 2017-02-27 paper/chapters/vae.tex
clarification to rep trick (and added citation)

@@
 % http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
-To be able to backpropagate the loss function through the VAE, it has to be differentiable and deterministic.
-However, because we only sample from the constructed distribution $q_\phi(z|x)$, we cannot backpropagate through the node due to the stochastic nature.
-In order to circumvent this restriction, the so-called "reparameterization trick" is applied.
+To be able to apply the backpropagation\cite{backprop:1988} algorithm to the loss function of the VAE without any injected noise it has to be both differentiable and deterministic.
+However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
+This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
+In order to circumvent this restriction, the so-called ``reparameterization trick'' (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) is applied.
 Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.
@@
 
-Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different probability distributions on $z$.
+Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different prior probability distributions on $z$.
 %The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.

### 2017-02-28 paper/chapters/abstract.tex
change abstract to be an actual abstract
improve abstract (more text, what we discuss, etc.)

@@
-\section{Abstract}
-\label{sec:abstract}
+\begin{abstract}
 %One of the key aspects of human intelligence is the ability to construct accurate and generalized internal representations by observing high-dimensional data.
@@
 Similar to how humans learn to generalize from sensory input, we discuss several models which can learn compact, internal representational structures.
-In this article, we will focus on two recently proposed architectures, variational autoencoders, and generative adversarial networks.
+After discussing generative modeling in general and its applications, we then introduce the sigmoid belief network which is a simple generative architecture.
+Then we focus on two more recently proposed architectures: the variational autoencoders (VAE), and generative adversarial networks (GAN).
+We discuss the key idea of both models and their respective learning and training procedure, including variational inference and adversarial training for VAE and GAN respectively.
+Additionally to the originally proposed models, we summarize several extensions and present experimental results for the variational autoencoder.
+
+%In this article, we will focus on two recently proposed architectures, variational autoencoders, and generative adversarial networks.
 
@@
 %Additionally to exploring these models and theoretical foundation, we provide an intuitive and simple application of a GAN.
+
+\end{abstract}

### 2017-02-28 paper/chapters/introduction.tex
fixed captions (short caption)
better caption for caption2img
more intro

@@
 
+%%%
+%%% intuition // generative vs discriminative
+%%%
+Intuitively, generative modeling is based on the idea that in order to produce data similar to previously seen data it helps to understand the data first~\cite{salakhutdinov:2009}.
+The case of deep learning suggests the use of deep, hierarchical architectures to capture dependencies within the data and reduce the high-dimensional input~\cite{reducing_dim:2006}.
+
+In contrast to discriminative models which handle the conditional distribution $p(y|x)$ of the labels $y$ given observations $x$, generative models use the joint distribution $p(x,y)$ over both labels and observations to describe the data~\cite{learning_dbn:2006}.\\\\
+
+%TODO more:
+% motivation
+
+%%%
+%%% applications
+%%%
+
 \begin{figure}[h]
   \includegraphics[width=\linewidth]{media/caption2img}
-  \caption[Caption to Image]{Images generated given a captioning text \cite{gan_t2i:2016}}
+  \caption[Caption-to-Image Samples]{Image Synthesis from provided caption~\cite{gan_t2i:2016}}
 \label{fig:caption2img}
 \end{figure}
-While sampling from generative models has by itself interesting applications, such as producing images given a caption as seen in Figure \ref{fig:caption2img}, there are many applications covering various different areas.
+%Generative modeling is a wide field involving capturing an unknown data distribution.\\
+
+% TODO (kinda similar to the first sentences in the introduction (and even abstract as well)
+Following the idea of deep models to learn hierarchical representations, these results should be applicable to generative models as well.
+% "learning deep generative models" gives these arguments!
+
+Sampling from generative models generally means sampling from a distribution which the model either implicitly or explicitly has learned.
+Interesting applications around sampling from these models include generating images given a caption as seen in Figure~\ref{fig:caption2img} or sampling 3D models from a learned latent representation~\cite{gan_3d:2016}.
+As generative models often have internal representations of the data distribution, these learned representational structures or features can be used in models built on top.
 A subset of these applications include
-image denoising, % TODO: citation
-image impainting, % TODO: citation
+image denoising~\cite{generative_denoising:2013}, % TODO: citation
 text-to-speech, % TODO: citation, e.g. WaveNet
 machine translation, % (neural?) TODO: e.g. NMT w/ VAE
-and semi-supervised learning, for example discussed in \ref{ssub:cvae}.
+and semi-supervised learning, for example discussed in~\ref{ssub:cvae}.
 \\\\
+
+\newpage
+
+%%%
+%%% history // other models
+%%%
+\todo[inline]{History}
+% traditional models (HMM, GMM)
+Traditional models include hidden Markov models used in speech recognition~\cite{hmm:1989} and Gaussian mixture models.
+
+% undirected networks (Boltzmann machine)
+Another class in generative models were undirected networks, for example various extensions of the Boltzmann machine~\cite{bm:1985} such as the restricted Boltzmann machine (RBM) and the deep Boltzmann machine~\cite{dbm:2009} were used.\\
+
+% directed networks (SBN, DBN)
+For directed networks, the sigmoid belief network (SBN) is an early model which is presented in~\ref{sec:sbn}.
+SBN has also been extended using an restricted Boltzmann machine resulting in Deep Belief Nets (DBN)~\cite{learning_dbn:2006}.
 
@@
 
-Generative models are in particular interesting as they
-
-\paragraph{Short-term applications}
 \begin{itemize}
@@
   \item \textbf{Semi-supervised learning} on datasets with a small amount of unlabelled data and a large amount of labeled data to improve model performance.
-Examples include the conditional variational autoencoder discussed in \ref{ssub:cvae}.
+Examples include the conditional variational autoencoder discussed in~\ref{ssub:cvae}.
   \item \textbf{Data exploration and understanding} is often needed especially for high-dimensional and complex data.
 By using models which reduce the dimensions by selecting only important features for generative modeling can help to improve data understanding.
-Figure \ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
+Figure~\ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
 \end{itemize}
@@
 
-The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure \ref{fig:dgm}.
+The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure~\ref{fig:dgm}.
 %as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.
@@
 \includegraphics{media/directed_graphical_model}
-  \caption{Directed graphical model}
+  \caption[Graphical Model]{Directed Graphical Model}
 \label{fig:dgm}
@@
   \small
+  \todo[inline]{arrow indicate stochastic dependency}
   The latent variable $z$ is a stochastic dependency of the observed data $x$.

### 2017-03-01 paper/chapters/sigmoid_belief_nets.tex
add todo for SBN
fixed captions (short caption)
describe the sigmoid func and weights using 'conditions'

@@
 \input{media/sbn_architecture}
-  \caption{Conceptual architecture of belief nets}
+  \caption[Architecture of Sigmoid Belief Nets]{Conceptual architecture of belief nets}
   \label{fig:sbn_arch}
@@
 \end{equation}
-In Equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.
+where 
+\begin{conditions}
+  \sigma(x) & $\frac{1}{1 + e^{-x}}$ (sigmoid function)\\
+  W_{i,j} & connection weights from state $s_i$ to $s_j$\\
+\end{conditions}
 
-To draw samples from this network, it is possible to simply evaluate Equation \ref{eq:sbn_node} for each visible node.
+To draw samples from this network, it is possible to simply evaluate Equation~\ref{eq:sbn_node} for each visible node.
 However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
@@
 Especially in deep networks, this kind of computation is infeasible as it requires computing exponentially many configurations.\\
+
+% TODO: connection to DBN, layer-wise greedy training
+% Hinton et al. (2006)
 

### 2017-03-01 paper/chapters/sigmoid_belief_nets.tex
fix order of indices
fixed bad formulation

@@
   \label{eq:sbn_node}
-p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
+p(s_i) = \sigma\bigg(\sum_{j<i}W_{j,i}s_j+b_i\bigg)
 \end{equation}
@@
   \sigma(x) & $\frac{1}{1 + e^{-x}}$ (sigmoid function)\\
-  W_{i,j} & connection weights from state $s_i$ to $s_j$\\
+  W_{j,i} & connection weights from state $s_j$ to $s_i$\\
 \end{conditions}
@@
 However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
-Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
+Learning means in this context to adjust the weights and biases of the network to be more likely to generate data similar to samples from the given training dataset.
 Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the states in the visible layer.\\

### 2017-03-01 paper/chapters/sigmoid_belief_nets.tex
fixed formulation to be conservative

@@
 
-Sigmoid Belief Networks (SBN) were proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985} shown in figure \ref{fig:sbn_arch}
+Sigmoid Belief Networks (SBN) were described in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985} shown in Figure \ref{fig:sbn_arch}
 In the case of SBNs however, all states are binary and the activation function is the sigmoid function.

### 2017-03-01 paper/chapters/vae.tex
more intro
fixed that formulation, I don't even understand Oo
rewritten approximated inference chapter

@@
 
-Properly defined emits the described probabilitic encoder variational parameters $\phi$ used to define the probability distribution $q_\phi(z|x)$.
+Such function approximator will emit the variational parameters $\phi$ of the probability encoder distribution $q_\phi(z|x)$.
 Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
@@
 
-\subsection{Variational Inference}
-Variational inference (VI) can be used to turn the inference problem into an optmization problem which we are able to approximate using gradient-based methods.
-%The calculus of variations is used to perform optimization on a family of functions.
-%As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems as seen in \ref{eq:intractable_posterior}.
+\subsection{Approximated Inference}
+A problem which often arises in directed generative models is the \emph{inference problem}.
+As discussed in the introduction~\ref{sec:introduction}, we assume that the observations are stochastically dependent on some latent variables.
+Inference is called the process of coming up with values of $z$ which can produce the observations, which formally means evaluating $p(z|x)$ where $z$ are the latent variables and $x$ observations.
+\begin{equation}
+  p(z|x) = \frac{p(x|z)p(z)}{p(x)}\\
+  \label{eq:intractable_posterior}
+\end{equation}
+where
+\begin{conditions}
+  p(z|x) & posterior distribution\\
+  p(x|z) & likelihood of $x$ given $z$\\
+  p(z)   & prior distribution of $z$\\
+\end{conditions}
+Using Bayes's rule we can expand the posterior distribution $p(z|x)$ as shown in Equation~\ref{eq:intractable_posterior}.
+Note that to evaluate the denominator $p(x)$, all hidden variables need to be marginalized: $p(x) = \int p(x,z) dz$.\\
+Due to the possibly large number of hidden variables, this integral can be high-dimensional and therefore difficult to calculate.
+This integral can be difficult to compute and needs to be approximated.
+However, for a large number of hidden variables, this integral is intractable to compute exactly and therefore needs to be approximated~\cite[p.~461-462]{bishop:2006}.
+In order to approximate those intractable posterior distributions, we discuss two commonly used techniques:  Markov chain Monte Carlo and variational inference.
 
-These intractable posteriors unfortunately arise in almost all interesting models and there are mostly two different approaches to approximate the posterior distribution.
-% MCMC
-Using Markov chain Monte Carlo (MCMC) methods, which construct a Markov chain with the desired attributes matching those of the posterior.
-The advantage of this approach is that it is guaranteed to asympotically approach the desired posterior, however it is difficult to reason about the convergence rate.
-For deep neural networks, running MCMC in the inner loop of optimization techniques is impractical.\\
+\paragraph{Markov chain Monte Carlo}
+MCMC methods are a class of techniques which can be used to sample from a large class of distribution where direct sampling is not applicable~\cite[Chapter~11.2]{bishop:2006}.
+For the variational autoencoder, where we want to approximate the posterior distribution $p(z|x)$ a method called \emph{Gibbs sampling}~\cite[Chapter~11.3]{bishop:2006} can be used to obtain samples from that conditional distribution by constructing a Markov chain with the desired equilibrium distribution.
+%
+One advantage of MCMC methods is that the resulting distribution is proven to be exactly the desired distribution~\cite[Chapter~11.3]{bishop:2006}.
+However, due to the problems of mixing between mixing modes~\cite[Chapter~17.5]{deeplearning:2016} and slow convergence rate\cite[p.~603]{deeplearning:2016} these methods are impractical for usage in the inner loop of optimization techniques.
+
+\paragraph{Variational Inference}
 % variational inference
-Using variational inference is the second approach, used in the variational autoencoder, which picks a specific distribution restricted to a family of distributions matching the posterior most closely.
-Advantages using this approach rather than MCMC is the greater speed at which the computation is performed.
-However it is not guaranteed to approximate the posterior accurately, if the true posterior can not be approximately expressed as a distribution of the family.
+Due to the possibly slow convergence of MCMC methods, variational inference (VI)~\cite[Chapter~19.4]{deeplearning:2016} is used in the variational autoencoder.
+VI works by picking a specific distribution restricted to a family of distributions which matches the posterior most closely.
+This means that VI is used to turn the inference problem into an optimization technique, allowing for gradient-based optimization.
+%TODO: more on advantages and disadvantages
+Using VI yields better performance at the sacrifice of theoretical guarantee of matching the posterior distribution.
 
@@
     \end{subfigure}
-    \caption{Variational autoencoder learning performance}
+    \caption{Manifold Learning in the VAE}
     \label{fig:vae_experiments}

### 2017-03-01 paper/chapters/sigmoid_belief_nets.tex
minor fixed (killed warnings)

@@
 
-Sigmoid Belief Networks (SBN) were described in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985} shown in Figure \ref{fig:sbn_arch}
+Sigmoid Belief Networks (SBN) were described in 1992 by R.M. Neal~\cite{neal:1992} representing a specific type of bayesian networks~\cite{pearl:1985} shown in Figure~\ref{fig:sbn_arch}
 In the case of SBNs however, all states are binary and the activation function is the sigmoid function.
@@
 
-To draw samples from this network, it is possible to simply evaluate Equation~\ref{eq:sbn_node} for each visible node.
+To draw samples from this network, it is possible to evaluate Equation~\ref{eq:sbn_node} for each visible node.
 However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
@@
 Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
-Especially in deep networks, this kind of computation is infeasible as it requires computing exponentially many configurations.\\
+Especially in deep networks, this kind of computation is infeasible as it requires computing exponentially many configurations.
 

### 2017-03-01 paper/chapters/vae.tex
fixed weird formulation

@@
 This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
-In order to circumvent this restriction, the so-called ``reparameterization trick'' (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) is applied.
+In order to circumvent this restriction, the reparameterization trick (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) can be applied.
 Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.

### 2017-03-01 paper/chapters/vae/cvae.tex
reformatted CVAE

--- 2017-01-06 paper/chapters/vae/cvae.tex
@@
 Instead of using VAEs in a completely unsupervised manner on a unlabeled dataset, it is also possible to learn from datasets where a small subset of datapoints has corresponding labels while most datapoints are still unlabeled.
-Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly.\\
-
-To adjust the VAE model to the semi-supervised case, we have to add an random variable $y$ corresponding to the corresponding label if available.
-Thus we can extend the distributions as shown below.
-
+Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly~\cite{cvae:2014}\cite{cvae_2:2015}.\\
+%
+To adjust the VAE model to the semi-supervised case, we have to add a random variable $y$ corresponding to the label if available.\\\\
+First, instead of using samples from the dataset alone we are now also conditioning on $y$, formally 
+$x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
+Then the probabilistic encoder and decoder can be written as $q_\theta(z|x,y)$ and $p_\phi(x|y,z)$ respectively.\\
+%
 %The modified setup of the semi-supervised case for the variational autoencoder can be formalized as follows.
@@
 %\begin{equation}
-
+%
 %\end{equation}
@@
 %A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks
-
+%
 %Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.
-
+%
 %To support conditional distributions, the following relations have to be rewritten:
-\begin{itemize}
-\item \textbf{Sampling from data space}: $x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
-\item \textbf{Probabilistic encoder}: $q_\theta(z|x,y)$ instead of just $q_\theta(z|x)$.\\
-\item \textbf{Probabilistic decoder}: $p_\phi(x|y,z)$ instead of just $p_\phi(x|z)$.\\
-\end{itemize}
-
-These modifications allows us to formalize the new objective functionas shown in equations \ref{eq:cvae_labels} and \ref{eq:cvae_unknown}.
+%
+These modifications allows us to reformulate the objective function $\mathcal{L}$ from Equation~\ref{eq:vae_obj} in equations \ref{eq:cvae_labels} and \ref{eq:cvae_unknown}.
 

### 2017-03-01 paper/chapters/gan.tex
added name for name

@@
 Essentially, BN works by normalizing each sample from a minibatch to the statistics of the whole batch.
-To improve training performance of GANs, a modified version of batch normalization has been proposed \cite{improved_gan:2016} for training of the generator.
+To improve training performance of GANs, a modified version of batch normalization called virtual batch normalization has been proposed \cite{improved_gan:2016} for training of the generator.
 Instead of normalizing each input $x$ to their corresponding minibatch, one reference batch is chosen prior to training which will be used to normalize all further input.

### 2017-03-01 paper/chapters/gan.tex
removed irrelevant, misleading info
reformulation

@@
 In computer vision, features are often used to identify relevant pieces of information in images for recognition and classification.
-They may refer to a set of points or edges in an image, which exactly is not clearly defined.
 With the recent advances in classification tasks performed by deep convolutional networks, automatically learned features have been a large part of that success.
-Feature matching is an technique designed to improve the learning procedure of GANs where not directly $\mathcal{L}_g$ is maximized but instead also learn the discriminator on intermediate layer activation outputs.
+Feature matching is an technique designed to improve the learning procedure of GANs where $\mathcal{L}_g$ is not not directly maximized but where the discriminator also learns on intermediate layer activation outputs.
 Using this modified procedure aids the generator to output data matching the statistics of the input data more closely.

### 2017-03-01 paper/chapters/conclusion.tex
restructured conclusion/summary

@@
 
-In this article we have taken a look at different directed models which perform generative modeling using deep neural networks.
-Variational autoencoders as well as generative adversarial networks have shown to produce good results using large datasets with unsupervised learning.
-Deep generative models have shown to learn rich internal representation of high-dimensional data in order to generate samples.
-Going further, as intelligent agents will need to have accurate and generalized internal representations inferred through sensory input to reason about the world, generative models are are likely to be the cornerstones of more advanced intelligent systems in the future.
+In this article we have taken a look at different directed models which perform generative modeling using deep neural networks.\\
+After a short overview over traditional generative models and models not covered in this article, we have taken a look at a wide area of applications using generative models.\\\\
+%
+As one of the first basic generative models, the sigmoid belief net lays the foundation for directed generative models and displays its weaknesses during inference and training.\\
+The variational autoencoders have been discussed in more detail using variational inference to turn the inference problem into an optimization problem and solving it there.
+VAE and its various extensions have been shown to be a flexible tool applicable for different tasks.\\
+%
+Generative adversarial networks introduced in Section~\ref{sec:gan} uses a game-theoretic approach to generative modeling. In recent years, GANs have been extended of which we discussed a subset.\\\\
+%
+Deep generative models have shown the capability to learn rich internal representation of high-dimensional data in order to generate samples using recent advances in deep learning.
+Going further, as intelligent agents will need to have accurate and generalized internal representations inferred through sensory input to reason about the world, generative models are likely to be the cornerstones of more advanced intelligent systems in the future.
 
-Both models explained in detail have been proposed in recent years with many proposed extensions which were not covered in this article, thus further improvements are likely to happen.
+%Both models explained in detail have been proposed in recent years with many proposed extensions which were not covered in this article, thus further improvements are likely to happen.
 

### 2017-03-01 paper/chapters/gan.tex
removed todo
minor changes to GAN part

@@
 %%% when introduced, game theory approach
-Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modeling.
+Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al.
+\cite{gan:2014} based on a game theoretic approach to generative modeling~\cite[Chapter~20.10.4]{deeplearning:2016}.
 %which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
@@
 
-Another formulation of the same game is to define the zero-sum game between the generative model and its adversary, the discriminator.
+Another formulation of the same game is to define the zero-sum game between the generative model and its adversary, the discriminator~\cite[Chapter~20.10.4]{deeplearning:2016}.
 The payoff of the discriminator can be described as the value function $V(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-V(G,D)$.\\
@@
 In the common case where both function approximators $G$ and $D$ are implemented using deep neural networks, GANs have been shown to be difficult to train using gradient descent.
-In fact, there is no guarantee that GANs will converge in the case of using non-convex functions to represent $G$ or $D$.
+In fact, there is no guarantee that GANs will converge in the case of using non-convex functions to represent $G$ or $D$~\cite[p.~700-701]{deeplearning:2016}.
 
@@
 %TheseAdditionally to these hypothetical thoughts, \cite{gan_distinguish_crit:2014} have shown that these methods may fail to converge.
-Acknowledging these problems, Salimans et al.\cite{improved_gan:2016} has studied the stability and performance of GAN in detail, resulting in a number of practical methods improving performance of the training procedure of GANs.
+
+Acknowledging these problems, Salimans et al.\cite{improved_gan:2016} have studied the stability and performance of GAN in detail, resulting in a number of practical methods improving performance of the training procedure of GANs.
 We will take a look at a selection of these methods in the following, focussing on deep neural networks as approximators in both $D$ and $G$.
@@
 Furthermore, as shown in Figure \ref{fig:dcgan_vector} they were able to perform vector artithmetic in the latent space with results indicating that their model was able to arrange the latent space in a useful way.
-\todo[inline]{Elaborate on how DCGAN arranges latent space}
+%\todo[inline]{Elaborate on how DCGAN arranges latent space}
 

### 2017-03-03 paper/chapters/abstract.tex
last-minute changes

@@
 
-Deep directed generative models leverage the possibilities of deep learning to learn rich, hierarchical structures from datasets in a possibly unsupervised context.
+Deep directed generative models leverage the possibilities of deep learning to learn rich, hierarchical structures from datasets in an unsupervised context.
 These models can be used to understand high-dimensional, real-world data and transform them into compact representations.
 Similar to how humans learn to generalize from sensory input, we discuss several models which can learn compact, internal representational structures.
-After discussing generative modeling in general and its applications, we then introduce the sigmoid belief network which is a simple generative architecture.
-Then we focus on two more recently proposed architectures: the variational autoencoders (VAE), and generative adversarial networks (GAN).
-We discuss the key idea of both models and their respective learning and training procedure, including variational inference and adversarial training for VAE and GAN respectively.
+After discussing generative modeling in general and its applications, we introduce the sigmoid belief network which is a simple generative architecture.
+We then focus on two more recently proposed architectures: variational autoencoders, and generative adversarial networks.
+We discuss the key idea of both models and their respective architecture and training procedure, including variational inference and adversarial training for the variational autoencoder and generative adversarial network respectively.
 Additionally to the originally proposed models, we summarize several extensions and present experimental results for the variational autoencoder.

### 2017-03-03 paper/chapters/conclusion.tex
last-minute changes

@@
 
-In this article we have taken a look at different directed models which perform generative modeling using deep neural networks.\\
+In this article, we have taken a look at various directed models which perform generative modeling using deep neural networks.\\
 After a short overview over traditional generative models and models not covered in this article, we have taken a look at a wide area of applications using generative models.\\\\

### 2017-03-03 paper/chapters/gan.tex
last-minute changes

@@
 Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al.
-\cite{gan:2014} based on a game theoretic approach to generative modeling~\cite[Chapter~20.10.4]{deeplearning:2016}.
+\cite{gan:2014} based on a game-theoretic approach to generative modeling~\cite[Chapter~20.10.4]{deeplearning:2016}.
 %which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).
@@
 One player is called the generator $G$ which tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually real data.
-It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.
+It is important to note that the generator gets to know whether the discriminator accepted the generated output as real or rejected it as fake.
 
@@
 \paragraph{Solution}
-The game-theoretic solution to the emerging situation where neither $G$ nor $D$ can improve its own success with respect to the outcome of the game gets called \emph{Nash equilibrium}\cite{game_theory:1994}.
+The game-theoretic solution to the emerging situation where neither $G$ nor $D$ can improve its own success with respect to the outcome of the game gets called Nash equilibrium\cite{game_theory:1994}.
 In the case of GANs, this situation arises when the generator produces samples which get rejected with probability $p=0.5$ by the discriminator.
 In the case of such equilibrium, the generator produces good enough samples to fool the generator in about half cases.\\\\
-\todo[inline]{Add some text showing the theoretical foundation of GANs}
 
@@
 The payoff of the discriminator can be described as the value function $V(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-V(G,D)$.\\
-This formulation allows us to define the reached nash equlibium in equation \ref{eq:zerosum_convergence}.
+This formulation allows us to define the reached nash equilibrium in equation \ref{eq:zerosum_convergence}.
 \begin{equation}
@@
 
-However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning \todo[inline]{early in training!
-show difference.} where the discriminator is able to reject samples from the generator with high confidence.\\
-Therefore is equation \ref{eq:l_g_2} more commonly used in practice \todo[inline]{citations! dcgan? improved gan?}.
+However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning where the discriminator is able to reject samples from the generator with high confidence.
+Therefore is Equation \ref{eq:l_g_2} advised to be used instead of Equation \ref{eq:l_g_1}~\cite{gan:2014}.\\
 
@@
   (c) After a step of learning $G$ with fixed $D$ has the generator learned to produce samples more similar to $p_{data}(x)$.\\
-  (d) The model has reached the nash equilibrium where $p_{model} = p_{data}$ and the discriminator can only distinguish between real and fake sample with $p=0.5$.
+  (d) The model has reached the nash equilibrium where $p_{model} = p_{data}$ and the discriminator can only distinguish between real and fake samples with $p=0.5$.
 \end{figure}
 
-Combining the two objectives into one value function yields equation \ref{eq:minimax_value} which is used to define minmax-game played by the two networks.
+Combining the two objectives into one value function yields equation \ref{eq:minimax_value} which is used to define the minmax-game played by the two networks.
 \begin{equation}
@@
 %\paragraph{Theoretical Foundation}
-Using a model with infinite training time as well as infinite capacity in both generator and discriminator allows to formulate the nash equilibrium in the generative adversarial network.
+Using a model with infinite training time as well as infinite capacity in both generator and discriminator allows formulating the nash equilibrium in the generative adversarial network.
 For any given generator $G$ and data points $x \sim p_{data}(x)$, the optimal discriminator $D$ is given by equation \ref{eq:opt_d}.
@@
 Most of the time during training will one network outperform the other player by such great margin that the gradients from the stronger player will vanish resulting in even slower training for the weaker network.
-Thus one of the common practices\cite{improved_gan:2016}\cite{dcgan:2015} for learning GANs is to freeze one network and prevent it from learning until the other player has catched up.
+Thus one of the common practices\cite{improved_gan:2016}\cite{dcgan:2015} for learning GANs is to freeze one network and prevent it from learning until the other player has caught up.
 %In practice the generator is far slower to train than the discriminator in most cases (citation).
 
-\subsubsection{Label Smoothing \cite{improved_gan:2016}}
+\subsubsection{Label Smoothing}
 \label{ssub:gan_label_smoothing}
-Another rather basic idea to improve gradient flow is to replace the binary classification from the discriminator with smoothed values $0.1/0.9$ or $\epsilon/1-\epsilon$.
+Another rather basic idea to improve gradient flow is to replace the binary classification from the discriminator with smoothed values $0.1/0.9$ or $\epsilon/1-\epsilon$\cite{improved_gan:2016}.
 This helps the generator with an informative gradient to improve learning performance.
@@
 With the recent advances in classification tasks performed by deep convolutional networks, automatically learned features have been a large part of that success.
-Feature matching is an technique designed to improve the learning procedure of GANs where $\mathcal{L}_g$ is not not directly maximized but where the discriminator also learns on intermediate layer activation outputs.
+Feature matching is an technique designed to improve the learning procedure of GANs where $\mathcal{L}_g$ is not directly maximized but where the discriminator also learns on intermediate layer activation outputs.
 Using this modified procedure aids the generator to output data matching the statistics of the input data more closely.
@@
 Deep convolutional networks\cite{lecun-89e} have yielded impressive results for image classification \cite{imagenet:2012} and have arguably been one of the main reasons for the recent success of deep learning.
-These techniques however have mostly been applied in a supervised learning context, not with unsupervised learning.
 
@@
 This leads to the suggestion that using deep convolutional networks helps boost the model performance.
-In fact is the deep convolution GAN proposed by Radford et al.\cite{dcgan:2015} one of the first GAN variants which results in visually pleasing results.
+In fact is the deep convolution GAN proposed by Radford et al.\cite{dcgan:2015} one of the first GAN variants which result in visually pleasing results.
 
-In order to use convolutional networks in the generator and discriminator, fractionally strided convolutions have been used to basically upsample the image.
-The exact details of the architecture is beyond the scope of this article, however in addition to the impressive generative performance the authors have shown that the model has learned to produce representional structures in latent space.
-Furthermore, as shown in Figure \ref{fig:dcgan_vector} they were able to perform vector artithmetic in the latent space with results indicating that their model was able to arrange the latent space in a useful way.
+In order to use convolutional networks in the generator and discriminator, fractionally strided convolutions (also called deconvolution~\cite{deconv:2010}) have been used to basically upsample the image.
+The exact details of the architecture is beyond the scope of this article, however in addition to the impressive generative performance the authors have shown that the model has learned to produce representational structures in latent space.
+Furthermore, as shown in Figure \ref{fig:dcgan_vector} they were able to perform vector arithmetic in the latent space with results indicating that their model was able to arrange the latent space in a useful way.
 %\todo[inline]{Elaborate on how DCGAN arranges latent space}

### 2017-03-03 paper/chapters/introduction.tex
last-minute changes

--- 2017-02-03 paper/chapters/introduction.tex
@@
 
-"What I cannot create, I do not understand." -- Richard Feynman \\\\
-
-Directed generative models are a wide class of models used for generating samples from an unknown distribution $p_{data}$.
-%Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
-The presented models in this article will approximate $p_{data}$ with $p_{model}$ which enables us to draw samples from the model.
-Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\
-
-\paragraph{Short-term applications}
-\begin{itemize}
-  \item \textbf{Data augmentation} is a technique to enrich an existing dataset by adding more derived information to provide more in-depth insight.
-  \item \textbf{Semi-supervised learning} on datasets with a small amount of unlabelled data and a large amount of labeled data to improve model performance.
-Examples include the conditional variational autoencoder discussed in \ref{ssub:cvae}.
-  \item \textbf{Data exploration and understanding} is often needed especially for high-dimensional and complex data.
-By using models which reduce the dimensions by selecting only important features for generative modeling can help to improve data understanding.
-Figure \ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
-\end{itemize}
-
-In addition to those short-term applications, there exist possibilities for more long-term aspects of generative modeling.
-For example with unsupervised representation learning, where statistical models are able to learn robust, accurate and generalized internal representations from high-dimensional sensory input can help power future agents to help understand real-world data, allowing these agents to take actions and reason about the real world.
-
-\newpage
-%Deep learning techniques have provided great success in applying gradient-based optimization methods to deep neural networks to achieve impressive results in discriminative models.
-%These models require however large amounts of labelled data, which is often expensive or infeasible to obtain.
-%Generative models can be used to learn \emph{representational} structures from rich, unlabelled data by having less parameters than the input data and are thus forced to capture only essential information.\\
-%\newpage
-
-%Prior to discussion individual models, we need to declare a few assumptions and definitions.
-%Additionally we assume that samples $x$ are also conditionally dependent on some hidden structure $z$: $x \sim p(x|z)$. 
-%\begin{equation}
-  %z \sim p(z)\\
-  %x \sim p(x|z)
-%\end{equation}
-
-%\paragraph{Sampling}
-%means for any given $z$ to draw a sample $x$ from our model.
-%Most commonly, sampling is the easiest requirement to fulfil.
-
-%\paragraph{Learning}
-%of the model parameters through maximum likelihood estimation generally involves $p(x)$ or its gradient.\\
-%To get $p(x)$, we first need to marginalize our the latent variables.
-%\begin{equation}
-  %p(x) = \sum_h p(x|z)p(z)
-%\end{equation}
-%Unfortunately, in the case of deep learning with many layers this is intractable, as we need to evaluate all model configurations.
-%via maximum likelihood generally involves the gradients of $p(x)$
-%\paragraph{Inference}
-% motivation / what we want exactly and why
-%Building intelligent systems requires having accurate and generalized information about the environment in order to perform actions or reason about the world.
-%Inferring structured and meaning information from unlabelled data, a process called unsupervised learning, remains a challenging task.
-
-%However, due to the massive increases in availability of data emerging from the web as well as the ongoing progress in computational power there has been significant progress using this data.
-
-%Generative modelling aims to generate new data similar to previously seen input, in a way which is sound with the internal structure of the data.
-
-%% on building reasonable representations
-%The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
-Models using deep learning often assume that observed data $x$ is stochastically dependent on some hidden structure $z$.
-To understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real-world data.
-
-\begin{itemize}
-  \item \textbf{Images} consisting of millions of pixels can be interpreted as data vectors with each pixel representing an own dimension.
-  However, these dimensions are highly correlated and structured, for example nearby pixels have with high probability similar values.
-  Looking at images of handwritten digits, the data can be reduced to a handful of variables which may include the digit number, stroke width, position or size.
-  Statistical models help us discover these relations and reduce them to representations in lower-dimensional space.
-  %\item commonly have thousands of dimensions, but the underlying structure is often far simpler.
-For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions.
-There are a few more dimensions, stroke width, cursive, position, size.
-But still far lower than 784.
-We will take a look at this example later on in \ref{sec:vae}.
-  \item \textbf{Audio data} is omni-present in the real world, however most data sources can be transformed from raw audio samples to spoken words or musical notes without losing too much information for reconstruction.
-  %\item \textbf{...}
-\end{itemize}
-% cope with
-%To develop systems learning from real world data we use statistical models which are able to infer the underlying structure from
-%Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
-%Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
-%Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\
-
-%With this assumption in mind we can take a look at some examples of input data.
-
-The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure \ref{fig:dgm}.
-%as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors.
-This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\
-
+%%%
+%%% intuition // generative vs discriminative
+%%%
+Intuitively, generative modeling is based on the idea that in order to produce data similar to samples from a dataset it helps to understand the data first~\cite[p.~720]{deeplearning:2016}.
+%The case of deep learning suggests the use of deep, hierarchical architectures to capture dependencies within the data and reduce the high-dimensional input~\cite{reducing_dim:2006}.
+In contrast to discriminative models which model the conditional distribution $p(y|x)$ of the labels $y$ given observations $x$, generative algorithms model the joint distribution $p(x,y)$ over both labels and observations to describe how the data was generated~\cite{learning_dbn:2006}.
+As there are only observations given, the generative task seems more difficult than classification where both input and desired output are given during training.~\cite[p.~695]{deeplearning:2016}\\\\
+%
+Due to the nature of generative models, they are unsupervised learning algorithms which mean that they try to learn the structure of samples from an unlabeled dataset\cite[p.~105-106]{deeplearning:2016}\cite{prob_unsupervised:1999}.
+%
+Generative models in machine learning using probabilistic models can be mostly separated into directed and undirected models~\cite[p.~77]{deeplearning:2016}.
+%
+This separation indicates how the model can be mapped to a graph with random variables as vertices and interactions between random variables as edges.
+Directed models induce a directed network graph while undirected models induce a network graph with undirected edges.\\\\
+%
+%
+In directed models, we assume that the observations $x$ are stochastically dependent on the underlying factors, also called latent variables $z$.
+This relationship can be expressed as the graphical model shown in Figure~\ref{fig:dgm}.
+Note that the latent variables can be structured hierarchically in order to express different levels of granularity in the features of the underlying factors and representations~\cite[p.~557]{deeplearning:2016}.
 \begin{figure}[htb]
 \centering
-\includegraphics{media/directed_graphical_model}
-  \caption{Directed graphical model}
+  \includegraphics[width=1.3cm]{media/directed_graphical_model}
+  \caption[Graphical Model]{Directed Graphical Model}
 \label{fig:dgm}
@@
   \small
-  The latent variable $z$ is a stochastic dependency of the observed data $x$.
+  The observed data $x$ is stochastically dependent on the latent variables $z$.\\
+  Based on~\cite[Chapter 8]{bishop:2006}.
 \end{figure}
 
+\subsection{Overview}
+% traditional models (HMM, GMM)
+Traditional models include hidden Markov models used for example in speech recognition~\cite{hmm:1989}, which are generative models~\cite[p.~878]{modern_approach:2009} and Gaussian mixture models~\cite[p.~190]{deeplearning:2016}\cite{schuster:1999}.\\\\
+%
+% undirected networks (Boltzmann machine)
+Undirected generative models include various extensions of the Boltzmann machine~\cite{bm:1985} such as the restricted Boltzmann machine (RBM), first described as Harmonium~\cite{smolensky:1986} and the deep Boltzmann machine~\cite{dbm:2009}\cite[Chapter~20.4]{deeplearning:2016}.\\\\
+%
+% directed networks (SBN, DBN)
+For directed models, the sigmoid belief network (SBN) is an early model described in 1992~\cite{neal:1992} which is presented in Section~\ref{sec:sbn}.
+SBN has also been extended with a restricted Boltzmann machine for the first two layers resulting in deep belief nets (DBN)~\cite{learning_dbn:2006}, which have been attributed to spark the beginning of current deep learning~\cite[p.~662]{deeplearning:2016}.
+DBN are partially directed models, as they contain both directed and undirected connections~\cite[p.~664]{deeplearning:2016}.\\\\
+%
+%Another class of generative models are auto-regressive networks (ARN), which 
+%\newpage
+%%%
+%%% applications
+%%%
+
+\subsection{Applications}
+
+\begin{figure}[t]
+  \includegraphics[width=\linewidth]{media/caption2img}
+  \caption[Caption-to-Image Samples]{Generative Text to Image Synthesis~\cite{gan_t2i:2016}}
+  \medskip
+  \small
+  The top row shows the provided captions and in the bottom row the corresponding synthesized images.
+\label{fig:caption2img}
+\end{figure}
+
+Generative models learn the joint distribution $p(x,y)$ and due to this fact they can be used to perform classification as well by expand $p(x,y)$ to $p(y|x)$ as indicated by Equation~\ref{eq:class}.
+\begin{equation}
+  \label{eq:class}
+  p(y|x) = \frac{p(x,y)}{p(x)}
+\end{equation}
+%
+One obvious application of generative models is to use the model to sample more data similar to samples from a given training dataset.\\
+%
+In computer vision tasks, generative models have been largely outperformed by much simpler models until at least 2011~\cite{thesis:2011}.
+In 2015 however, convolutional networks have been shown to surpass those models in image synthesis~\cite{synthesis:2015}.
+Other tasks involving image modeling include performing super-resolution~\cite{superres:2017}, constructing 3D models from 2D images~\cite{gan_3d:2016}, predicting the next video frames~\cite{gan_video:2016}, and synthesizing new images from captions as shown in Figure~\ref{fig:caption2img}.\\
+%
+In the area of reinforcement learning, generative models offer a way to perform action space exploration~\cite{rl_variational:2016}.
+Generative adversarial networks discussed in Section~\ref{sec:gan} have been shown to be used in imitation learning, where the generative model is used to imitate previously seen interactions~\cite{imitation_learning:2016}.\\
+%
+SampleRNN~\cite{samplernn:2016}, which is a recurrent neural network combined with an auto-regressive network have been shown to produce good generated audio samples.\\\\
+%
+Recently, generative models have also shown to improve semi-supervised learning performance significantly~\cite{cvae:2014}.
+Semi-supervised learning algorithms use only small amount of labeled data, which is usually difficult to obtain and large amounts of unlabeled data, which is easier to obtain in most cases.
+
 \newpage
-%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.
-
-%After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two recent promising frameworks, the variational autoencoder (VAE \ref{sec:vae}) and the generative adversarial network (GAN \ref{sec:gan}).
-
-%Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
-%More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
-%This assumption is reasonable given the following examples:\\
-
-%Generative models have been proposed as one of the most promising approaches towards
-%learning representations of real world data by some of the leading researchers (YannLeCun, OpenAI blogpost).
-
-%In recent years supervised learning has yielded impressive results,
-%however for these models to succeed huge amount of labelled data is needed.
-%Unsupervised learning, that is learning from unlabelled data, has been expressed
-%as one hurdles toward general articial intelligence (citation needed).
-%Early approaches toward this problem however have shown that these problems are very hard to learn (intractable?).
-
-%In generative models the approach is different in a way that
-%we search for a good internal representation of the data.
-%This resembles the way humans learn about the world where we incrementally build
-%a model of the world.

### 2017-03-03 paper/chapters/sigmoid_belief_nets.tex
last-minute changes

@@
   Bayesian networks, also called belief nets are an acyclic directed graph where blue nodes indicate hidden variables and red nodes represent observed variables.
-  Arrows indicate stochastic dependency.
-  Even though most belief nets are divided into many layers in the context of deep learning, the original definition does not specify that and thus indicates the dashed arrow the possibility of inter-layer dependencies.
+  Even though most belief nets are divided into many layers in the context of deep learning~\cite[Chapter 20.10.1]{deeplearning:2016}, the original definition does not specify that and thus indicates the dashed arrow the possibility of inter-layer dependencies.
 \end{figure}
@@
 In the case of SBNs however, all states are binary and the activation function is the sigmoid function.
-They are one of the earliest neural networks used for generative modeling, predating all other presented models in this article.
+They are one of the earliest neural networks used for generative modeling, predating all other presented directed models in this article.
 %As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
@@
 To draw samples from this network, it is possible to evaluate Equation~\ref{eq:sbn_node} for each visible node.
-However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
+However, prior to that the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem~\cite[p.~695]{deeplearning:2016}.\\
 Learning means in this context to adjust the weights and biases of the network to be more likely to generate data similar to samples from the given training dataset.
-Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the states in the visible layer.\\
-
-Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
+Meanwhile, statistical inference is tasked to infer the states of the hidden nodes based on the states in the visible layer.\\
+%
+Due to each state being statistically dependent on all ancestors, inference means to evaluate all possible configurations of every previous node.
 Especially in deep networks, this kind of computation is infeasible as it requires computing exponentially many configurations.

### 2017-03-03 paper/chapters/vae.tex
last-minute changes

@@
 Using gradient-based optimization techniques enables the encoder network to be trained to output distributions similar to the true posterior $p(z|x)$.
-In addition to this inference network there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.
+In addition to this inference network, there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.
 
@@
 Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
-This architecture is shown in Figure \ref{fig:vae_architecture} with a one-dimensonal gaussian prior.
+This architecture is shown in Figure \ref{fig:vae_architecture} with a one-dimensional Gaussian prior.
 \\
@@
 \subsection{Approximated Inference}
-A problem which often arises in directed generative models is the \emph{inference problem}.
-As discussed in the introduction~\ref{sec:introduction}, we assume that the observations are stochastically dependent on some latent variables.
+A problem which often arises in directed generative models is the \emph{inference problem}~\cite[Chapter~16.6]{deeplearning:2016}.
+As discussed in \ref{sec:introduction}, we assume that the observations are stochastically dependent on some latent variables.
 Inference is called the process of coming up with values of $z$ which can produce the observations, which formally means evaluating $p(z|x)$ where $z$ are the latent variables and $x$ observations.
@@
 % variational inference
-Due to the possibly slow convergence of MCMC methods, variational inference (VI)~\cite[Chapter~19.4]{deeplearning:2016} is used in the variational autoencoder.
+Due to the possibly slow convergence of MCMC methods, variational inference (VI)~\cite[Chapter~19.4]{deeplearning:2016}\cite[Chapter~10.1]{bishop:2006} is used in the variational autoencoder~\cite{vae:2014}.
 VI works by picking a specific distribution restricted to a family of distributions which matches the posterior most closely.
-This means that VI is used to turn the inference problem into an optimization technique, allowing for gradient-based optimization.
+In the case of the VAE, we use neural networks to select the parameters of the distributions to turn the inference problem into an optimization problem, allowing for gradient-based optimization.
+Due to the restriction of VI to a specific probability distribution class and neural networks as non-convex function approximators, it is not guaranteed to converge.
+%Variational inference is however restricted to the specific class of distributions, which means that while we are able to approximate distribution in a given family (e.g.
+normal distributions), if the target distribution is not in that family there is no way of convergence.
 %TODO: more on advantages and disadvantages
-Using VI yields better performance at the sacrifice of theoretical guarantee of matching the posterior distribution.
+Conclusively, even though MCMC offers approximation with guaranteed convergence~\cite[Chapter~11.2]{bishop:2006}, in practice it is often slow to converge while VI offers reasonable good approximation at greater speed~\cite{vae:2014}.
 
-More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are the hidden and $x$ the observed variables into an optimization problem, where we can use gradient-based methods.
+\newpage
 
 \subsubsection{Kullback-Leibler Divergence}
-In order to approximate one distribution $Q$ to another distribution $P$, there needs to be a measurement of how close $Q$ is to $P$.
-One commonly used measurement is called Kullback-Leibler (KL) divergence\cite{kl_div:1951} which enables the variational inference methods to choose appropiate distributions.
+In order to approximate one distribution $Q$ to another distribution $P$, we need to have a measurement of how close $Q$ is to $P$.
+One commonly used measurement is called Kullback-Leibler (KL) divergence~\cite{kl_div:1951} which enables the variational inference methods to choose appropriate distributions.
 %\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions.
@@
 \end{equation}
-\todo[inline]{Add some description to the equation above, logarithmic difference between P and Q, bla bla expectation}
+%\todo{Add some description to the equation above, logarithmic difference between P and Q, bla bla expectation}
 %For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
@@
 $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the variational inference.
-
-\newpage
 
@@
 
-However we are not able to optimize Equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
+However, we are not able to optimize Equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
 The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
@@
 
-\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative KL divergence between the learned variational distribution $q_\phi(z|x)$ and the latent space prior distribution $p(z)$.
+% excluded due to no references found :/
+%\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative KL divergence between the learned variational distribution $q_\phi(z|x)$ and the latent space prior distribution $p(z)$.
 
@@
 
-Using the example of handwritten digits it appears why this regularizer term $\mathcal{L}_{reg}$ is important for the model to learn generalized representations in latent space.
-Suppose we have two data points from the dataset which both represent the same handwritten digit.
-Ideally, our model would learn representations in latent space to be fairly close to ensure the model has learned to generalize the semantics.
+%Using the example of handwritten digits it appears why this regularizer term $\mathcal{L}_{reg}$ is important for the model to learn generalized representations in latent space.
+%Suppose we have two data points from the dataset which both represent the same handwritten digit.
+%Ideally, our model would learn representations in latent space to be fairly close to ensure the model has learned to generalize the semantics.
 Without $\mathcal{L}_{reg}$ the model would assign each data example a different representation in latent space with a distribution collapsing to that single point.
@@
 
-%Suppose two data points with similar input
+%%Suppose two data points with similar input
 %Without $\mathcal{L}_{reg}$ the encoder would encode data points we want to have similar representations, fo
@@
 
-\vspace{2cm}
+%\vspace{2cm}
 
-\paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
-This error is measured using the negative log-likelihood of the conditional probability distribution $p_\theta(x|z)$ where $z$ is sampled from $q_\phi(z | x)$.\\
-$\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
+%\paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
+%This error is measured using the negative log-likelihood of the conditional probability distribution $p_\theta(x|z)$ where $z$ is sampled from $q_\phi(z | x)$.\\
+%$\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.
 
@@
 However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
-This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
-In order to circumvent this restriction, the reparameterization trick (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) can be applied.
+This poses a problem because the gradient can not be computed accurately due to the added noise.\\
+In order to circumvent this restriction, the reparameterization trick~\cite{vae:2014}\cite{dlgm:2014}\cite[chapter~20.9]{deeplearning:2016} can be applied.
 Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$.
@@
 Figure \ref{fig:rep_trick} expresses the reparameterization trick applied to an arbitrary function $f$ with the parameters $\theta$.\\
-Using this trick, the whole VAE model can be trained using gradient-based optimization techniques.\\\\
+Using this trick, the whole VAE model can be trained using gradient-based optimization techniques.
 
-Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different prior probability distributions on $z$.
+%Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different prior probability distributions on $z$.
+
 %The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
@@
 To allow analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.
-
-%To allow the computation of the KL divergence between $q_\theta(z|x)$ and the latent prior distribution $p(z)$ in closed form, the prior is an isotropic gaussian $\mathcal{N}(0, 1)$.
-
+The learning procedure of a variational autoencoder is shown in Algorithm~\ref{alg:vae} which is a simplified version of the algorithm in the original paper~\cite{vae:2014}.
 \begin{algorithm}
@@
 \end{algorithm}
+%
+\subsection{Experiments}
+In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained using TensorFlow~\cite{tf:2015} on the MNIST dataset implementing a fully-connected network.%\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}.
+Both experiments as shown in Figure~\ref{fig:vae_experiments} have been performed with two hidden layers of size 500.
+Rectified linear units~\cite{relu:2000} were used as activation functions in this network.
 
-\subsection{Experiments}
-In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}.
-Results of this experiments are shown in Figure \ref{fig:vae_experiments}.
 %The model has been trained on the MNIST dataset \cite{mnist} with satisfactory generative performance as shown in figure \ref{fig:vae_experiments}
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[h!]
     \centering
@@
         \includegraphics[height=7cm]{media/vae_manifold.pdf}
-        \caption{Learned manifold using 2-dimensional latent space}
+        \caption{Learned manifold}
     \end{subfigure}%
@@
         \includegraphics[height=7cm]{media/vae_samples.pdf}
-        \caption{Generated images by sampling from latent space}
+        \caption{Randomly generated images by sampling}
     \end{subfigure}
@@
     The shown manifold has been plotted using samples uniformly samples in the range $[-3,-3]^2$ and then reconstructed using the decoder network $p_\theta(x|z)$.
-The results have shown that the model has learned to arrange the latent space in a sensible way by placing digits with high similarity near each other.\\
-    (b) generative performance qualitively evaluated using samples in the data space by random sampling $z \sim p(z)$ and decoding $z$ using the decoder network.
-For this experiment an 50-dimensional latent space has been used.
+    (b) generative performance qualitatively evaluated using samples in the data space by random sampling $z \sim p(z)$ and decoding $z$ using the decoder network.
+For this experiment a 50-dimensional latent space has been used.
 \end{figure*}

### 2017-03-03 paper/chapters/vae/cvae.tex
last-minute changes

@@
-Instead of using VAEs in a completely unsupervised manner on a unlabeled dataset, it is also possible to learn from datasets where a small subset of datapoints has corresponding labels while most datapoints are still unlabeled.
+Instead of using VAEs in a completely unsupervised manner on an unlabeled dataset, it is also possible to learn from datasets where a small subset of data points has corresponding labels while most data points are still unlabeled.
 Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly~\cite{cvae:2014}\cite{cvae_2:2015}.\\
@@
 
-The final objective function for whole dataset is the summation between right-hand terms of equations \ref{eq:cvae_labels} for the labelled subset and \ref{eq:cvae_unknown} for the unlabelled subset.
+The final objective function forthe  whole dataset is the summation between right-hand terms of equations \ref{eq:cvae_labels} for the labeled subset and \ref{eq:cvae_unknown} for the unlabelled subset.

### 2017-03-03 paper/chapters/vae/iwae.tex
remove dumb emphasis
last-minute changes

@@
 % write about IWAE, extends these k passes with importance weights
-Importance Weighted Autoencoders (IWAE\cite{iwae:2015}) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
+The Importance Weighted Autoencoder\cite{iwae:2015} (IWAE) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
 % add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
@@
 
-Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum as $k$ approaches infinity in equation \ref{eq:iwae}
+Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimum as $k$ approaches infinity in equation \ref{eq:iwae}
 \begin{equation}
